---
title: "raw_data_cleaning"
output: html_document
date: "2025-03-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
```

```{r}
raw_data <- read.delim(file = "../../data/raw_data/neuropeptides_KW-0527.tsv", sep="\t", na = c("", "NA", "#N/A"))
```

```{r}
names(raw_data)
```
```{r}
raw_data %>%
  distinct(Subcellular.location..CC.)
```

Filtering out receptors:
```{r}
raw_data_no_recep <- raw_data %>%
  filter(!str_detect(Protein.names, "receptor")) 
```

# extracting all possible localizations:
```{r}
all_locations <- raw_data_no_recep %>%
  mutate(
    extracted_locations = str_extract_all(Subcellular.location..CC., "(?<=SUBCELLULAR LOCATION: |\\.\\s)[^\\{]+") # Extract locations
  ) %>%
  unnest(extracted_locations) %>%
  mutate(
    extracted_locations = str_remove_all(extracted_locations, "Note=.*"),  # Remove "Note="
    extracted_locations = str_replace_all(extracted_locations, "\\[.*?\\]:\\s*", ""),  # Remove "[text]:"
    extracted_locations = str_replace_all(extracted_locations, "SUBCELLULAR LOCATION:\\s*", ""), # Properly remove "SUBCELLULAR LOCATION:"
    extracted_locations = str_split(extracted_locations, "[.,;]")  # Split locations by ".", "," or ";"
  ) %>%
  unnest(extracted_locations) %>%
  mutate(
    extracted_locations = str_trim(extracted_locations), # Remove leading/trailing spaces
    extracted_locations = na_if(extracted_locations, ""), # Remove empty locations
    extracted_locations = na_if(extracted_locations, "NA") # Remove "NA" as a location
  ) %>%
  distinct(Entry.Name, extracted_locations) %>% # Ensure unique locations per ID
  na.omit()

expanded_df <- all_locations %>%
  pivot_wider(
    names_from = extracted_locations,
    values_from = extracted_locations,
    values_fn = function(x) TRUE,  # Set detected locations to TRUE
    values_fill = list(extracted_locations = NA)  # Fill others with NA
  )

head(expanded_df)
```

```{r}
expanded_df <- expanded_df %>%
  select(where(~ sum(!is.na(.)) >= 10))

expanded_df %>%
  summarise_all(~ sum(!is.na(.)))
```

# Note: the below code needs to be revised. there are multiple isoforms of the peptides and therefore, multiple localizations within the cell.
```{r}
clean_data <- raw_data %>%
  mutate(subcellular_location = ifelse(str_detect(Subcellular.location..CC., "LOCATION: Secreted"), "Secreted", Subcellular.location..CC.)) %>%
  select(-Subcellular.location..CC.) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Cell membrane"), "Cell membrane", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: [Isoform 1]: Cell membrane"), "Cell membrane", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Cytoplasmic vesicle"), "Cytoplasmic vesicle", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Rough endoplasmic reticulum"), "Endoplasmic reticulum", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Endoplasmic reticulum"), "Endoplasmic reticulum", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Cytoplasm"), "Cytoplasm", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Membrane"), "Membrane", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Nucleus"), "Nucleus", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Mitochondrion"), "Mitochondrion", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Endomembrane"), "Endomembrane", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Peroxisome"), "Peroxisome", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "LOCATION: Apical cell membrane"), "Apical cell membrane", subcellular_location)) %>%
  mutate(subcellular_location = ifelse(str_detect(subcellular_location, "Secreted"), "Secreted", subcellular_location)) 


clean_data %>%
  distinct(subcellular_location) %>%
  pull()
```

```{r}
clean_data <- raw_data_no_recep %>%
  select(-Subcellular.location..CC.) %>%
  left_join(expanded_df)
```


```{r}
clean_data <- clean_data %>%
  mutate(protein_names = str_extract(Protein.names, "^[^\\[]+")) %>%
  select(-Protein.names)
```
Looking at all the values in the protein_names column
```{r}
clean_data %>%
  distinct(protein_names)
```
# making the table longer by putting all locations into a column called "Localization"
```{r}
clean_data <- clean_data %>%
  pivot_longer(
    cols = Secreted:Membrane,  
    names_to = "Localization",  
    values_to = "Present"  # Temporary column to store TRUE/NA
  ) %>%
  filter(Present == TRUE) %>%  # Keep only rows where location is present
  select(-Present)  # Drop the "Present" column
```

# let's look the number of values per column and delete any columns with less than 10 entries. 
```{r}
clean_data <- clean_data %>%
  select(where(~ sum(!is.na(.)) >= 10))

clean_data %>%
  summarise_all(~ sum(!is.na(.)))
```

# Let's look at the function column
```{r}
clean_data %>%
  group_by(Gene.Ontology..molecular.function.) %>%
  summarise(n = n())
```

```{r}
clean_data_2 <- clean_data %>%
  mutate(Gene.Ontology..molecular.function. = str_split(Gene.Ontology..molecular.function., ";\\s*")) %>%  # Split by ";"
  unnest(Gene.Ontology..molecular.function.) %>%  # Expand into separate rows
  mutate(Gene.Ontology..molecular.function. = str_remove(Gene.Ontology..molecular.function., "\\[GO:\\d+\\]")) %>%  # Remove [GO:xxxxx]
  rename(molecular_function_go = Gene.Ontology..molecular.function.) %>%  # Rename column
  mutate(molecular_function_go = str_trim(molecular_function_go))  # Remove any extra spaces

clean_data_2 %>%
  select(molecular_function_go) %>%
  group_by(molecular_function_go) %>%
  summarise(n = n())
```

# repeating the same process with gene ontology biological process

```{r}
clean_data_3 <- clean_data_2 %>%
  mutate(Gene.Ontology..biological.process. = str_split(Gene.Ontology..biological.process., ";\\s*")) %>%  # Split by ";"
  unnest(Gene.Ontology..biological.process.) %>%  # Expand into separate rows
  mutate(Gene.Ontology..biological.process. = str_remove(Gene.Ontology..biological.process., "\\[GO:\\d+\\]")) %>%  # Remove [GO:xxxxx]
  rename(biological_process_go = Gene.Ontology..biological.process.) %>%  # Rename column
  mutate(biological_process_go = str_trim(biological_process_go))  # Remove any extra spaces

clean_data_3 %>%
  select(biological_process_go) %>%
  group_by(biological_process_go) %>%
  summarise(n = n())
```

# Now let's look at the keywords column
```{r}
clean_data_4 <- clean_data_3 %>%
  mutate(Keywords = str_split(Keywords, ";\\s*")) %>%  # Split by ";"
  unnest(Keywords) %>%  # Expand into separate rows
  mutate(Keywords = str_trim(Keywords))  # Remove any extra spaces

clean_data_4 %>%
  select(Keywords) %>%
  group_by(Keywords) %>%
  summarise(n = n())
```

$ looking at protein families
```{r}
raw_data_no_recep %>%
  select(Protein.families) %>%
  group_by(Protein.families) %>%
  summarise(n = n())
```
# Now let's look at the function column:
```{r}
clean_data_4 <- clean_data_4 %>%
  

function_df <- raw_data_no_recep %>%
  select(Function..CC.) %>%
  group_by(Function..CC.) %>%
  summarise(n = n())
```

Cleaning function column
```{r}
clean_data_5 <- clean_data_4 %>%
  mutate(Function..CC. = str_remove(Function..CC., "^FUNCTION:\\s*")) %>%  # Remove "FUNCTION:"
  mutate(Function..CC. = str_remove(Function..CC., "\\{ECO:.*?\\}")) %>%  # Remove "{ECO:...}"
  mutate(Function..CC. = str_trim(Function..CC.)) %>%  # Remove leading/trailing spaces
  rename(Function = Function..CC.) # Rename column for clarity

clean_data_5 %>%
  distinct(Function) 

clean_data_5 <- clean_data_5 %>%
  filter(Protein.families != "G-protein coupled receptor 1 family") %>%
  filter(Length < 600) 
```

# Now let's get as many datasets as we can.

Go terms (molecular function)
```{r}
grouped_go_molecular_function <- clean_data_5 %>%
  select(Sequence, molecular_function_go) %>%
  distinct() %>%
  na.omit() %>%
  group_by(molecular_function_go) %>%
  count() %>%
  filter(n > 49) %>%
  rename("biological_concept" = molecular_function_go,
         "sequence_count" = n)
```



Go terms (biological process)
```{r}
grouped_go_bio_process <- clean_data_5 %>%
  select(Sequence, biological_process_go) %>%
  distinct() %>%
  na.omit() %>%
  group_by(biological_process_go) %>%
  count() %>%
  filter(n > 49) %>%
  rename("biological_concept" = biological_process_go,
         "sequence_count" = n)
```

Protein family
```{r}
grouped_protein_family <- clean_data_5 %>%
  select(Sequence, Protein.families) %>%
  distinct() %>%
  na.omit() %>%
  group_by(Protein.families) %>%
  count() %>%
  filter(n > 49) %>%
  rename("biological_concept" = Protein.families,
         "sequence_count" = n)
```

Signal peptide present 
```{r}
count_signal_peptide <- clean_data_5 %>%
  select(Sequence, Signal.peptide) %>%
  distinct() %>%
  mutate(signal = ifelse(!is.na(Signal.peptide), TRUE, FALSE)) %>%
  group_by(signal) %>%
  count() %>%
  rename("signal_present" = signal,
         "sequence_count" = n)
  
```

Now grouping by keywords
```{r}
grouped_keywords <- clean_data_5 %>%
  select(Sequence, Keywords) %>%
  distinct() %>%
  na.omit() %>%
  group_by(Keywords) %>% 
  count() %>%
  filter(n > 49) %>%
  rename("biological_concept" = Keywords,
         "sequence_count" = n) %>%
  filter(!biological_concept %in% c("Reference proteome", "Direct protein sequencing", "Repeat")) 
  
```

Now, let's group by localization:
```{r}
grouped_localization <- clean_data_5 %>%
  select(Sequence, Localization) %>%
  distinct() %>%
  na.omit() %>%
  group_by(Localization) %>% 
  count() %>%
  #filter(n > 49) %>% 
  rename("biological_concept" = Localization,
         "sequence_count" = n)

```


Now let's make figures for each of these groups. (barplots)
```{r}
mol_func_plot <- grouped_go_molecular_function %>%
  ggplot(aes(x = sequence_count, y = fct_reorder(biological_concept, sequence_count))) +
  geom_col(fill = "maroon") +
  scale_x_continuous(
    name = "Number of sequences",
    limits = c(0, 4000),
    breaks = c(0, 1000, 2000, 3000, 4000),
    expand = expansion(mult = c(0, 0.13)),
  ) +
  scale_y_discrete(
    name = NULL,
    expand = expansion(mult = c(0, 0))
  ) +
  labs(title = "GO Terms: Molecular Function") +
  geom_text(aes(label = sequence_count),
            hjust = -0.1,
            vjust = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0),
        panel.grid.major.y = element_blank(),  
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(color = "grey70", size = 0.1),
        panel.grid.major.x = element_line(color = "grey70", size = 0.1),
        #axis.line.x.bottom = element_line(color = "grey70", size = 0.15)
        ) 

mol_func_plot
ggsave(filename = "../../figures/mol_func_plot.png", plot = mol_func_plot, width = 9, height = 5)
```
```{r}
bio_proc_plot <- grouped_go_bio_process %>%
  ggplot(aes(x = sequence_count, y = fct_reorder(biological_concept, sequence_count))) +
  geom_col(fill = "maroon") +
  scale_x_continuous(
    name = "Number of sequences",
    limits = c(0, 12000),
    breaks = c(0, 2000, 4000, 6000, 8000, 10000, 12000),
    expand = expansion(mult = c(0, 0.13)),
  ) +
  scale_y_discrete(
    name = NULL,
    expand = expansion(mult = c(0, 0))
  ) +
  labs(title = "GO Terms: Biological Process") +
  geom_text(aes(label = sequence_count),
            hjust = -0.1,
            vjust = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0),
        panel.grid.major.y = element_blank(),  
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(color = "grey70", size = 0.1),
        panel.grid.major.x = element_line(color = "grey70", size = 0.1),
        #axis.line.x.bottom = element_line(color = "grey70", size = 0.15)
        ) 

bio_proc_plot
ggsave(filename = "../../figures/bio_proc_plot.png", plot = bio_proc_plot, width = 7, height = 8)
```
```{r}
prot_fam_plot <- grouped_protein_family %>%
  ggplot(aes(x = sequence_count, y = fct_reorder(biological_concept, sequence_count))) +
  geom_col(fill = "maroon") +
  scale_x_continuous(
    name = "Number of sequences",
    expand = expansion(mult = c(0, 0.13)),
  ) +
  scale_y_discrete(
    name = NULL,
    expand = expansion(mult = c(0, 0))
  ) +
  labs(title = "Protein Family") +
  geom_text(aes(label = sequence_count),
            hjust = -0.1,
            vjust = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0),
        panel.grid.major.y = element_blank(),  
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(color = "grey70", size = 0.1),
        panel.grid.major.x = element_line(color = "grey70", size = 0.1)
        ) 

prot_fam_plot
ggsave(filename = "../../figures/prot_fam_plot.png", plot = prot_fam_plot, width = 7, height = 4)
```

```{r}
keyword_plot <- grouped_keywords %>%
  ggplot(aes(x = sequence_count, y = fct_reorder(biological_concept, sequence_count))) +
  geom_col(fill = "maroon") +
  scale_x_continuous(
    name = "Number of sequences",
    expand = expansion(mult = c(0, 0.13)),
    limits = c(0, 12000),
    breaks = c(0, 2000, 4000, 6000, 8000, 10000, 12000)
  ) +
  scale_y_discrete(
    name = NULL,
    expand = expansion(mult = c(0, 0))
  ) +
  labs(title = "Uniprot Keywords") +
  geom_text(aes(label = sequence_count),
            hjust = -0.1,
            vjust = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0),
        panel.grid.major.y = element_blank(),  
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(color = "grey70", size = 0.1),
        panel.grid.major.x = element_line(color = "grey70", size = 0.1)
        ) 

keyword_plot
ggsave(filename = "../../figures/keyword_plot.png", plot = keyword_plot, width = 7, height = 4)
```


```{r}
local_plot <- grouped_localization %>%
  ggplot(aes(x = sequence_count, y = fct_reorder(biological_concept, sequence_count))) +
  geom_col(fill = "maroon") +
  scale_x_continuous(
    name = "Number of sequences",
    expand = expansion(mult = c(0, 0.13)),
    limits = c(0, 12000),
    breaks = c(0, 2000, 4000, 6000, 8000, 10000, 12000)
  ) +
  scale_y_discrete(
    name = NULL,
    expand = expansion(mult = c(0, 0))
  ) +
  labs(title = "Localization") +
  geom_text(aes(label = sequence_count),
            hjust = -0.1,
            vjust = 0.5) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0),
        panel.grid.major.y = element_blank(),  
        panel.grid.minor.y = element_blank(),
        panel.grid.minor.x = element_line(color = "grey70", size = 0.1),
        panel.grid.major.x = element_line(color = "grey70", size = 0.1)
        ) 

local_plot
ggsave(filename = "../../figures/local_plot.png", plot = local_plot, width = 7, height = 4)
```

# Turning the clean_data_5 dataset into a fasta
```{r}
for_fasta <- clean_data_5 %>%
  select(Entry, Sequence) %>%
  distinct(Sequence, .keep_all = TRUE)

# Create FASTA-formatted lines: ">ID" on one line, followed by the sequence on the next
fasta_lines <- apply(for_fasta, 1, function(row) {
  paste0(">", row["Entry"], "\n", row["Sequence"])
})

# Write these lines to a .fasta file
writeLines(fasta_lines, con = "../../data/neuropeptide_sequences.fasta")
```

# let's look at the deeploc results and compare them to the Uniprot localization annotations.

```{r}
deeploc_data <- read.csv(file = "../../data/combined_deeploc_results.csv")

```


Now, let's do an inner join between clean_data_5 and the deeploc data:
```{r}
loc_df <- clean_data_5 %>%
  select(Entry, Localization) %>%
  distinct() %>%
  rename("uniprot_localization" = Localization)

loc_df %>%
  group_by(uniprot_localization) %>%
  count()
```

```{r}
deeploc_df <- deeploc_data %>%
  select(Protein_ID, Localizations) %>%
  rename("Entry" = Protein_ID,
         "deeploc_localization" = Localizations)

deeploc_df %>%
  group_by(deeploc_localization) %>%
  count()

```

Now joining the datasets together:
```{r}
joined_loc <- inner_join(deeploc_df, loc_df)
```

First, I want to separate multiple localization values in the deeploc column
```{r}
joined_loc <- joined_loc %>%
  separate_rows(deeploc_localization, sep = "\\|") %>%
  mutate(deeploc_localization = str_trim(deeploc_localization))  # optional: remove surrounding whitespace

joined_loc %>%
  group_by(deeploc_localization) %>%
  count()
```
Now I need to change some of the entries so that they match between the deeploc and uniprot data. 
```{r}
joined_loc_2 <- joined_loc %>%
  mutate(uniprot_localization = ifelse(uniprot_localization == "Secreted", "Secreted/Extracellular", uniprot_localization)) %>%
  mutate(deeploc_localization = ifelse(deeploc_localization == "Extracellular", "Secreted/Extracellular", deeploc_localization)) %>%
  mutate(uniprot_localization = ifelse(uniprot_localization == "Membrane", "Cell membrane", uniprot_localization)) %>%
  mutate(uniprot_localization = ifelse(uniprot_localization == "Rough endoplasmic reticulum", "Endoplasmic reticulum", uniprot_localization))

joined_loc_2 %>%
  filter(uniprot_localization != deeploc_localization) %>%
  distinct(uniprot_localization, deeploc_localization)
```

Now let's plot a heatmap of the predicted and annotated localizations:
```{r}
for_heat_plot <- joined_loc_2 %>%
  group_by(deeploc_localization, uniprot_localization) %>%
  summarize(count = n(), .groups = "drop")

for_heat_plot_2 <- for_heat_plot %>%
  group_by(uniprot_localization) %>%
  mutate(freq = count/sum(count))

for_heat_plot_2 <- for_heat_plot_2 %>%
  mutate(
    deeploc_localization = factor(deeploc_localization),
    uniprot_localization = factor(uniprot_localization))

for_heat_plot_2 <- for_heat_plot_2 %>%
  mutate(
    deeploc_localization = fct_relevel(deeploc_localization, "Secreted/Extracellular", "Cell membrane", "Endoplasmic reticulum", "Nucleus"),
    uniprot_localization = fct_relevel(uniprot_localization, "Secreted/Extracellular",  "Cell membrane", "Endoplasmic reticulum", "Nucleus")) 
```

```{r}
# Get the levels in order (so we can reference the lower-left corner)
x_levels <- levels(for_heat_plot_2$deeploc_localization)
y_levels <- levels(for_heat_plot_2$uniprot_localization)

# Pick the 4 lowest levels (leftmost and bottommost)
x_rect <- x_levels[1:4]
y_rect <- y_levels[1:4]


heatmap_loc_plot <- for_heat_plot_2 %>%
  ggplot(aes(x = deeploc_localization, y = uniprot_localization, fill = freq)) +
  geom_tile(color = "white") +  
  geom_text(aes(label = count), color = "white", size = 3) +
  scale_x_discrete(
    name = "DeepLoc localization prediction",
    expand = expansion(add = c(0,0))
  ) +
  scale_y_discrete(
    name = "Uniprot localization annotation",
    expand = expansion(add = c(0,0))
  ) +
  scale_fill_gradient(
    low = "#b2dfdb",   # for low counts
    high = "#014d4e",   # for high counts (darkest)
    name = "Frequency"
  ) +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("rect",
           xmin = 0.5, xmax = 4.5,   # first 4 x-levels
           ymin = 0.5, ymax = 4.5,   # first 4 y-levels
           fill = NA,                # no fill
           color = "maroon",
           size = 0.6) 

heatmap_loc_plot

ggsave(filename = "../../figures/local_heatmap_plot.png", plot = heatmap_loc_plot, width = 8, height = 6)
```

I want to see how many neuropeptides are less than 600 residues in length
```{r}
clean_data_5 %>%
  select(Sequence) %>%
  distinct() %>%
  count()
```
I need to plot a histogram of peptide size length, so that I can then sample the non-neuropeptides to have the same size distribution.


```{r}
for_hist_plot <- clean_data_5 %>%
  select(Sequence, Length) %>%
  distinct()

hist_plot <- for_hist_plot %>%
  ggplot(aes(x = Length)) +
  geom_histogram(
    binwidth = 50,
    center = 25,
    color = "gray55",
    fill = "cornsilk") +
  scale_x_continuous(
    expand = expansion(mult = c(0,0.01))
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0,0.1))
  ) +
  geom_text(
    stat = "bin",
    binwidth = 50,
    center = 25,
    aes(label = ..count..),
    vjust = -0.5  # move text slightly above the bar
  ) +
  theme_cowplot()

hist_plot
```
Now I need a file with the counts per bin:
```{r}

binned_counts <- for_hist_plot %>%
  mutate(
    bin = cut(
      Length,
      breaks = seq(0, 600, by = 50),  # start at 0 to align with center=25
      include.lowest = TRUE,
      include.highest = TRUE,
      right = FALSE,  # [0,50), [50,100), etc.
      labels = paste(seq(0, 550, by = 50), seq(49, 599, by = 50), sep = "-")
    )
  ) %>%
  count(bin, name = "sequence_count")

print(binned_counts)

write_csv(binned_counts, "../../data/sequence_counts_per_bin.csv")
```

```{r}
sum <- binned_counts %>%
  mutate(sum = sum(sequence_count))

sum
```

Lets get our first concept positive training group.
The first group will be neuropeptide hormone activity (broad)

```{r}
neuropeptide_hormone_activity <- clean_data_5 %>%
  select(Entry, Entry.Name, Sequence, Length, molecular_function_go) %>%
  distinct() %>%
  filter(molecular_function_go == "neuropeptide hormone activity")

```

Now I need to save these sequences as a fasta file. 


```{r}
# Create FASTA-formatted lines: ">ID" on one line, followed by the sequence on the next
fasta_lines <- apply(neuropeptide_hormone_activity, 1, function(row) {
  paste0(">", row["Entry"], "\n", row["Sequence"])
})

# Write these lines to a .fasta file
writeLines(fasta_lines, con = "../../data/neuropeptide_hormone_activity_1.fasta")
```

Now we need to see the sequence length distribution of the neuropeptides in 
order to match it when sampling the non-neuropeptide sequences.

```{r}
hist_plot_group1 <- neuropeptide_hormone_activity %>%
  ggplot(aes(x = Length)) +
  geom_histogram(
    binwidth = 50,
    center = 25,
    color = "gray55",
    fill = "cornsilk") +
  scale_x_continuous(
    expand = expansion(mult = c(0,0.01))
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0,0.1))
  ) +
  geom_text(
    stat = "bin",
    binwidth = 50,
    center = 25,
    aes(label = ..count..),
    vjust = -0.5  # move text slightly above the bar
  ) +
  theme_cowplot()

hist_plot_group1

```
```{r}
binned_counts <- neuropeptide_hormone_activity %>%
  mutate(
    bin = cut(
      Length,
      breaks = seq(0, 600, by = 50),  # start at 0 to align with center=25
      include.lowest = TRUE,
      include.highest = TRUE,
      right = FALSE,  # [0,50), [50,100), etc.
      labels = paste(seq(0, 550, by = 50), seq(49, 599, by = 50), sep = "-")
    )
  ) %>%
  count(bin, name = "sequence_count")

print(binned_counts)

write_csv(binned_counts, "../../data/sequence_counts_per_bin_group_1.csv")
```


# Let's look at how many sequences have the "peptide" annotated.

```{r}
peptides_ann <- clean_data_5 %>%
  select(Entry, Sequence, Peptide) %>%
  distinct() %>%
  na.omit(Peptide)
```


Here is a list of all the peptide starts and ends. 
```{r}
peptides_ann_2 <- peptides_ann %>%
  # extract all “123..456” substrings as a list-column
  mutate(range_str = str_extract_all(Peptide, "\\d+\\.\\.\\d+")) %>%
  # turn each element of that list into its own row
  unnest_longer(range_str) %>%
  # split “123..456” into two new columns
  separate(range_str, into = c("peptide_start","peptide_end"), sep = "\\.\\.") %>%
  # convert to integer
  mutate(
     start = as.integer(peptide_start),
     end   = as.integer(peptide_end)) %>%
  select(Entry, Sequence, peptide_start, peptide_end)

```

How much data is left after I filter out the starts at 1. 

```{r}
not_1_start <- peptides_ann_2 %>%
  filter(peptide_start != 1)

```

get Signal sequence info:
```{r}
signal_info <- clean_data_5 %>%
  select(Entry, Signal.peptide) %>%
  distinct() %>%
  mutate(range_str = str_extract_all(Signal.peptide, "\\d+\\.\\.\\d+")) %>%
  unnest_longer(range_str) %>%
  separate(range_str, into = c("signal_start","signal_end"), sep = "\\.\\.") %>%
  mutate(
     start = as.integer(signal_start),
     end   = as.integer(signal_end)) %>%
  select(Entry, signal_start, signal_end)
```

Joining the peptide and signal info dataframes together.

```{r}
peptide_data_full <- peptides_ann_2 %>%
  left_join(signal_info)
```

```{r}
right_after_signal <- peptide_data_full %>%
  mutate(peptide_after_signal = ifelse(as.numeric(signal_end) + 1 == peptide_start, TRUE, FALSE))
```  
  
Let's see how many peptides are NOT right after the signal"
```{r}
right_after_signal %>%
  filter(peptide_after_signal == FALSE) %>%
  count()

```
```{r}


```





