---
title: "crf_analysis.Rmd"
output: html_document
date: "2025-07-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(tidytext)    # for reorder_within() and scale_x_reordered()
```


#=========================================================================================
# VALIDATION RESULTS
#=========================================================================================

Quick peak at the crf validation results for the 40% hom group:
```{r}
crf_40 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/crf_results_40_hom_dup.csv")
crf_50 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/crf_results_50_hom_dup.csv")
crf_60 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/crf_results_60_hom_dup.csv")
crf_70 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/crf_results_70_hom_dup.csv")
crf_80 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/crf_results_80_hom_dup.csv")
crf_90 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/crf_results_90_hom_dup.csv")

```

combining data into a single df
```{r}

combined <- bind_rows(
  `40` = crf_40,
  `50` = crf_50,
  `60` = crf_60,
  `70` = crf_70,
  `80` = crf_80,
  `90` = crf_90,
  .id  = "perc_homology"
) %>%
  mutate(
    # make sure perc_homology is numeric, not a string
    perc_homology = as.numeric(perc_homology)
  )
```

Now I want a line plot of validation accuracy across number of features. Each percent homology
will be a line of a different color. 
```{r}

ggplot(combined, aes(x = num_features, y = validation, color = factor(perc_homology))) +
  geom_line(size = 1) +
  scale_color_manual(
    name   = "training/validation\nset homology (%)",
    values = qualitative_hcl(
      n       = length(unique(combined$perc_homology)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Validation Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```
#=========================================================================================
# TEST RESULTS
#=========================================================================================

Test set done. Let's plot the test results.

```{r}
crf_40_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_crf_results/crf_test_results_40_hom.csv")
crf_50_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_crf_results/crf_test_results_50_hom.csv")
crf_60_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_crf_results/crf_test_results_60_hom.csv")
crf_70_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_crf_results/crf_test_results_70_hom.csv")
crf_80_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_crf_results/crf_test_results_80_hom.csv")
crf_90_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_crf_results/crf_test_results_90_hom.csv")

```

```{r}
combined_test <- bind_rows(
  `40` = crf_40_test,
  `50` = crf_50_test,
  `60` = crf_60_test,
  `70` = crf_70_test,
  `80` = crf_80_test,
  `90` = crf_90_test,
  .id  = "perc_homology"
) %>%
  mutate(
    # make sure perc_homology is numeric, not a string
    perc_homology = as.numeric(perc_homology)
  )
```

```{r}
ggplot(combined_test, aes(x = num_features, y = balanced_accuracy, color = factor(perc_homology))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "training/validation\nset homology (%)",
    values = qualitative_hcl(
      n       = length(unique(combined$perc_homology)),
      palette = "Set 2"
    )
  ) +
  scale_x_continuous(limits = c(0, 200)) +
  labs(
    x = "Number of Features",
    y = "Test Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```

Loading validation results
```{r}
percent_identity <- 40
base_dir <- file.path(
  "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results",
  "crf_csv_results",
  paste0("graphpart_", percent_identity)
)

# 1) Locate all per‐fold validation CSVs
files <- list.files(
  path       = base_dir,
  pattern    = "crf_results_excluded_fold_\\d+\\.csv$",
  recursive  = TRUE,
  full.names = TRUE
)

# 2) Read & combine
all_val_results <- files %>%
  set_names() %>% 
  map_dfr(~ {
    df <- read_csv(.x, show_col_types = FALSE)
    fold_num <- str_extract(basename(.x), "\\d+") %>% as.integer()
    df %>% mutate(fold = fold_num)
  })

```

plotting the validation results:
```{r}
ggplot(all_val_results, aes(x = num_features, y = validation, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Average Validation Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )

```



Loading all the test results from the 40% nested k-fold cross validation
```{r}
library(tidyverse)

# 1) Point to the directory where your fold sub-folders live:
base_dir <- "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/crf_csv_results/graphpart_40"

# 2) Find all the CSV files (one per fold):
files <- list.files(
  path       = base_dir, 
  pattern    = "^crf_test_results_fold_\\d+\\.csv$", 
  recursive  = TRUE,
  full.names = TRUE
)

# 3) Read & combine, extracting the fold number from each filename:
all_results <- files %>% 
  set_names() %>% 
  map_dfr(read_csv, .id = "file", show_col_types = FALSE) %>% 
  mutate(
    fold = str_extract(basename(file), "\\d+")
  ) %>% 
  select(fold, num_features, test_accuracy)
```

Ploting the graph:
```{r}
ggplot(all_results, aes(x = num_features, y = test_accuracy, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Test Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```

Getting df with the average test accuracy:
```{r}
avg_results <- all_results %>%
  # if test_accuracy is character, first turn it numeric:
  mutate(test_accuracy = as.numeric(test_accuracy)) %>%
  group_by(num_features) %>%
  summarise(
    avg_test_accuracy = mean(test_accuracy, na.rm = TRUE),
    .groups = "drop"
  )
```

plotting the average test result:
```{r}
ggplot(
  data = avg_results,
  aes(x = num_features, y = avg_test_accuracy)
  ) +
  geom_line(size = 0.8, color = "darkmagenta") +
  labs(
    x = "Number of Features",
    y = "Average Test Accuracy (%)"
  ) +
  theme_minimal()
```

Detecting the final number of features I need to use:
(Elbow/knee method)

```{r}
avg_results <- avg_results %>% arrange(num_features)
slopes   <- diff(avg_results$avg_test_accuracy) / diff(avg_results$num_features)
threshold <- 0.001  # e.g. 0.1% gain per 10 features
elbow_idx <- which(slopes < threshold)[1] + 0
opt_n     <- avg_results$num_features[elbow_idx]
opt_n

```
Let's just use 175 features. 


















