---
title: "deeploc_analysis"
output: html_document
date: "2025-04-09"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
```

#---------------------------------------------------------------------------------------------------------------------------------------
# IMPORTANT! Go to clean_KW-0527_data.Rmd and rerun code to get the "clean_data_5" dataset!!!!!!!!!!
#---------------------------------------------------------------------------------------------------------------------------------------

# let's look at the deeploc results and compare them to the Uniprot localization annotations.

```{r}
deeploc_data <- read.csv(file = "../../data/combined_deeploc_results.csv")

```


Now, let's do an inner join between clean_data_5 and the deeploc data:
```{r}
loc_df <- clean_data_5 %>%
  select(Entry, Localization) %>%
  distinct() %>%
  rename("uniprot_localization" = Localization)

loc_df %>%
  group_by(uniprot_localization) %>%
  count()
```

```{r}
deeploc_df <- deeploc_data %>%
  select(Protein_ID, Localizations) %>%
  rename("Entry" = Protein_ID,
         "deeploc_localization" = Localizations)

deeploc_df %>%
  group_by(deeploc_localization) %>%
  count()

```

Now joining the datasets together:
```{r}
joined_loc <- inner_join(deeploc_df, loc_df)
```

First, I want to separate multiple localization values in the deeploc column
```{r}
joined_loc <- joined_loc %>%
  separate_rows(deeploc_localization, sep = "\\|") %>%
  mutate(deeploc_localization = str_trim(deeploc_localization))  # optional: remove surrounding whitespace

joined_loc %>%
  group_by(deeploc_localization) %>%
  count()
```

Now I need to change some of the entries so that they match between the deeploc and uniprot data. 
```{r}
joined_loc_2 <- joined_loc %>%
  mutate(uniprot_localization = ifelse(uniprot_localization == "Secreted", "Secreted/Extracellular", uniprot_localization)) %>%
  mutate(deeploc_localization = ifelse(deeploc_localization == "Extracellular", "Secreted/Extracellular", deeploc_localization)) %>%
  mutate(uniprot_localization = ifelse(uniprot_localization == "Membrane", "Cell membrane", uniprot_localization)) %>%
  mutate(uniprot_localization = ifelse(uniprot_localization == "Rough endoplasmic reticulum", "Endoplasmic reticulum", uniprot_localization))

joined_loc_2 %>%
  filter(uniprot_localization != deeploc_localization) %>%
  distinct(uniprot_localization, deeploc_localization)
```


Now let's plot a heatmap of the predicted and annotated localizations:
```{r}
for_heat_plot <- joined_loc_2 %>%
  group_by(deeploc_localization, uniprot_localization) %>%
  summarize(count = n(), .groups = "drop")

for_heat_plot_2 <- for_heat_plot %>%
  group_by(uniprot_localization) %>%
  mutate(freq = count/sum(count))

for_heat_plot_2 <- for_heat_plot_2 %>%
  mutate(
    deeploc_localization = factor(deeploc_localization),
    uniprot_localization = factor(uniprot_localization))

for_heat_plot_2 <- for_heat_plot_2 %>%
  mutate(
    deeploc_localization = fct_relevel(deeploc_localization, "Secreted/Extracellular", "Cell membrane", "Endoplasmic reticulum", "Nucleus"),
    uniprot_localization = fct_relevel(uniprot_localization, "Secreted/Extracellular",  "Cell membrane", "Endoplasmic reticulum", "Nucleus")) 
```


```{r}
# Get the levels in order (so we can reference the lower-left corner)
x_levels <- levels(for_heat_plot_2$deeploc_localization)
y_levels <- levels(for_heat_plot_2$uniprot_localization)

# Pick the 4 lowest levels (leftmost and bottommost)
x_rect <- x_levels[1:4]
y_rect <- y_levels[1:4]


heatmap_loc_plot <- for_heat_plot_2 %>%
  ggplot(aes(x = deeploc_localization, y = uniprot_localization, fill = freq)) +
  geom_tile(color = "white") +  
  geom_text(aes(label = count), color = "white", size = 3) +
  scale_x_discrete(
    name = "DeepLoc localization prediction",
    expand = expansion(add = c(0,0))
  ) +
  scale_y_discrete(
    name = "Uniprot localization annotation",
    expand = expansion(add = c(0,0))
  ) +
  scale_fill_gradient(
    low = "#b2dfdb",   # for low counts
    high = "#014d4e",   # for high counts (darkest)
    name = "Frequency"
  ) +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  annotate("rect",
           xmin = 0.5, xmax = 4.5,   # first 4 x-levels
           ymin = 0.5, ymax = 4.5,   # first 4 y-levels
           fill = NA,                # no fill
           color = "maroon",
           linewidth = 0.6) 

heatmap_loc_plot

ggsave(filename = "../../figures/local_heatmap_plot.png", plot = heatmap_loc_plot, width = 8, height = 6)
```

Now let's look at the new deeplock results for the non-neuropeptides.

```{r}
deeploc_data_non_neuro <- read.csv(file = "../../data/combined_deeploc_results_non_neuro.csv")

```


```{r}
deeploc_non_neuro <- deeploc_data_non_neuro %>%
  select(Protein_ID, Localizations) %>%
  rename("Entry" = Protein_ID,
         "deeploc_localization" = Localizations)

deeploc_non_neuro %>%
  group_by(deeploc_localization) %>%
  count()

```

I want to again separate multiple localization values in the deeploc column
```{r}
deeploc_non_neuro_sep <- deeploc_non_neuro %>%
  separate_rows(deeploc_localization, sep = "\\|") %>%
  mutate(deeploc_localization = str_trim(deeploc_localization))  # optional: remove surrounding whitespace

deeploc_non_neuro_sep %>%
  group_by(deeploc_localization) %>%
  count()
```























