---
title: "Figure_6"
output: html_document
date: "2026-02-12"
---
```{r}
library(tidyverse)
library(cowplot)
library(ggridges)
```

LogReg l18 320 feature model prediction results:
```{r}
mouse_pred <- read.csv("../../data/fig_6_mouse_secretome/layer_18_secretome_mouse_log_reg_320_predictions.csv")

worm_pred <- read.csv("../../data/fig_6_c_elegans_secretome/layer_18_secretome_c_elegans_log_reg_320_predictions.csv")

fly_pred <- read.csv("../../data/fig_6_drosophila_secretome/layer_18_secretome_drosophila_log_reg_320_predictions.csv")

zebra_fish_pred <- read.csv("../../data/fig_6_zebrafish_secretome/layer_18_secretome_zebrafish_log_reg_320_predictions_fixed.csv")
```

Neuropeptide flag files:
```{r}
mouse_flags <- read.csv("../../data/fig_6_mouse_secretome/secretome_neuropeptide_flags.csv")

worm_flags <- read.csv("../../data/fig_6_c_elegans_secretome/secretome_neuropeptide_flags.csv")

fly_flags <- read.csv("../../data/fig_6_drosophila_secretome/secretome_neuropeptide_flags.csv")

fish_flags <- read.csv("../../data/fig_6_zebrafish_secretome/secretome_neuropeptide_flags.csv")
```

Loading in the annotations:

```{r}
library(dplyr)

mouse_ann <- read.csv("../../data/fig_6_mouse_secretome/secretome_annotations_mouse_clean.csv")

worm_ann  <- read.csv("../../data/fig_6_c_elegans_secretome/secretome_annotations_c_elegans_clean.csv")

fly_ann   <- read.csv("../../data/fig_6_drosophila_secretome/secretome_annotations_drosophila_clean.csv")

fish_ann   <- read.csv("../../data/fig_6_zebrafish_secretome/secretome_annotations_zebrafish_clean.csv")
```


Defining a function for cleaning my annotations
```{r}
clean_secretome_annotations <- function(secretome_annotations) {
  
  secretome_annotations %>%
    dplyr::select(
      Entry,
      Sequence,
      Length,
      Signal.peptide,
      Peptide,
      Protein.names,
      Function..CC.,
      Gene.Ontology..molecular.function.
    ) %>%
    dplyr::rename(
      Signal_peptide   = Signal.peptide,
      Protein_names    = Protein.names,
      Protein_function = Function..CC.,
      Go_function      = Gene.Ontology..molecular.function.
    ) %>%
    dplyr::select(
      Entry,
      Protein_names,
      Protein_function,
      Go_function,
      Signal_peptide,
      Peptide,
      Sequence,
      Length
    )
}

```

Applying my function to all three secretomes:
```{r}
mouse_clean <- clean_secretome_annotations(mouse_ann)
worm_clean  <- clean_secretome_annotations(worm_ann)
fly_clean   <- clean_secretome_annotations(fly_ann)
fish_clean   <- clean_secretome_annotations(fish_ann)

```

FINAL MOUSE DF:
```{r}
mouse_final <- mouse_clean %>%
  inner_join(mouse_pred, by = "Entry") %>%   # keep only predicted peptides
  left_join(mouse_flags, by = "Entry") %>%   # add neuropeptide column 
  select(Entry, probability, neuropeptide, everything())

```

FINAL C.ELEGANS DF:
```{r}
worm_final <- worm_clean %>%
  inner_join(worm_pred, by = "Entry") %>%
  left_join(worm_flags, by = "Entry") %>%
  select(Entry, probability, neuropeptide, everything())

```

FINAL DROSOPHILA DF:
```{r}
fly_final <- fly_clean %>%
  inner_join(fly_pred, by = "Entry") %>%
  left_join(fly_flags, by = "Entry") %>%
  select(Entry, probability, neuropeptide, everything())

```

FINAL ZEBRA_FISH DF:
```{r}
fish_final <- fish_clean %>%
  inner_join(zebra_fish_pred, by = "Entry") %>%
  left_join(fish_flags, by = "Entry") %>%
  select(Entry, probability, neuropeptide, everything())

```

Helper function for counting bins and plotting:
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)
library(cowplot)

make_binned_plot <- function(df, prob_col = "highest_prob") {
  
  df_prepared <- df %>%
    mutate(
      neuropeptide = ifelse(neuropeptide == TRUE,
                            "Neuropeptide",
                            "Non-neuropeptide"),
      prob_bin = cut(
        .data[[prob_col]],
        breaks = seq(0, 1, by = 0.1),
        include.lowest = TRUE
      )
    )
  
  binned_long <- df_prepared %>%
    count(prob_bin, neuropeptide)
  
  label_df <- binned_long %>%
    group_by(prob_bin) %>%
    mutate(prop = n / sum(n)) %>%
    filter(neuropeptide == "Neuropeptide") %>%
    summarise(
      prop_np = sum(prop),
      .groups = "drop"
    )
  
  plot <- binned_long %>%
    mutate(
      neuropeptide = factor(
        neuropeptide,
        levels = c("Neuropeptide", "Non-neuropeptide")
      )
    ) %>%
    ggplot(aes(x = fct_rev(prob_bin), y = n,
               fill = fct_relevel(neuropeptide,
                                  "Neuropeptide",
                                  "Non-neuropeptide"))) +
    geom_col(
      position = position_fill(reverse = TRUE),
      color = "grey40",
      size = 0.3,
      alpha = 0.85
    ) +
    scale_y_continuous(
      breaks = seq(0, 1, by = 0.2),
      labels = percent_format(accuracy = 1),
      expand = expansion(mult = c(0, 0))
    ) +
    scale_x_discrete(
      expand = expansion(mult = c(0, 0))
    ) +
    geom_text(
      data = label_df,
      aes(
        x = prob_bin,
        y = prop_np,
        label = percent(prop_np, accuracy = 1)
      ),
      color = "white",
      size = 4,
      vjust = -0.5,
      inherit.aes = FALSE
    ) +
    scale_fill_manual(
      values = c(
        "Non-neuropeptide" = "#3d81c4",
        "Neuropeptide" = "#ffb145"
      )
    ) +
    labs(
      x = "Predicted probability bin",
      y = "Proportion of proteins",
      fill = NULL
    ) +
    theme_cowplot() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "top",
      legend.direction = "horizontal"
    )
  
  return(plot)
}

```


Generating plots for each species:
```{r}
plot_mouse <- make_binned_plot(mouse_final, prob_col = "probability")
plot_worm  <- make_binned_plot(worm_final, prob_col = "probability")
plot_fly   <- make_binned_plot(fly_final, prob_col = "probability")
plot_fish   <- make_binned_plot(fish_final, prob_col = "probability")

plot_mouse
plot_worm
plot_fly
plot_fish

```
Combining three dataset:
```{r}
library(dplyr)
library(ggplot2)
library(forcats)

mouse_final$species <- "M.\u00A0musculus"
worm_final$species  <- "C.\u00A0elegans"
fly_final$species   <- "D.\u00A0melanogaster"
fish_final$species   <- "D.\u00A0rerio"

combined_df <- bind_rows(
  mouse_final,
  worm_final,
  fly_final,
  fish_final
) %>%
  mutate(
    neuropeptide = ifelse(neuropeptide == TRUE,
                          "Neuropeptide",
                          "Non-neuropeptide"),
    neuropeptide = factor(
      neuropeptide,
      levels = c("Non-neuropeptide", "Neuropeptide")),
    species = factor(species,
                     levels = c("D.\u00A0rerio", "D.\u00A0melanogaster", "C.\u00A0elegans", "M.\u00A0musculus"))

  )

```


```{r}
library(ggplot2)
library(ggridges)
library(scales)
library(cowplot)

plot_density <- ggplot(
  combined_df,
  aes(
    x = probability,
    y = species,
    fill = neuropeptide
  )
) +
  geom_density_ridges(
    alpha = 0.7,
    scale = 0.9,
    rel_min_height = 0.01,
    color = "grey40",
    size = 0.3
  ) +
  geom_vline(
  xintercept = 0.7,
  color = "#7a0019",   # dark maroon
  linewidth = 0.75,
  linetype = "dashed"
 ) +
  scale_x_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.1),
    labels = number_format(accuracy = 0.1),
    expand = expansion(mult = c(0,0.01))
  ) +
  scale_y_discrete( 
    expand = expansion(mult = c(0,0))
  ) +
  scale_fill_manual(
    values = c(
      "Non-neuropeptide" = "#64a1de",
      "Neuropeptide" = "#fca226"
    )
  ) +
  labs(
    x = "Predicted probability",
    y = NULL,
    fill = NULL
  ) +
  theme_cowplot() +
  theme(
    legend.position = "top",
    axis.text.y = element_text(
      angle = 90,
      hjust = 0,     # anchor at tick
      vjust = 0.6,   # center vertically
      size = 14,
      face = "italic"
    ),
    panel.grid.major.x = element_line(color = "grey85"),
    panel.grid.major.y = element_line(color = "grey85"),
    panel.grid.minor = element_blank()
  )

plot_density

```

Worst neuropeptide predictions for mouse secretome
```{r}
mouse_final %>%
  filter(neuropeptide == TRUE) %>%
  arrange(probability) %>%
  transmute(
    Entry = Entry,
    probability = probability,
    neuropeptide = neuropeptide,
    Protein_names = Protein_names
  ) %>%
  slice_head(n = 15)
```

Worst neuropeptide predictions for zebrafish secretome
```{r}
fish_final %>%
  filter(neuropeptide == FALSE) %>%
  arrange(desc(probability)) %>%
  transmute(
    Entry = Entry,
    probability = probability,
    neuropeptide = neuropeptide,
    Protein_names = Protein_names
  ) %>%
  slice_head(n = 15)
```

# Getting 8 files for the outliers for each species:
```{r}
worm_final_filtered <- worm_final %>%
  filter(neuropeptide == FALSE, probability >= 0.7) %>%
  arrange(desc(probability)) %>%
  select(Entry, probability, neuropeptide, Protein_names, everything())
```

```{r}
write.csv(
  worm_final_filtered,
  file = "../../data/outliers/worm_prob>=0.7_FALSE.csv",
  row.names = FALSE
)
```




Now, let's make confusion matrices for each plot.
```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(cowplot)

make_confusion_plot <- function(df, species_name, threshold = 0.7) {
  
  df_eval <- df %>%
    mutate(
      actual = ifelse(neuropeptide == TRUE,
                      "Neuro.",
                      "Non-neuro."),
      predicted = ifelse(probability >= threshold,
                         "Neuro.",
                         "Non-neuro.")
    )
  
  conf_mat <- df_eval %>%
    count(actual, predicted)
  
  total_n <- nrow(df_eval)
  
  ggplot(conf_mat, aes(x = predicted, y = actual, fill = n)) +
    geom_tile(color = "white") +
    geom_text(
      aes(label = n),
      size = 5,
      fontface = "bold",
      color = ifelse(conf_mat$n > max(conf_mat$n) * 0.6, "white", "black")
    ) +
    scale_fill_gradient(low = "white", high = "#752132") +
    labs(
      subtitle = paste0("N = ", total_n),
      x = "Predicted label",
      y = "True label"
    ) +
    scale_y_discrete(expand = expansion(mult = c(0,0))) +
    scale_x_discrete(expand = expansion(mult = c(0,0))) +
    theme_cowplot(14) +
    theme(
      plot.title = element_text(hjust = 0.5),
      axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
      legend.position = "none",
      plot.subtitle = element_text(
        hjust = 0.5,        # center horizontally
        size = 17,          # control font size
        face = "bold"       # optional
      )
    )
}



```

Generating the plots for each of the 3 secretomes
```{r}
plot_mouse_cm <- make_confusion_plot(mouse_final, "M. musculus")
plot_worm_cm  <- make_confusion_plot(worm_final, "C. elegans")
plot_fly_cm   <- make_confusion_plot(fly_final, "D. melanogaster")
plot_fish_cm   <- make_confusion_plot(fish_final, "D. rerio")

plot_mouse_cm
plot_worm_cm
plot_fly_cm
plot_fish_cm

```

Stacking confusion plots:
```{r}
confusion_stack <- plot_grid(
  plot_mouse_cm,
  plot_worm_cm,
  plot_fly_cm,
  plot_fish_cm,
  labels = c("b.", "c.", "d.", "e."),
  ncol = 1,
  align = "v"
)
```

combining with density plot:
```{r}
compound_plot <- plot_grid(
  plot_density,
  confusion_stack,
  labels = c("a.", ""),
  ncol = 2,
  rel_widths = c(2, 1)
)

compound_plot

```

Saving the plot:
```{r}
ggsave(
  filename = "../../figures/paper_figures/figure_6_other_secretomes.png",
  plot = compound_plot,
  width = 9,
  height = 9.4,
  dpi = 400
)
```


Now, we need to calculate the statistics:
```{r}
compute_metrics <- function(df, threshold = 0.7) {
  
  df_eval <- df %>%
    mutate(
      actual = ifelse(neuropeptide == TRUE, 1, 0),
      predicted = ifelse(probability >= threshold, 1, 0)
    )
  
  TP <- sum(df_eval$actual == 1 & df_eval$predicted == 1)
  TN <- sum(df_eval$actual == 0 & df_eval$predicted == 0)
  FP <- sum(df_eval$actual == 0 & df_eval$predicted == 1)
  FN <- sum(df_eval$actual == 1 & df_eval$predicted == 0)
  
  accuracy  <- (TP + TN) / (TP + TN + FP + FN)
  precision <- ifelse((TP + FP) == 0, NA, TP / (TP + FP))
  recall    <- ifelse((TP + FN) == 0, NA, TP / (TP + FN))
  f1        <- ifelse(is.na(precision) | is.na(recall) | (precision + recall) == 0,
                      NA,
                      2 * precision * recall / (precision + recall))
  
  tibble(
    TP = TP,
    TN = TN,
    FP = FP,
    FN = FN,
    Accuracy = accuracy,
    Precision = precision,
    Recall = recall,
    F1 = f1
  )
}

```

```{r}
mouse_metrics <- compute_metrics(mouse_final, threshold = 0.7)
worm_metrics  <- compute_metrics(worm_final, threshold = 0.7)
fly_metrics   <- compute_metrics(fly_final, threshold = 0.7)
fish_metrics   <- compute_metrics(fish_final, threshold = 0.7)
```


Putting it all into a table:
```{r}
metrics_table <- bind_rows(
  mouse_metrics %>% mutate(Species = "M. musculus"),
  worm_metrics  %>% mutate(Species = "C. elegans"),
  fly_metrics   %>% mutate(Species = "D. melanogaster"),
  fish_metrics   %>% mutate(Species = "D. rerio")
) %>%
  select(Species, Accuracy, Precision, Recall, F1, TP, FP, FN, TN)

metrics_table

```



