---
title: "Figure_1_PCA"
output: html_document
date: "2026-01-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)

# 1) Load averaged embeddings
emb <- read.csv("../../data/esm2_secretome_embeddings_averaged.csv",
                check.names = FALSE)

# 2) Load neuropeptide flags (these are the revised ones from Thomas and Helena)
secr_neuro_flags <- read.csv("../../data/secretome_neuroflags_revised_final.csv")

# Inspect columns to find the flag column name
names(secr_neuro_flags)

```

```{r}
flag_col <- "neuropeptide"   # <- change if your flag column has a different name

# 3) Join embeddings with neuropeptide labels
df <- emb %>%
  inner_join(secr_neuro_flags, by = "Entry")

# All embedding columns (those starting with "e_")
emb_cols <- grep("^e_", names(df), value = TRUE)
length(emb_cols)  # should be 1280
nrow(df)
```


```{r}
# 4) PCA on the embedding features
pca <- prcomp(df[, emb_cols], scale. = TRUE)

# 5) Build a data frame with PC1/PC2 and labels
scores_esm_average <- as_tibble(pca$x[, 1:2]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]]
  )

head(scores_esm_average)

```

```{r}
library(dplyr)
library(ggplot2)

scores_esm_average <- scores_esm_average %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

esm_average_plot <- ggplot() +
  # Invisible layer just to force legend creation
  geom_point(
    data = scores_esm_average,
    aes(x = PC1, y = PC2, fill = type),
    alpha = 0,
    shape = 21,
    show.legend = TRUE
  ) +
  geom_point(
    data = filter(scores_esm_average, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#062c5e",
    shape = 21,
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  geom_point(
    data = filter(scores_esm_average, type == "Neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#ff5e00",
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5
  ) +
  scale_fill_manual(
    name   = "",
    values = c(
      "Non-neuropeptide" = "#477aad",
      "Neuropeptide"     = "#ffb145"
    )
  ) +
  guides(fill = guide_legend()) +   # ← THIS FIXES IT
  theme_cowplot() +
  labs(
    x = "PC1",
    y = "PC2"
  ) +
  theme(legend.position = "top")


esm_average_plot


```
Testing legend:
```{r}
esm_average_plot$guides

```


#==================================================================================================
# Now we are plotting the ESM2 embeddings (maxpooled embeddings)
#==================================================================================================

```{r}
library(tidyverse)

# 1) Load averaged embeddings
emb <- read.csv("../../data/esm2_secretome_embeddings_maxpooled.csv",
                check.names = FALSE)

# 2) Load neuropeptide flags (these are the revised ones from Thomas and Helena)
secr_neuro_flags <- read.csv("../../data/secretome_neuroflags_revised_final.csv")

# Inspect columns to find the flag column name
names(secr_neuro_flags)

```

```{r}
flag_col <- "neuropeptide"   # <- change if your flag column has a different name

# 3) Join embeddings with neuropeptide labels
df <- emb %>%
  inner_join(secr_neuro_flags, by = "Entry")

# All embedding columns (those starting with "e_")
emb_cols <- grep("^e_", names(df), value = TRUE)
length(emb_cols)  # should be 1280
nrow(df)

```


```{r}
# 4) PCA on the embedding features
pca <- prcomp(df[, emb_cols], scale. = TRUE)

# 5) Build a data frame with PC1/PC2 and labels
scores_esm_maxpool <- as_tibble(pca$x[, 1:2]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]]
  )

head(scores_esm_maxpool)

```

```{r}
library(dplyr)
library(ggplot2)

scores_esm_maxpool <- scores_esm_maxpool %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

esm_maxpool_plot <- ggplot() +
  # Non-neuropeptides: blue fill, black outline, under
  geom_point(
    data = filter(scores_esm_maxpool, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#062c5e",          # outline color
    shape = 21,               # filled circle with border
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  # Neuropeptides: orange fill, black outline, on top
  geom_point(
    data = filter(scores_esm_maxpool, type == "Neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#ff5e00",          # outline color
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5              # outline thickness
  ) +
  scale_fill_manual(
    name   = "",
    values = c("Non-neuropeptide" = "#477aad",
               "Neuropeptide"     = "#ffb145")
  ) +
  theme_cowplot() +
  labs(
    x = "PC1",
    y = "PC2"
  ) 

```


#==================================================================================================
# Now we are plotting the SAE embeddings (averaged embeddings)
#==================================================================================================

```{r setup-pca, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tibble)

# 1) Load averaged SAE embeddings
sae_emb <- read.csv(
  "../../data/sae_layer18_secretome_embeddings_averaged_1201.csv",
  check.names = FALSE
)

dim(sae_emb)  # sanity check: ~1201 x 10241

# 2) Load neuropeptide flags (these are the revised ones from Thomas and Helena)
secr_neuro_flags <- read.csv("../../data/secretome_neuroflags_revised_final.csv")
names(secr_neuro_flags)  # check column names

```


```{r join-flags}
flag_col <- "neuropeptide"  # adjust if needed

# 3) Join embeddings with neuropeptide labels
df <- sae_emb %>%
  inner_join(secr_neuro_flags, by = "Entry")

# 4) Embedding columns: all starting with "e_"
emb_cols <- grep("^e_", names(df), value = TRUE)
length(emb_cols)  # should be 10240
nrow(df)

```


```{r pca-sae}
# 5) PCA on SAE embeddings

# Compute SD for each embedding column
emb_matrix <- as.matrix(df[, emb_cols])
sds <- apply(emb_matrix, 2, sd)

# Keep only non-constant columns
non_const_cols <- emb_cols[sds > 0]
length(non_const_cols)  # should be slightly less than 10240

# Optionally inspect how many were dropped
cat("Dropped", sum(sds == 0), "constant columns\n")

# Run PCA on filtered matrix
pca_sae <- prcomp(df[, non_const_cols], scale. = TRUE)

# Variance explained
var_explained <- (pca_sae$sdev^2) / sum(pca_sae$sdev^2)
pc1_var <- round(var_explained[1] * 100, 1)
pc2_var <- round(var_explained[2] * 100, 1)

# 6) Build scores data frame with PC1/PC2 and labels
scores_sae_average <- as_tibble(pca_sae$x[, 1:2]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]]
  )

colnames(scores_sae_average)[1:2] <- c("PC1", "PC2")

head(scores_sae_average)

```

```{r}
library(dplyr)
library(ggplot2)

scores_sae_average <- scores_sae_average %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

sae_average_plot <- ggplot() +
  # Non-neuropeptides: blue fill, dark blue outline, under
  geom_point(
    data = dplyr::filter(scores_sae_average, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#062c5e",          # outline color
    shape = 21,                 # filled circle with border
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  # Neuropeptides: orange fill, orange outline, on top
  geom_point(
    data = dplyr::filter(scores_sae_average, type == "Neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#ff5e00",          # outline color
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5               # outline thickness
  ) +
  scale_fill_manual(
    name   = "",
    values = c("Non-neuropeptide" = "#477aad",
               "Neuropeptide"     = "#ffb145")
  ) +
  theme_cowplot() +
  labs(
    x = "PC1",
    y = "PC2"
  ) 
sae_average_plot
```
#===================================================================================
# Now let's plot SAE embeddings (maxpooled). instead of averaged
#===================================================================================

```{r setup-pca, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tibble)

# 1) Load averaged SAE embeddings
sae_emb <- read.csv(
  "../../data/sae_layer18_secretome_embeddings_maxpooled_1201.csv",
  check.names = FALSE
)

dim(sae_emb)  # sanity check: ~1201 x 10241

# 2) Load neuropeptide flags (these are the revised ones from Thomas and Helena)
secr_neuro_flags <- read.csv("../../data/secretome_neuroflags_revised_final.csv")
names(secr_neuro_flags)  # check column names

```


```{r join-flags}
flag_col <- "neuropeptide"  # adjust if needed

# 3) Join embeddings with neuropeptide labels
df <- sae_emb %>%
  inner_join(secr_neuro_flags, by = "Entry")

# 4) Embedding columns: all starting with "e_"
emb_cols <- grep("^e_", names(df), value = TRUE)
length(emb_cols)  # should be 10240
nrow(df)
```


```{r pca-sae}
# 5) PCA on SAE embeddings

# Compute SD for each embedding column
emb_matrix <- as.matrix(df[, emb_cols])
sds <- apply(emb_matrix, 2, sd)

# Keep only non-constant columns
non_const_cols <- emb_cols[sds > 0]
length(non_const_cols)  # should be slightly less than 10240

# Optionally inspect how many were dropped
cat("Dropped", sum(sds == 0), "constant columns\n")

# Run PCA on filtered matrix
pca_sae <- prcomp(df[, non_const_cols], scale. = TRUE)

# Variance explained
var_explained <- (pca_sae$sdev^2) / sum(pca_sae$sdev^2)
pc1_var <- round(var_explained[1] * 100, 1)
pc2_var <- round(var_explained[2] * 100, 1)

# 6) Build scores data frame with PC1/PC2 and labels
scores_sae_maxpool <- as_tibble(pca_sae$x[, 1:2]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]]
  )

colnames(scores_sae_maxpool)[1:2] <- c("PC1", "PC2")

head(scores)

```

```{r}
library(dplyr)
library(ggplot2)

scores_sae_maxpool <- scores_sae_maxpool %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

sae_maxpool_plot <- ggplot() +
  # Non-neuropeptides: blue fill, dark blue outline, under
  geom_point(
    data = dplyr::filter(scores_sae_maxpool, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#062c5e",          # outline color
    shape = 21,                 # filled circle with border
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  # Neuropeptides: orange fill, orange outline, on top
  geom_point(
    data = dplyr::filter(scores_sae_maxpool, type == "Neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#ff5e00",          # outline color
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5               # outline thickness
  ) +
  scale_fill_manual(
    name   = "",
    values = c("Non-neuropeptide" = "#477aad",
               "Neuropeptide"     = "#ffb145")
  ) +
  theme_cowplot() +
  labs(
    x = "PC1",
    y = "PC2"
  ) 
sae_maxpool_plot
```
#=======================================================================================
# Let's now find the distances between the centroids
#=======================================================================================
```{r}
library(dplyr)

centroid_distance <- function(scores_df) {
  # Compute centroids
  centroids <- scores_df %>%
    group_by(type) %>%
    summarise(
      PC1 = mean(PC1),
      PC2 = mean(PC2),
      .groups = "drop"
    )

  # Extract centroids
  c_np  <- centroids %>% filter(type == "Neuropeptide")
  c_non <- centroids %>% filter(type == "Non-neuropeptide")

  # Euclidean distance between centroids
  dist <- sqrt(
    (c_np$PC1 - c_non$PC1)^2 +
    (c_np$PC2 - c_non$PC2)^2
  )

  list(
    centroids = centroids,
    distance  = dist
  )
}
```

Applying to SAE maxpool PCA
```{r}
esm_centroid <- centroid_distance(scores_sae_maxpool)

esm_centroid$centroids
esm_centroid$distance
```
```{r}
esm_centroid <- centroid_distance(scores_esm_average)

esm_centroid$centroids
esm_centroid$distance
```
```{r}
esm_centroid <- centroid_distance(scores_sae_average)

esm_centroid$centroids
esm_centroid$distance
```

```{r}
esm_centroid <- centroid_distance(scores_esm_maxpool)

esm_centroid$centroids
esm_centroid$distance
```
Helper Function for the silhouette scores:
```{r}
library(cluster)
library(dplyr)

silhouette_pc12 <- function(scores_df) {
  # Ensure correct ordering
  scores_df <- scores_df %>%
    filter(!is.na(PC1), !is.na(PC2), !is.na(type))

  # Encode class labels as integers
  labels <- as.integer(factor(scores_df$type))

  # Distance matrix in PC1–PC2 space
  dmat <- dist(scores_df[, c("PC1", "PC2")])

  # Compute silhouette
  sil <- silhouette(labels, dmat)

  # Return mean silhouette score
  mean(sil[, "sil_width"])
}
```

Calculating silhouette scores:
```{r}
sil_results <- tibble(
  Model   = c("ESM-2", "ESM-2", "SAE", "SAE"),
  Pooling = c("Mean", "Max", "Mean", "Max"),
  Silhouette_PC1_PC2 = c(
    silhouette_pc12(scores_esm_average),
    silhouette_pc12(scores_esm_maxpool),
    silhouette_pc12(scores_sae_average),
    silhouette_pc12(scores_sae_maxpool)
  )
)

sil_results
```




#===================================================
# Final compound PCA plot:
#===================================================


```{r}
legend_source <- ggplot(
  scores_esm_average,
  aes(x = PC1, y = PC2, fill = type)
) +
  geom_point(shape = 21, size = 3) +
  scale_fill_manual(
    name = "",
    values = c(
      "Non-neuropeptide" = "#477aad",
      "Neuropeptide"     = "#ffb145"
    )
  ) +
  theme_void() +
  theme(
    legend.position = "right"
  ) +
  guides(fill = guide_legend(nrow = 1))   # ← forces horizontal layout

legend <- cowplot::get_legend(legend_source)
legend



```

```{r}
library(cowplot)

panel_grid <- plot_grid(
  esm_average_plot + theme(legend.position = "none"),
  esm_maxpool_plot + theme(legend.position = "none"),
  sae_average_plot + theme(legend.position = "none"),
  sae_maxpool_plot + theme(legend.position = "none"),
  ncol = 2,
  labels = c("a.", "b.", "c.", "d.")
)

final_figure <- plot_grid(
  legend,
  panel_grid,
  ncol = 1,
  rel_heights = c(0.2, 1)
)

final_figure
```


```{r}
library(ggplot2)
ggsave(
  filename = "../../figures/paper_figures/Figure_1_final_PCA.png",
  plot = final_figure,
  width = 7,
  height = 7,
  dpi = 400
)
```




#------------------end of figure 1----------------------------------



Now, I am making the PCA of the 320 top features for the maxpooled SAE embeddings: (FINAL VERSION FOR PAPER)
```{r setup-pca, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tibble)
library(readr)

# 1) Load max-pooled SAE embeddings
sae_emb <- read.csv(
  "../../data/sae_layer18_secretome_embeddings_maxpooled_1201.csv",
  check.names = FALSE
)

dim(sae_emb)  # e.g., ~1201 x 10241 (Entry + 10240 features)

# 2) Load neuropeptide flags (revised)
secr_neuro_flags <- read.csv("../../data/secretome_neuroflags_revised_final.csv")
names(secr_neuro_flags)

# Column in secr_neuro_flags that indicates neuropeptide
flag_col <- "neuropeptide"  # adjust if needed
```

```{r}
# 3) Join embeddings with neuropeptide labels
df <- sae_emb %>%
  inner_join(secr_neuro_flags, by = "Entry")

# 4) Embedding columns: all starting with "e_"
emb_cols <- grep("^e_", names(df), value = TRUE)

length(emb_cols)  # should be 10240
nrow(df)          # e.g., 562

# Make sure the flag column is logical
# (works if it's TRUE/FALSE, 0/1, "0"/"1", "TRUE"/"FALSE")
df[[flag_col]] <- as.logical(as.integer(as.character(df[[flag_col]])) %||%
                              ifelse(as.character(df[[flag_col]]) %in% c("TRUE","True","true"), 1, 0))
```

```{r}
rank_df <- read_csv(
  "/Volumes/T7 Shield/layer_18_F1_scores/winning_thresholds_l18.csv",
  show_col_types = FALSE
)

names(rank_df)

# Rank by winning_F1_score (highest first), take top 320
top_features <- rank_df %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 320)

# 'feature' is a numeric index (1..10240)
top_indices <- as.integer(top_features$feature)

length(top_indices)
head(top_indices)
```

```{r}
# Sanity check: indices are within range
if (max(top_indices) > length(emb_cols) || min(top_indices) < 1) {
  stop("Some feature indices are out of range of embedding columns.")
}

# Map indices to column names
top_feats_in_df <- emb_cols[top_indices]

length(top_feats_in_df)  # should be 320
head(top_feats_in_df)

```

```{r}
# Matrix of just the selected top features
emb_matrix_top <- as.matrix(df[, top_feats_in_df, drop = FALSE])

# Drop zero-variance columns
sds_top <- apply(emb_matrix_top, 2, sd, na.rm = TRUE)
non_const_cols_top <- colnames(emb_matrix_top)[sds_top > 0]

cat("Dropped", sum(sds_top == 0), "constant columns among top features\n")

if (length(non_const_cols_top) < 2) {
  stop("Fewer than 2 non-constant columns remain; PCA cannot be computed.")
}

# PCA on selected, non-constant features
pca_sae_top <- prcomp(emb_matrix_top[, non_const_cols_top, drop = FALSE], scale. = TRUE)

# Variance explained
var_explained_top <- (pca_sae_top$sdev^2) / sum(pca_sae_top$sdev^2)
pc1_var_top <- round(var_explained_top[1] * 100, 1)
pc2_var_top <- round(var_explained_top[2] * 100, 1)

# Build scores with PC1 & PC2 and labels
scores_top <- as_tibble(pca_sae_top$x[, 1:2, drop = FALSE]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]]
  )

colnames(scores_top)[1:2] <- c("PC1", "PC2")

head(scores_top)
```

```{r}
scores_top <- scores_top %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

ggplot() +
  # Non-neuropeptides: blue fill, dark blue outline, under
  geom_point(
    data = dplyr::filter(scores_top, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#062c5e",
    shape = 21,
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  # Neuropeptides: orange fill, orange outline, on top
  geom_point(
    data = dplyr::filter(scores_top, type == "Neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#ff5e00",
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5
  ) +
  scale_fill_manual(
    name   = "",
    values = c("Non-neuropeptide" = "#477aad",
               "Neuropeptide"     = "#ffb145")
  ) +
  theme_cowplot() +
  labs(
    x = paste0("PC1 (", pc1_var_top, "%)"),
    y = paste0("PC2 (", pc2_var_top, "%)"),
    title = "SAE embeddings (top 320 features)"
  )
```

Looking at the centroid and silhouette scores:
```{r metrics-sae-top320, message=FALSE, warning=FALSE}
library(cluster)
library(dplyr)

# Ensure labels are clean
scores_top <- scores_top %>%
  mutate(
    is_neuropeptide = as.logical(is_neuropeptide),
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

# --- Centroid distance in PC1–PC2 ---
centroids_top <- scores_top %>%
  group_by(type) %>%
  summarise(
    c1 = mean(PC1, na.rm = TRUE),
    c2 = mean(PC2, na.rm = TRUE),
    n  = dplyr::n(),
    .groups = "drop"
  )

# sanity check: should have exactly 2 groups
if (nrow(centroids_top) != 2) {
  stop("Expected exactly two groups for centroid calculation. Check neuropeptide labels.")
}

centroid_distance_top <- sqrt(
  (centroids_top$c1[1] - centroids_top$c1[2])^2 +
  (centroids_top$c2[1] - centroids_top$c2[2])^2
)

cat("Group sizes:\n")
print(centroids_top %>% select(type, n))
cat("\nCentroids:\n")
print(centroids_top %>% select(type, c1, c2))
cat("\nCentroid distance (PC1–PC2):", round(centroid_distance_top, 5), "\n")

# --- Silhouette score in PC1–PC2 ---
pc_mat_top <- as.matrix(scores_top[, c("PC1", "PC2")])

# cluster::silhouette expects integer cluster labels
cluster_labels_top <- as.integer(scores_top$type)  # 1/2

# Need at least 2 points per group to compute silhouette meaningfully
tab <- table(cluster_labels_top)
if (any(tab < 2)) {
  warning("One of the groups has <2 points; silhouette may be undefined or uninformative.")
}

sil_obj_top <- silhouette(cluster_labels_top, dist(pc_mat_top))
sil_mean_top <- mean(sil_obj_top[, "sil_width"])

cat("Mean silhouette (PC1–PC2):", round(sil_mean_top, 8), "\n")

```
This results is a little depressing. Let's explore other PCs

```{r silhouette-more-pcs, message=FALSE, warning=FALSE}
library(cluster)

# labels: 1/2
cluster_labels <- as.integer(scores_top$type)

sil_for_k <- function(k) {
  pcs <- as.matrix(pca_sae_top$x[, 1:k, drop = FALSE])
  mean(silhouette(cluster_labels, dist(pcs))[, "sil_width"])
}

ks <- c(2, 5, 10, 20, 50, 100, 200, min(320, ncol(pca_sae_top$x)))
sil_vals <- sapply(ks, sil_for_k)

data.frame(k_PCs = ks, mean_silhouette = sil_vals)

```

Let's look at the global centroid distance. 
```{r centroid-distance-320d-tidyverse, message=FALSE, warning=FALSE}
library(tidyverse)

# Top-feature matrix (non-constant only), standardized
Xz <- emb_matrix_top %>%
  as_tibble() %>%
  select(all_of(non_const_cols_top)) %>%
  mutate(across(everything(), scale)) %>%
  mutate(is_neuropeptide = as.logical(df[[flag_col]]))

# Centroids (means) in standardized feature space
centroids_320d <- Xz %>%
  group_by(is_neuropeptide) %>%
  summarise(across(all_of(non_const_cols_top), ~ mean(.x, na.rm = TRUE)), .groups = "drop")

# Ensure we have exactly two groups
stopifnot(nrow(centroids_320d) == 2)

# Euclidean distance between centroids
centroid_dist_320d <- centroids_320d %>%
  select(-is_neuropeptide) %>%
  as.matrix() %>%
  { .[1, ] - .[2, ] } %>%
  { sqrt(sum(.^2)) }

centroids_320d %>%
  mutate(group = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide")) %>%
  select(group, everything())

cat("Centroid distance in standardized", length(non_const_cols_top), "D space:",
    round(centroid_dist_320d, 5), "\n")

```

```{r centroid-distance-320d-raw-tidyverse, message=FALSE, warning=FALSE}
library(tidyverse)

# Same as above but WITHOUT standardization (often less interpretable)
Xraw <- emb_matrix_top %>%
  as_tibble() %>%
  select(all_of(non_const_cols_top)) %>%
  mutate(is_neuropeptide = as.logical(df[[flag_col]]))

centroids_raw <- Xraw %>%
  group_by(is_neuropeptide) %>%
  summarise(across(all_of(non_const_cols_top), ~ mean(.x, na.rm = TRUE)), .groups = "drop")

stopifnot(nrow(centroids_raw) == 2)

centroid_dist_raw <- centroids_raw %>%
  select(-is_neuropeptide) %>%
  as.matrix() %>%
  { .[1, ] - .[2, ] } %>%
  { sqrt(sum(.^2)) }

cat("Centroid distance in raw", length(non_const_cols_top), "D space:",
    round(centroid_dist_raw, 5), "\n")

```




#----------------- end of top 320 feaures PCA figure---------------------------------






```{r sae-layer18-density-plot, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)  # for theme_cowplot, optional but matches your style

scores <- scores %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

# Background = non-neuropeptides, foreground = neuropeptides
nonneu_data <- dplyr::filter(scores, type == "Non-neuropeptide")
neuro_data  <- dplyr::filter(scores, type == "Neuropeptide")

# Sizes for legend
sizes <- c("Non-neuropeptide" = 1,
           "Neuropeptide"     = 2)

ggplot() +
  # Light background points for non-neuropeptides
  geom_point(
    data = nonneu_data,
    aes(PC1, PC2, size = "Non-neuropeptide"),
    color = "#CAD8EF",
    stroke = 0.2
  ) +
  # 2D density polygons for non-neuropeptides
  stat_density_2d(
    data = nonneu_data,
    aes(PC1, PC2, fill = factor(after_stat(level))),
    geom     = "polygon",
    contour  = TRUE,
    bins     = 10,
    key_glyph = "blank"
  ) +
  # Neuropeptides as highlighted points on top
  geom_point(
    data = neuro_data,
    aes(PC1, PC2, size = "Neuropeptide"),
    shape = 21,
    color = "#70685d",
    fill  = "#ffb22e",
    stroke = 0.25
  ) +
  # Density fill palette (no legend for fill)
  scale_fill_manual(
    values = c(
      '#BACAE7', '#ACBCD8', '#A2B1CD', '#93A1BC', '#8494AF',
      '#7787A1', '#6A7B95', '#60708A', '#52647F', '#3C597E'
    ),
    guide = "none"
  ) +
  # Size legend only for the two groups
  scale_size_manual(values = sizes) +
  guides(
    size = guide_legend(
      override.aes = list(
        fill  = c("#CAD8EF", "#ffb22e"),
        color = c("#8aa4ba", "#70685d")
      )
    )
  ) +
  scale_x_continuous(name = "PC1") +
  scale_y_continuous(name = "PC2") +
  labs(
    size  = "",
    title = "SAE embeddings (layer 18 model)"
  ) +
  theme_cowplot(16) +
  theme(
    legend.position = c(0.8, 0.2)
  )
```

Neuropeptides basically live in the same region of SAE space as the majority of the secretome.
They’re a small, slightly shifted cloud inside a larger blob, not a totally separate island.

#=====================================================================================================
# read ranking file and make PCA with top 320 features 
#=====================================================================================================

```{r select-top-features, message=FALSE}
library(dplyr)
library(readr)

rank_df <- read_csv("/Volumes/T7 Shield/layer_18_F1_scores/winning_thresholds_l18.csv",
                    show_col_types = FALSE)

names(rank_df)

# Rank by winning_F1_score (highest first), take top 320
top_features <- rank_df %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 320)

# 'feature' is a numeric index (1..10240)
top_indices <- as.integer(top_features$feature)
length(top_indices)
head(top_indices)

```


```{r align-top-features}
# Sanity check: indices are within range
range(top_indices)
length(emb_cols)

if (max(top_indices) > length(emb_cols) || min(top_indices) < 1) {
  stop("Some feature indices are out of range of embedding columns.")
}

# Map indices to column names
top_feats_in_df <- emb_cols[top_indices]

length(top_feats_in_df)  # should be 320
head(top_feats_in_df)
```
```{r pca-sae-top320}
# Matrix of just the selected top features
emb_matrix_top <- as.matrix(df[, top_feats_in_df])

# Drop zero-variance columns
sds_top <- apply(emb_matrix_top, 2, sd)
non_const_cols_top <- top_feats_in_df[sds_top > 0]

cat("Dropped", sum(sds_top == 0), "constant columns among top features\n")

# PCA on selected, non-constant features
pca_sae_top <- prcomp(df[, non_const_cols_top], scale. = TRUE)

# Variance explained
var_explained_top <- (pca_sae_top$sdev^2) / sum(pca_sae_top$sdev^2)
pc1_var_top <- round(var_explained_top[1] * 100, 1)
pc2_var_top <- round(var_explained_top[2] * 100, 1)

# Build scores with PC1 & PC2 and labels
scores_top <- as_tibble(pca_sae_top$x[, 1:2]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]]
  )

colnames(scores_top)[1:2] <- c("PC1", "PC2")

head(scores_top)
```


---

## Plot PC1 vs PC2 using the top-320 PCA

```{r plot-sae-pca-top320}
scores_top <- scores_top %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

ggplot() +
  # Non-neuropeptides: blue fill, dark blue outline, under
  geom_point(
    data = dplyr::filter(scores_top, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#062c5e",          # outline color
    shape = 21,                 # filled circle with border
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  # Neuropeptides: orange fill, orange outline, on top
  geom_point(
    data = dplyr::filter(scores_top, type == "Neuropeptide"),
    aes(x = PC1, y = PC2, fill = type),
    color = "#ff5e00",          # outline color
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5
  ) +
  scale_fill_manual(
    name   = "",
    values = c("Non-neuropeptide" = "#477aad",
               "Neuropeptide"     = "#ffb145")
  ) +
  theme_minimal() +
  labs(
    x = paste0("PC1 (", pc1_var_top, "%)"),
    y = paste0("PC2 (", pc2_var_top, "%)"),
    title = "SAE embeddings (top 320 features)"
  )
```

```{r sae-pca-density-plot, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(cowplot)  # for theme_cowplot

# Starting from scores_top (PC1, PC2, is_neuropeptide)
scores_top <- scores_top %>%
  mutate(
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

# Split into background (non-neuropeptide) and foreground (neuropeptide)
nonneu_data <- filter(scores_top, type == "Non-neuropeptide")
neuro_data  <- filter(scores_top, type == "Neuropeptide")

# Sizes for legend
sizes <- c("Non-neuropeptide" = 1,
           "Neuropeptide"     = 2)

p_sae_density <- ggplot() +
  # Light background points for non-neuropeptides
  geom_point(
    data = nonneu_data,
    aes(PC1, PC2, size = "Non-neuropeptide"),
    color = "#CAD8EF",
    stroke = 0.2
  ) +
  # 2D density polygons ("2D histogram" style) for non-neuropeptides
  stat_density_2d(
    data = nonneu_data,
    aes(PC1, PC2, fill = factor(after_stat(level))),
    geom     = "polygon",
    contour  = TRUE,
    bins     = 10,
    key_glyph = "blank"
  ) +
  # Neuropeptides as highlighted points on top
  geom_point(
    data = neuro_data,
    aes(PC1, PC2, size = "Neuropeptide"),
    shape = 21,
    color = "#70685d",
    fill  = "#ffb22e",
    stroke = 0.25
  ) +
  # Custom fill scale for the density levels (no legend for fill)
  scale_fill_manual(
    values = c(
      '#BACAE7', '#ACBCD8', '#A2B1CD', '#93A1BC', '#8494AF',
      '#7787A1', '#6A7B95', '#60708A', '#52647F', '#3C597E'
    ),
    guide = "none"
  ) +
  # Sizes for legend
  scale_size_manual(values = sizes) +
  guides(
    size = guide_legend(
      override.aes = list(
        fill  = c("#CAD8EF", "#ffb22e"),
        color = c("#8aa4ba", "#70685d")
      )
    )
  ) +
  scale_x_continuous(name = "PC1") +
  scale_y_continuous(name = "PC2") +
  labs(size = "", title = "SAE embeddings (top 320 features)") +
  theme_cowplot(16) +
  theme(
    legend.position = "none"
  )

p_sae_density
```
Neuropeptides basically live in the same region of SAE space as the majority of the secretome.
They’re a small, slightly shifted cloud inside a larger blob, not a totally separate island.

```{r}
embedding_cols <- grep("^e_", names(df), value = TRUE)
length(embedding_cols)      # how many embedding columns?
head(embedding_cols)        # first few
tail(embedding_cols)        # last few

```


# which PC is aligned with the label? 
```{r}
# scores_top from your PCA on top 320 features
label_num <- as.numeric(scores_top$is_neuropeptide)  # 0/1

cors <- apply(pca_sae_top$x, 2, function(pc) cor(pc, label_num))
head(sort(cors^2, decreasing = TRUE))
```


```{r}
# cors: correlation of each PC with the label_num
# rank PCs by |correlation|, largest first
order_idx <- order(abs(cors), decreasing = TRUE)
order_idx[1:5]   # see the top 5 PCs by label correlation

best_pc  <- order_idx[1]  # this is 1 (PC1)
second_pc <- order_idx[2] # next best

best_pc
second_pc

scores_top_lab <- as_tibble(pca_sae_top$x[, c(1, second_pc)]) %>%
  mutate(
    Entry = df$Entry,
    is_neuropeptide = df[[flag_col]],
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

colnames(scores_top_lab)[1:2] <- c("PC1", "PC_best2")

ggplot() +
  geom_point(
    data = dplyr::filter(scores_top_lab, type == "Non-neuropeptide"),
    aes(x = PC1, y = PC_best2, fill = type),
    color = "#062c5e",
    shape = 21,
    alpha = 0.4,
    size = 2,
    stroke = 0.5
  ) +
  geom_point(
    data = dplyr::filter(scores_top_lab, type == "Neuropeptide"),
    aes(x = PC1, y = PC_best2, fill = type),
    color = "#ff5e00",
    shape = 21,
    alpha = 0.9,
    size = 2,
    stroke = 0.5
  ) +
  scale_fill_manual(
    name = "",
    values = c("Non-neuropeptide" = "#477aad",
               "Neuropeptide"     = "#ffb145")
  ) +
  theme_minimal() +
  labs(
    x = "PC1",
    y = paste0("PC", second_pc, " (2nd-best label aligned)"),
    title = "SAE embeddings (top 320, PC1 vs label-aligned PC)"
  )

```



Looking at the 320 features more closely:

```{r sae-logit-score, message=FALSE, warning=FALSE}
library(glmnet)
library(dplyr)
library(ggplot2)

X <- as.matrix(df[, non_const_cols_top])
y <- as.numeric(df[[flag_col]])  # TRUE/FALSE -> 1/0
y <- ifelse(y, 1, 0)

set.seed(1)
fit <- cv.glmnet(X, y, family = "binomial", alpha = 0)

# Predicted probability of being a neuropeptide
p_hat <- as.numeric(predict(fit, newx = X, s = "lambda.min", type = "response"))

score_df <- tibble(
  prob_neuropeptide = p_hat,
  is_neuropeptide   = df[[flag_col]]
)

ggplot(score_df, aes(x = prob_neuropeptide, fill = is_neuropeptide)) +
  geom_density(alpha = 0.4) +
  scale_fill_manual(values = c("FALSE" = "#477aad", "TRUE" = "#ffb145"),
                    name   = "",
                    labels = c("Non-neuropeptide", "Neuropeptide")) +
  theme_minimal() +
  labs(
    x = "Predicted probability (neuropeptide)",
    y = "Density",
    title = "Classifier separation using top 320 SAE features"
  )
```



If those features are really good for prediction, this plot will show **two density curves shifted apart**. That’s the “separation” your classifier sees, even if PCA doesn’t show two distinct blobs.

---

## 2) Supervised projection: LDA on the top-320 SAE features

LDA explicitly finds directions that separate classes. So instead of “does PC1 separate labels?”, you’re now saying “give me the axis that *best* separates neuropeptides from others.”

```{r sae-lda-fixed3, message=FALSE, warning=FALSE}
library(MASS)
library(dplyr)
library(ggplot2)

# 1) Labels
y <- as.factor(df[[flag_col]])  # TRUE/FALSE -> factor
table(y)

# 2) Start with globally non-constant top features
X_all <- df[, non_const_cols_top, drop = FALSE]

# 3) Drop features that are (almost) constant within any class
eps <- 1e-4  # tolerance for "constant within group"

const_within_groups <- sapply(X_all, function(col) {
  # within-class SDs
  sds <- tapply(col, y, sd)
  # treat NA or very small SD as constant
  any(is.na(sds) | sds < eps)
})

cat("Features dropped as constant-within-group (tol =", eps, "):",
    sum(const_within_groups), "\n")

good_cols <- names(const_within_groups)[!const_within_groups]
cat("Features kept for LDA:", length(good_cols), "\n")

# If we dropped too many and are left with nothing, fail gracefully
if (length(good_cols) == 0) {
  stop("No features left after filtering constant-within-group predictors.")
}

# 4) Build X matrix with the filtered predictors
X <- as.matrix(X_all[, good_cols, drop = FALSE])

# 5) Fit LDA
lda_fit <- lda(x = X, grouping = y)

# 6) LDA scores
lda_scores <- as.data.frame(predict(lda_fit)$x) %>%
  mutate(
    is_neuropeptide = df[[flag_col]],
    type = if_else(is_neuropeptide, "Neuropeptide", "Non-neuropeptide"),
    type = factor(type, levels = c("Non-neuropeptide", "Neuropeptide"))
  )

# 7) Plot: LD1 vs LD2 if available, else 1D density on LD1
if ("LD2" %in% names(lda_scores)) {
  ggplot(lda_scores, aes(x = LD1, y = LD2, fill = type)) +
    geom_point(
      shape = 21,
      alpha = 0.7,
      size = 2,
      color = "black",
      stroke = 0.3
    ) +
    scale_fill_manual(
      values = c("Non-neuropeptide" = "#477aad",
                 "Neuropeptide"     = "#ffb145"),
      name = ""
    ) +
    theme_minimal() +
    labs(
      x = "LD1",
      y = "LD2",
      title = "LDA projection (top 320 SAE features, filtered with tolerance)"
    )
} else {
  ggplot(lda_scores, aes(x = LD1, fill = type)) +
    geom_density(alpha = 0.4) +
    scale_fill_manual(
      values = c("Non-neuropeptide" = "#477aad",
                 "Neuropeptide"     = "#ffb145"),
      name = ""
    ) +
    theme_minimal() +
    labs(
      x = "LD1",
      y = "Density",
      title = "LDA projection (top 320 SAE features, filtered with tolerance)"
    )
}
```