---
title: "figure_5"
output: html_document
date: "2026-01-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)

# master_entries <- read.csv(
#   "../../data/final_human_secratome_entries_1201.csv",
#   stringsAsFactors = FALSE
# ) %>%
#   select(Entry)

master_entries <- read.csv(
  "../../data/secretome_neuroflags_revised_final.csv",  #using the neuropeptide flags as the master entrires list
  stringsAsFactors = FALSE
) %>%
  select(Entry)
```


```{r}
# 2) Load neuropeptide flags (these are the revised ones from Thomas and Helena)

secr_neuro_flags <- read.csv(
  "../../data/secretome_neuroflags_revised_final.csv",
  stringsAsFactors = FALSE
) %>%
  select(
    Entry,
    neuropeptide
  )

```


```{r}
library(tidyverse)

our_model_preds <- read_csv(
  "../../data/layer_18_secretome_1201_log_reg_320_predictions.csv"
) %>%
  mutate(
    Entry = str_extract(Entry, "(?<=\\|)[^|]+(?=\\|)")
  ) %>%
  dplyr::rename(
    our_model_prob = probability
  )

```


```{r}
library(dplyr)

neuropredplm_preds <- read.csv(
  "../../data/neuropredPLM_secretome_predictions.csv",
  stringsAsFactors = FALSE
) %>%
  select(
    Entry,
    neuropredplm_class = predicted_class,
    neuropredplm_max   = max_activation,
    neuropredplm_mean  = mean_activation
  )

neuropredfrl_preds <- read.csv(
  "../../data/neuropredFRL_secretome_predictions.csv",
  stringsAsFactors = FALSE
) %>%
  select(
    Entry,
    neuropredfrl_score = prediction_score,
    neuropredfrl_class = prediction
  )

secretome_all <- master_entries %>%
  left_join(secr_neuro_flags,  by = "Entry") %>%
  left_join(our_model_preds,   by = "Entry") %>%
  left_join(neuropredplm_preds,by = "Entry") %>%
  left_join(neuropredfrl_preds,by = "Entry")

dim(secretome_all)

```

# cleaning up the data:
```{r}
secretome_all_2 <- secretome_all %>%
  select(Entry, neuropeptide, our_model_prob, neuropredfrl_score, neuropredfrl_class, neuropredplm_class) %>%
  mutate(neuropredplm_class = neuropredplm_class == 1) %>%
  mutate(neuropredfrl_class = neuropredfrl_class == "Neuropeptide") %>%
  mutate(our_model_class = our_model_prob >= 0.7)
```

Plotting out model (layer 18 InterPLM- 320 features log reg model)
```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(forcats)

# 1) Bin the probabilities into 10% buckets
binned <- secretome_all_2 %>%
  filter(!is.na(our_model_prob), !is.na(neuropeptide)) %>%
  mutate(
    prob_bin = cut(
      our_model_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = TRUE,
      labels = paste0(
        seq(0, 90, 10), "%–", seq(10, 100, 10), "%"
      )
    )
  ) %>%
  group_by(prob_bin) %>%
  summarise(
    total = n(),
    np_count = sum(neuropeptide == TRUE),
    prop_np = ifelse(total > 0, np_count / total, 0),
    .groups = "drop"
  ) %>%
  mutate(
    prob_bin = fct_explicit_na(prob_bin, na_level = "Unbinned")
  )

# 2) Plot: bar height = proportion neuropeptides; label "np/total" above bars
ggplot(binned, aes(x = fct_rev(prob_bin), y = prop_np)) +
  geom_col(
    fill  = "lavender",
    color = "grey40",
    size  = 0.3
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0,0)),
    labels = percent_format(accuracy = 1),
    limits = c(0, 0.9)
  ) +
  labs(
    x = "Predicted probability bin",
    y = "Neuropeptide proportion"
  ) +
  theme_cowplot() +
  theme(
    axis.text.x      = element_text(angle = 45, hjust = 1)
  )

```
```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(forcats)

binned_long <- secretome_all_2 %>%
  filter(!is.na(our_model_prob), !is.na(neuropeptide)) %>%
  mutate(
    prob_bin = cut(
      our_model_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = FALSE,  # makes bins [a, b)
      labels = c(
        sprintf("[%.1f-%.1f)", seq(0, 0.8, by = 0.1), seq(0.1, 0.9, by = 0.1)),
        "[0.9-1.0]"    # last bin closed to include 1.0
      )
    ),
    prob_bin = fct_rev(prob_bin),
    neuropeptide = factor(
      neuropeptide,
      levels = c(TRUE, FALSE),
      labels = c("Neuropeptide", "Non-neuropeptide")
    )
  ) %>%
  count(prob_bin, neuropeptide)

```


```{r}

label_df <- binned_long %>%
  group_by(prob_bin) %>%
  mutate(prop = n / sum(n)) %>%
  filter(neuropeptide == "Neuropeptide") %>%
  summarise(
    prop_np = sum(prop),
    .groups = "drop"
  )


plot_fig5 <- binned_long %>%
mutate(
    neuropeptide = factor(
      neuropeptide,
      levels = c("Neuropeptide", "Non-neuropeptide")
    )
  ) %>%
ggplot(aes(x = fct_rev(prob_bin), y = n, fill = fct_relevel(neuropeptide, "Neuropeptide", "Non-neuropeptide"))) + 
  geom_col(
    position = position_fill(reverse = TRUE),
    color = "grey40",
    size = 0.3,
    alpha = 0.85
  ) +
  scale_y_continuous(
    breaks = seq(0, 1, by = 0.2),
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, 0))
  ) +
  scale_x_discrete(
    expand = expansion(mult = c(0, 0))
  ) +
  geom_text(
  data = label_df,
  aes(
    x = prob_bin,
    y = prop_np,
    label = percent(prop_np, accuracy = 1)
  ),
  color = "white",
  size = 4,
  vjust = -0.5,
  inherit.aes = FALSE
  ) +
  scale_fill_manual(
    values = c( "Non-neuropeptide" = "#3d81c4",
                "Neuropeptide" = "#ffb145"
                )
  ) +
  labs(
    x = "Predicted probability bin",
    y = "Proportion of proteins",
    fill = NULL
  ) +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top",
    legend.direction = "horizontal"
  )

plot_fig5
```

Ok, now I am making the count plot (will go under the plot above)

```{r}
plot_fig5_counts <- binned_long %>%
  filter(neuropeptide == "Neuropeptide") %>%
  ggplot(aes(x = fct_rev(prob_bin), y = n)) +
  geom_col(
    fill  = "#ffb145",
    color = "grey40",
    size  = 0.5,
    alpha = 0.85
  ) +
  geom_text(
    aes(label = n),
    color = "black",
    vjust = -0.5,
    size = 4
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.1))
  ) +
  scale_x_discrete(
    expand = expansion(mult = c(0, 0.06))
  ) +
  labs(
    x = "Predicted probability bin",
    y = "Neuropeptide count"
  ) +
  theme_cowplot() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

plot_fig5_counts

```

Now let's make the cofusion matrices for the neuropred-frl and neuropred-plm models.

NeuroPredPLM:
```{r}

colnames(secretome_all_2)
```

```{r}
library(dplyr)

conf_mat <- secretome_all_2 %>%
  filter(!is.na(neuropeptide), !is.na(neuropredplm_class)) %>%
  count(
    truth = neuropeptide,
    prediction = neuropredplm_class
  )

conf_mat
```
```{r}
neuropredplm_plot <- conf_mat %>%
  mutate(
    truth = factor(
      truth,
      levels = c(TRUE, FALSE),
      labels = c("Neuro.", "Non-neuro.")
    ),
    prediction = factor(
      prediction,
      levels = c(TRUE, FALSE),
      labels = c("Neuro.", "Non-neuro.")
    )
  ) %>%
  ggplot(aes(x = prediction, y = truth, fill = n)) +
  geom_tile(color = "white") +
  scale_x_discrete(expand = expansion(mult = c(0,0))) +
  scale_y_discrete(expand = expansion(mult = c(0,0))) +
  geom_text(
    aes(
      label = n,
      color = n > max(n) * 0.6   # threshold for switching to white
    ),
    size = 5
  ) +
  scale_color_manual(
    values = c(`TRUE` = "white", `FALSE` = "black"),
    guide = "none"
  ) +
  scale_fill_gradient(low = "#f7fbff", high = "#014687") +
  labs(
    x = "Predicted label",
    y = "True label",
    fill = "Count"
  ) +
  theme_cowplot() +
  theme(
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5)
  )
neuropredplm_plot
```

```{r}
library(dplyr)

conf_mat <- secretome_all_2 %>%
  filter(!is.na(neuropeptide), !is.na(neuropredfrl_class)) %>%
  count(
    truth = neuropeptide,
    prediction = neuropredfrl_class
  )

conf_mat
```
```{r}
neuropredfrl_plot <- conf_mat %>%
  mutate(
    truth = factor(
      truth,
      levels = c(TRUE, FALSE),
      labels = c("Neuro.", "Non-neuro.")
    ),
    prediction = factor(
      prediction,
      levels = c(TRUE, FALSE),
      labels = c("Neuro.", "Non-neuro.")
    )
  ) %>%
  ggplot(aes(x = prediction, y = truth, fill = n)) +
  geom_tile(color = "white") +
  scale_x_discrete(expand = expansion(mult = c(0,0))) +
  scale_y_discrete(expand = expansion(mult = c(0,0))) +
  geom_text(
    aes(
      label = n,
      color = n > max(n) * 0.6   # threshold for switching to white
    ),
    size = 5
  ) +
  scale_color_manual(
    values = c(`TRUE` = "white", `FALSE` = "black"),
    guide = "none"
  ) +
  scale_fill_gradient(low = "#f7fbff", high = "#014687") +
  labs(
    x = "Predicted label",
    y = "True label",
    fill = "Count"
  ) +
  theme_cowplot() +
  theme(
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5)
  )
neuropredfrl_plot
```

```{r}
library(dplyr)

conf_mat <- secretome_all_2 %>%
  filter(!is.na(neuropeptide), !is.na(our_model_class)) %>%
  count(
    truth = neuropeptide,
    prediction = our_model_class
  )

conf_mat
```

```{r}
our_model_plot <- conf_mat %>%
  mutate(
    truth = factor(
      truth,
      levels = c(TRUE, FALSE),
      labels = c("Neuro.", "Non-neuro.")
    ),
    prediction = factor(
      prediction,
      levels = c(TRUE, FALSE),
      labels = c("Neuro.", "Non-neuro.")
    )
  ) %>%
  ggplot(aes(x = prediction, y = truth, fill = n)) +
  geom_tile(color = "white") +
  scale_x_discrete(expand = expansion(mult = c(0,0))) +
  scale_y_discrete(expand = expansion(mult = c(0,0))) +
  geom_text(
    aes(
      label = n,
      color = n > max(n) * 0.6   # threshold for switching to white
    ),
    size = 5
  ) +
  scale_color_manual(
    values = c(`TRUE` = "white", `FALSE` = "black"),
    guide = "none"
  ) +
  scale_fill_gradient(low = "#f7fbff", high = "#014687") +
  labs(
    x = "Predicted label",
    y = "True label",
    fill = "Count"
  ) +
  theme_cowplot() +
  theme(
    axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5)
  )
our_model_plot
```





Calculating the statistics for each model:

```{r}
calc_classification_metrics <- function(df, truth_col, pred_col) {

  df <- df %>%
    dplyr::filter(
      !is.na({{ truth_col }}),
      !is.na({{ pred_col }})
    )

  TP <- sum(df[[deparse(substitute(truth_col))]] == TRUE  &
            df[[deparse(substitute(pred_col))]]  == TRUE)

  TN <- sum(df[[deparse(substitute(truth_col))]] == FALSE &
            df[[deparse(substitute(pred_col))]]  == FALSE)

  FP <- sum(df[[deparse(substitute(truth_col))]] == FALSE &
            df[[deparse(substitute(pred_col))]]  == TRUE)

  FN <- sum(df[[deparse(substitute(truth_col))]] == TRUE  &
            df[[deparse(substitute(pred_col))]]  == FALSE)

  tibble::tibble(
    Accuracy  = (TP + TN) / (TP + TN + FP + FN),
    Precision = TP / (TP + FP),
    Recall    = TP / (TP + FN),
    F1_score  = 2 * Precision * Recall / (Precision + Recall)
  )
}

```

```{r}
library(dplyr)

metrics_neuropredplm <- calc_classification_metrics(
  secretome_all_2,
  truth_col = neuropeptide,
  pred_col  = neuropredplm_class
)

metrics_neuropredfrl <- calc_classification_metrics(
  secretome_all_2,
  truth_col = neuropeptide,
  pred_col  = neuropredfrl_class
)

metrics_l18 <- calc_classification_metrics(
  secretome_all_2,
  truth_col = neuropeptide,
  pred_col  = our_model_class
)

bind_rows(
  NeuropredPLM        = metrics_neuropredplm,
  NeuropredFRL        = metrics_neuropredfrl,
  `L18 InterPLM LogReg` = metrics_l18,
  .id = "Model"
)
```


#-------------------- making a compound figure -----------------------------------

```{r}
#plot_fig5, plot_fig5_counts, our_model_plot, neuropredfrl_plot, neuropredplm_plot

our_model_noleg <- our_model_plot + theme(legend.position = "none")
neuropredfrl_noleg <- neuropredfrl_plot + theme(legend.position = "none")
neuropredplm_noleg <- neuropredplm_plot + theme(legend.position = "none")


```

```{r}
library(cowplot)

confmat_legend <- get_legend(
  our_model_plot +
    theme(
      legend.position = "right",
      legend.title = element_text(size = 10),
      legend.text  = element_text(size = 9)
    )
)

```

```{r}
bottom_row <- plot_grid(
  our_model_noleg,
  neuropredfrl_noleg,
  neuropredplm_noleg,
  confmat_legend,
  nrow = 1,
  rel_widths = c(1, 1, 1, 0.35),
  labels = c("c.", "d.", "e.", ""),
  label_size = 14,
  label_fontface = "bold",
  label_x = 0.02,
  label_y = 0.98,
  hjust = 0,
  vjust = 1
)
```

```{r}
top_row <- plot_grid(
  plot_fig5,
  plot_fig5_counts,
  nrow = 1,
  rel_widths = c(1, 1),
  labels = c("a.", "b."),
  label_size = 14,
  label_fontface = "bold",
  label_x = 0.02,
  label_y = 0.98,
  hjust = 0,
  vjust = 1
)
```

```{r}
final_figure <- plot_grid(
  top_row,
  bottom_row,
  ncol = 1,
  rel_heights = c(1.2, 1)
)

final_figure
```

Saving the plot:
```{r}
ggsave(
  filename = "../../figures/paper_figures/figure_5_predictions.png",
  plot = final_figure,
  width = 11,
  height = 7,
  dpi = 400
)
```


#-------------------------------------------------------------------------------
# SUPPLEMENTARY FIGURE
#-------------------------------------------------------------------------------


```{r}
library(dplyr)
library(ggplot2)
library(cowplot)

x_thresh <- 0.5
y_thresh <- 0.5

plot_df <- secretome_all_2 %>%
  filter(
    !is.na(our_model_prob),
    !is.na(neuropredfrl_score)
  ) %>%
  mutate(
    type = if_else(neuropeptide, "Neuropeptide", "Non-neuropeptide")
  )

sup_cor_plot <- ggplot(plot_df, aes(x = our_model_prob, y = neuropredfrl_score)) +

  # Non-neuropeptides (draw first)
  geom_point(
    data = filter(plot_df, type == "Non-neuropeptide"),
    aes(fill = type),
    color  = "#062c5e",
    shape  = 21,
    alpha  = 0.4,
    size   = 2,
    stroke = 0.5
  ) +

  # Neuropeptides (draw on top)
  geom_point(
    data = filter(plot_df, type == "Neuropeptide"),
    aes(fill = type),
    color  = "#ff5e00",
    shape  = 21,
    alpha  = 0.9,
    size   = 2,
    stroke = 0.5
  ) +

  # Quadrants
  geom_vline(xintercept = x_thresh, linetype = "dashed", color = "grey40") +
  geom_hline(yintercept = y_thresh, linetype = "dashed", color = "grey40") +

  # Scales
  scale_fill_manual(
    name   = "",
    values = c(
      "Non-neuropeptide" = "#477aad",
      "Neuropeptide"     = "#ffb145"
    )
  ) +
  scale_x_continuous(
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.03))
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    expand = expansion(mult = c(0, 0.03))
  ) +

  labs(
    x = "InterPLM LogReg prediction",
    y = "Neuropred-FRL prediction"
  ) +

  theme_cowplot() +
  theme(
    legend.position = "top",
    legend.direction = "horizontal"
  )

sup_cor_plot

```
```{r}
df_cor <- secretome_all_2 %>%
  dplyr::filter(
    !is.na(our_model_prob),
    !is.na(neuropredfrl_score)
  )
```

```{r}
pearson_cor <- cor.test(
  df_cor$our_model_prob,
  df_cor$neuropredfrl_score,
  method = "pearson"
)

pearson_cor$estimate
pearson_cor$p.value

```

```{r}
ggsave(
  filename = "../../figures/paper_figures/sup_fig_3_cor.png",
  plot = sup_cor_plot,
  width = 8,
  height = 7,
  dpi = 400
)
```


#----------- I need to look at the top predictions (their annotations) ----------------------

Adding the secretome annotations:

Loading the secratome annotations:
```{r}
secratome_annotations <- read.delim(file = "../../data/secratome_annotations.tsv", sep="\t", na = c("", "NA", "#N/A"))
```

Let's clean the secratome annotations and then join the log reg predictions. 

```{r}
final_secratome_df <- secratome_annotations %>%
  dplyr::select(
    Entry,
    Sequence,
    Length,
    Signal.peptide,
    Peptide,
    Protein.names,
    Function..CC.,
    Gene.Ontology..molecular.function.
  ) %>%
  dplyr::rename(
    Signal_peptide   = Signal.peptide,
    Protein_names    = Protein.names,
    Protein_function = Function..CC.,
    Go_function      = Gene.Ontology..molecular.function.
  ) %>%
  dplyr::select(
    Entry,
    Protein_names,
    Protein_function,
    Go_function,
    Signal_peptide,
    Peptide,
    Sequence,
    Length
  )
```


```{r restrict-to-flagged-entries, message=FALSE, warning=FALSE}
library(tidyverse)

final_with_preds <- final_secratome_df %>%
  # Limit to only entries present in the flags file
  inner_join(secr_neuro_flags, by = "Entry") %>%
  # Add predictions for those entries
  left_join(our_model_preds %>% select(Entry, our_model_prob), by = "Entry") %>%
  select(Entry, our_model_prob, neuropeptide, Protein_names, everything())

# Sanity checks
final_with_preds %>%
  summarise(
    n_total = n(),
    n_with_pred = sum(!is.na(our_model_prob)),
    n_missing_pred = sum(is.na(our_model_prob))
  )

final_with_preds %>%
  count(neuropeptide)

```
```{r}
nrow(final_with_preds)
```

```{r top-15-by-prob, message=FALSE, warning=FALSE}
library(tidyverse)

top15 <- final_with_preds %>%
  filter(!is.na(our_model_prob)) %>%
  arrange(desc(our_model_prob)) %>%
  transmute(
    Entry = Entry,
    prediction = our_model_prob,
    neuropeptide = neuropeptide,
    Protein_names = Protein_names
  ) %>%
  slice_head(n = 20)

top15 %>% print(n = 20)
```
Now, here are the worst ones (neuropeptides that recieved <0.9)

```{r missed-neuropeptides-human, message=FALSE, warning=FALSE}
library(tidyverse)

missed_neuropeptides <- final_with_preds %>%
  filter(
    neuropeptide == TRUE,
    !is.na(our_model_prob),
    our_model_prob < 0.9
  ) %>%
  arrange(our_model_prob) %>%  # lowest (worst) first
  transmute(
    Entry = Entry,
    prediction = our_model_prob,
    neuropeptide = neuropeptide,
    Protein_names = Protein_names
  )

# How many were "missed" under this threshold?
missed_neuropeptides %>%
  summarise(n_missed = n())

# Print the worst 20 (optional)
missed_neuropeptides %>% slice_head(n = 30) %>% print(n = 30)

# Or save/export if needed:
# write_csv(missed_neuropeptides, "../../results/human_secretome_missed_neuropeptides_lt0p9.csv")
```

Here are the famouse neuropeptides:
```{r}
known_neuro <- final_with_preds %>%
  filter(Entry %in% c("A6XGL2", "P01178", "P61278", "P01185", "P05019", "P01303")) %>%
  select(Entry, our_model_prob, Protein_names)
known_neuro
```

# Peptides that are NOT neuropeptides
```{r}
 high_pred_false <- final_with_preds %>%
  filter(our_model_prob >= 0.9) %>%
  filter(neuropeptide == FALSE)

high_pred_false

 low_pred_true <- final_with_preds %>%
  filter(our_model_prob < 0.9) %>%
  filter(neuropeptide == TRUE)
 
 low_pred_true
  
 write.csv(
  high_pred_false, "../../data/secretome_high_pred_false.csv",
  row.names = FALSE
)
 
 write.csv(
  low_pred_true, "../../data/secretome_low_pred_true.csv",
  row.names = FALSE
)
```



```{r boxplot-length-by-bin, message=FALSE, warning=FALSE}
library(tidyverse)

bin_labels <- c(
  sprintf("[%.1f-%.1f)", seq(0, 0.8, by = 0.1), seq(0.1, 0.9, by = 0.1)),
  "[0.9-1.0]"  # include 1.0
)

final_with_preds %>%
  filter(!is.na(our_model_prob), !is.na(Length)) %>%
  mutate(
    prob_bin = cut(
      our_model_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = FALSE,
      labels = bin_labels
    ),
    prob_bin = factor(prob_bin, levels = bin_labels)  # force correct order
  ) %>%
  ggplot(aes(x = prob_bin, y = Length)) +
  geom_boxplot(outlier.alpha = 0.25) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Predicted probability bin",
    y = "Peptide length (aa)",
    title = "Peptide length distribution across predicted probability bins"
  )

```



size distribution before clean-up:
```{r boxplot-length-by-bin-precleanup, message=FALSE, warning=FALSE}
library(tidyverse)

# Bin labels in ascending order
bin_labels <- c(
  sprintf("[%.1f-%.1f)", seq(0, 0.8, by = 0.1), seq(0.1, 0.9, by = 0.1)),
  "[0.9-1.0]"  # include 1.0
)

# Pre-cleanup join: annotations + old predictions (no flag filtering)
precleanup_with_preds <- final_secratome_df %>%
  inner_join(
    our_model_preds %>% select(Entry, our_model_prob),
    by = "Entry"
  )

# Sanity check: should be ~1200 rows
precleanup_with_preds %>% summarise(n = n())

# Plot
precleanup_with_preds %>%
  filter(!is.na(our_model_prob), !is.na(Length)) %>%
  mutate(
    prob_bin = cut(
      our_model_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = FALSE,
      labels = bin_labels
    ),
    prob_bin = factor(prob_bin, levels = bin_labels)
  ) %>%
  ggplot(aes(x = prob_bin, y = Length)) +
  geom_boxplot(outlier.alpha = 0.25) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(
    x = "Predicted probability bin",
    y = "Peptide length (aa)",
    title = "Pre-cleanup peptide length distribution across predicted probability bins"
  )

```
```{r}
nrow(precleanup_with_preds)
```

