---
title: "group_2_part_2"
output: html_document
date: "2025-05-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
library(glue)
options(scipen = 999)
library(colorspace)
```

This script re-labels the positive dataset to also include domain labels. Next, it looks at F1 scores produced by a hybrid model-
where the precision is single residue-based and the recall is domain-based. 


Note: go to "group_2_analysis.Rmd and find remake the right_after_signal df!!!!!!
Then, continue with the following code:

=========================== POSITIVE DATASET ===========================================================================

Getting the final cleavage site df for creating labels (positive)
```{r}
final_cleavage_site_df <- right_after_signal %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) # filtering out mature peptides
```

Getting the positive data labels for the cleavage sites
```{r}
library(stringr)

positive_labels <- final_cleavage_site_df %>%
  # for each row, build a list of the 4 flanking positions
  rowwise() %>%
    mutate(
      # build a tiny tibble of flank positions + their labels
      flank = list(tibble::tibble(
        residue_number = c(as.numeric(peptide_start) - 2,
                           as.numeric(peptide_start) - 1,
                           as.numeric(peptide_end)   + 1,
                           as.numeric(peptide_end)   + 2),
        cleavage       = c("cleavage_1",   
                           "cleavage_1",
                           "cleavage_2",
                           "cleavage_2"),
        cleavage_pos       = c("first\nouter",   
                           "first\ninner",
                           "second\ninner",
                           "second\nouter")
      ))
    ) %>%
    unnest(cols = c(flank)) %>% 
    # drop any cleavage_1 when start == 1 or after_signal == TRUE
    filter(!(cleavage == "cleavage_1" & (peptide_start == 1 | peptide_after_signal))) %>%
    # drop any cleavage_2 when end == Length
    filter(!(cleavage == "cleavage_2" & peptide_end == Length)) %>%
    # also drop any positions that fall outside 1…Length
    filter(residue_number >= 1, residue_number <= Length) %>%
    # extract the amino acid at each flank position
    mutate(residue = str_sub(Sequence, residue_number, residue_number))
    
positive_labels_2 <- positive_labels %>%
  select(Entry, Sequence, Length, residue, residue_number) %>%
  distinct()

positive_labels_save <- positive_labels_2 %>%
  select(Entry, residue, residue_number) %>%
  ungroup() %>%
  distinct()

```


```{r}
positive_labels_domain <- positive_labels %>%
  # define each window (domain) by Entry + cleavage site + flank type
  # within the same sequence, same peptide (start and end) and cleavage site, this is a single domain
  mutate(domain_string = paste(Entry, peptide_start, peptide_end, cleavage, sep = "_")) %>%
  group_by(domain_string) %>%
    # this gives each domain a unique integer
    mutate(domain_id = cur_group_id()) %>%
  ungroup() #%>%
  #select(-domain_string)
```

saving the new positive labels (for hybrid model)
```{r}
positive_labels_save_domain <- positive_labels_domain %>%
  select(Entry, residue, residue_number, domain_id) %>%
  distinct()

write_csv(positive_labels_save_domain, "../../data/concept_groups/sequences/group_2/training/group_2_positive_domain.csv")

```

====================================================================================================
# ========================== Modified F1 scores computed in python. ================================
====================================================================================================
To compute F1 scores (using single residues/positions for the precision and domains for the recall), 
run the following script in python: "get_training_F1_scores_modified.py"


```{r}
winning_F1_modified <- read.csv(file = "../../data/concept_groups/F1_scores/group_2/modified_F1/winning_thresholds.csv")
# new modified: 
# old unmodified: winning F1 is 0.67 (feature 4614)
```

Now let's look at the winning features for this modified method (per-residue precision and domain recall)

#--------------------------- FEATURE 7986 -------------------------------------------------------------------
```{r}
feature <- 7986
```

Loading ALL activations for feature 7986 (won with F1 of 0.88)
```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

Some additional F1 info about feature 7986:

Threshold -> F1
0.0 -> 0.882
0.15 -> 0.624
0.30 -> 0.112
0.40 -> 0.011
0.50 -> 0.0
0.60 -> 0.0
0.80 -> 0.0

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```
```{r}

means <- feature_df %>%
  group_by(dataset) %>%
  summarise(mean_activation = mean(activation, na.rm=TRUE)) %>%
  pull(mean_activation)

feature_df %>%
  ggplot(aes(x = dataset, y = activation, color = dataset, fill = dataset)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    #geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Dataset",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -1,
      hjust     = 3,
      size      = 4,
      color     = "black"
    ) +
    geom_hline(yintercept = means, 
               color = "gray50", 
               linetype = "dashed",
               size       = 0.5) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```
What is the percent zeros in both datasets?
```{r}
feature_df %>%
  group_by(dataset) %>%
  summarise(
    n               = n(),
    non_zero        = sum(activation != 0),
    percent_non_zero = non_zero / n * 100
  )
```

For feature 7986, I want to see what the distribution of activations is for the first outer, first inner, second inner and second outer cleavage residues. 
NOTE: get the cleavage_positions df from "group_2_analysis.Rmd"

```{r}
# joining this dataframe to the activations df
joined_cleavage_positions <- left_join(cleavage_positions, feature_df)
```
Now let's plot the distribution of activations per cleavage site position:
```{r, warning=FALSE}
joined_cleavage_positions <- joined_cleavage_positions %>%
  mutate(
    cleavage_pos = fct_relevel(
      cleavage_pos,
      "first\nouter",
      "first\ninner",
      "second\ninner",
      "second\nouter"
    )
  )

joined_cleavage_positions %>%
  ggplot(aes(x = cleavage_pos, y = activation, color = cleavage_pos, fill = cleavage_pos)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Cleavage site",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -0.5,
      hjust     = 2,
      size      = 4,
      color     = "black"
    ) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```

#--------------------------- FEATURE 260 -------------------------------------------------------------------

```{r}
feature <- 260
```

```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```
```{r}

means <- feature_df %>%
  group_by(dataset) %>%
  summarise(mean_activation = mean(activation, na.rm=TRUE)) %>%
  pull(mean_activation)

feature_df %>%
  ggplot(aes(x = dataset, y = activation, color = dataset, fill = dataset)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    #geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Dataset",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -1,
      hjust     = 3,
      size      = 4,
      color     = "black"
    ) +
    geom_hline(yintercept = means, 
               color = "gray50", 
               linetype = "dashed",
               size       = 0.5) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```
What is the percent zeros in both datasets?
```{r}
feature_df %>%
  group_by(dataset) %>%
  summarise(
    n               = n(),
    non_zero        = sum(activation != 0),
    percent_non_zero = non_zero / n * 100
  )
```

For feature 260, I want to see what the distribution of activations is for the first outer, first inner, second inner and second outer cleavage residues. 
NOTE: get the cleavage_positions df from "group_2_analysis.Rmd"

```{r}
# joining this dataframe to the activations df
joined_cleavage_positions <- left_join(cleavage_positions, feature_df)
```
Now let's plot the distribution of activations per cleavage site position:
```{r, warning=FALSE}
joined_cleavage_positions <- joined_cleavage_positions %>%
  mutate(
    cleavage_pos = fct_relevel(
      cleavage_pos,
      "first\nouter",
      "first\ninner",
      "second\ninner",
      "second\nouter"
    )
  )

joined_cleavage_positions %>%
  ggplot(aes(x = cleavage_pos, y = activation, color = cleavage_pos, fill = cleavage_pos)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Cleavage site",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -0.5,
      hjust     = 2,
      size      = 4,
      color     = "black"
    ) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```
#--------------------------- FEATURE 9365 -------------------------------------------------------------------

```{r}
feature <- 9365
```

```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```
```{r}

means <- feature_df %>%
  group_by(dataset) %>%
  summarise(mean_activation = mean(activation, na.rm=TRUE)) %>%
  pull(mean_activation)

feature_df %>%
  ggplot(aes(x = dataset, y = activation, color = dataset, fill = dataset)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    #geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Dataset",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -1,
      hjust     = 3,
      size      = 4,
      color     = "black"
    ) +
    geom_hline(yintercept = means, 
               color = "gray50", 
               linetype = "dashed",
               size       = 0.5) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```
What is the percent zeros in both datasets?
```{r}
feature_df %>%
  group_by(dataset) %>%
  summarise(
    n               = n(),
    non_zero        = sum(activation != 0),
    percent_non_zero = non_zero / n * 100
  )
```

For feature 7986, I want to see what the distribution of activations is for the first outer, first inner, second inner and second outer cleavage residues. 
NOTE: get the cleavage_positions df from "group_2_analysis.Rmd"

```{r}
# joining this dataframe to the activations df
joined_cleavage_positions <- left_join(cleavage_positions, feature_df)
```
Now let's plot the distribution of activations per cleavage site position:
```{r, warning=FALSE}
joined_cleavage_positions <- joined_cleavage_positions %>%
  mutate(
    cleavage_pos = fct_relevel(
      cleavage_pos,
      "first\nouter",
      "first\ninner",
      "second\ninner",
      "second\nouter"
    )
  )

joined_cleavage_positions %>%
  ggplot(aes(x = cleavage_pos, y = activation, color = cleavage_pos, fill = cleavage_pos)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Cleavage site",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -0.5,
      hjust     = 2,
      size      = 4,
      color     = "black"
    ) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```










