---
title: "group_1_analysis"
output: html_document
date: "2025-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
```

# Group 1: Neuropeptide hormone activity

## R Markdown

========================================================================================
Overall: (percent of features that had ANY activation)
----------------------------------------------------------------------------------------
31.62% of the features had some activation in the positive (neuropeptide hormone activity)
31.64% of the features had some activation in the negative dataset (non-neuropeptides)

for 70% of the features, the maximum activation was zero
========================================================================================

Now, let's find the proportions of non-zero activation along the length of the proteins.

```{r}
activation_summary_data <- read.csv(file = "../../data/concept_groups/stats/group_1/summary_activation_bins_positive.csv")

```


# These are the proportion of all non-zero activations within a specific sequence range along the seqence. 
(for the positive dataset!)

```{r}
# Compute proportion of non-zero activations
activation_summary_data <- activation_summary_data %>%
  mutate(proportion_non_zero = non_zero_count / (zero_count + non_zero_count))

# Plot
proportion_plot <- activation_summary_data %>%
  ggplot(aes(x = position_bin, y = proportion_non_zero)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = round(proportion_non_zero, 4)), angle = 45, vjust = -0.3, hjust = 0, size = 3) +
  labs(
    title = "Proportion of Non-Zero Activations per Position Bin",
    x = "Position Bin",
    y = "Proportion Non-Zero"
  ) +
  scale_x_discrete(expand = expansion(mult = c(0,0.05))) +
  ylim(0, 0.016) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

proportion_plot
```

We can see that the embeddings (10,240 features) are extremely sparse! About 1% of the features per residue have some activation. 

We also want to see the number of sequences that fall into each bin- it is not the same. 
(for the positive dataset!)

```{r}
sequence_count_plot <- activation_summary_data %>%
  ggplot(aes(x = position_bin, y = sequence_count)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  geom_text(aes(label = sequence_count), 
            vjust = -0.3, size = 3) +
  labs(
    title = "Number of Sequences per Position Bin",
    x = "Position Bin",
    y = "Sequence Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

sequence_count_plot
```


Now, taking the features with any activation (max-pooled data- where we only kept the highest activation for each feature in a sequence), we want to look at the distribution of max activations across the length of the protein. 
**(for the positive dataset!)**

Loading data- HUGE! do it once! DO NOT RERUN!!
```{r}
activation_distribution_data <- read.csv(file = "../../data/concept_groups/stats/group_1/activation_distribution_bins.csv")

```

Now let's plot boxplots of max activation values across the sequences
```{r}

activation_distribution_data <- activation_distribution_data %>%
  mutate(bin_start = as.numeric(str_extract(bin, "^\\d+"))) %>%
  arrange(bin_start) %>%
  mutate(bin = factor(bin, levels = unique(bin))) %>%
  select(-bin_start) 

activation_distribution_boxplot <- activation_distribution_data %>%
  ggplot(aes(x = bin, y = max_activation)) +
  geom_boxplot(fill = "lightblue", outlier.color = "maroon", outlier.size = 1) +
  labs(
    title = "Distribution of Max Activations per Position Bin",
    x = "Position Bin",
    y = "Max Activation"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

activation_distribution_boxplot
```

## I did not normalize the activations!!! This will be important! You want normalized activations so that you can compare features between sequences. 

For now, let's look at the number of activations per bin. 

```{r}
activation_counts <- activation_distribution_data %>%
  group_by(bin) %>%
  summarise(activation_count = n()) %>%
  ungroup()

activation_counts <- activation_counts %>%
  mutate(bin_start = as.numeric(str_extract(bin, "^\\d+"))) %>%
  arrange(bin_start) %>%
  mutate(bin = factor(bin, levels = unique(bin))) %>%
  select(-bin_start) 

activation_count_plot <- activation_counts %>%
  ggplot(aes(x = bin, y = activation_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = activation_count), angle = 90, vjust = 0.5, hjust = -0.05, size = 3.5) +
  labs(
    title = "Number of Max Activations per Position Bin",
    x = "Position Bin",
    y = "Activation Count"
  ) +
  ylim(0, 2500000) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

activation_count_plot

```

Now let's repeat the last two plots but with the negative dataset!! 

Loading data- HUGE! do it once! DO NOT RERUN!!
```{r}
activation_distribution_data_neg <- read.csv(file = "../../data/concept_groups/stats/group_1/activation_distribution_bins_negative.csv")

```

Now let's plot boxplots of max activation values across the negative dataset sequences
```{r}

activation_distribution_data_neg <- activation_distribution_data_neg %>%
  mutate(bin_start = as.numeric(str_extract(bin, "^\\d+"))) %>%
  arrange(bin_start) %>%
  mutate(bin = factor(bin, levels = unique(bin))) %>%
  select(-bin_start) 

activation_distribution_boxplot <- activation_distribution_data_neg %>%
  ggplot(aes(x = bin, y = max_activation)) +
  geom_boxplot(fill = "lightblue", outlier.color = "maroon", outlier.size = 1) +
  labs(
    title = "Distribution of Max Activations per Position Bin",
    x = "Position Bin",
    y = "Max Activation"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

activation_distribution_boxplot
```
```{r}
activation_counts <- activation_distribution_data_neg %>%
  group_by(bin) %>%
  summarise(activation_count = n()) %>%
  ungroup()

activation_counts <- activation_counts %>%
  mutate(bin_start = as.numeric(str_extract(bin, "^\\d+"))) %>%
  arrange(bin_start) %>%
  mutate(bin = factor(bin, levels = unique(bin))) %>%
  select(-bin_start) 

activation_count_plot <- activation_counts %>%
  ggplot(aes(x = bin, y = activation_count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = activation_count), angle = 90, vjust = 0.5, hjust = -0.05, size = 3.5) +
  labs(
    title = "Number of Max Activations per Position Bin",
    x = "Position Bin",
    y = "Activation Count"
  ) +
  ylim(0, 2500000) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

activation_count_plot

```
Next, I wanted to look at the features that activate across the majority of neuropeptides and see if any of those are exclusive to the neuropeptides. 

```{r}
top_50_features <- read.csv(file = "../../data/concept_groups/stats/group_1/top50_feature_stats.csv")
```

First, we want to filter for the positive dataset. Then we want to look at the top 50 features (across the x-axis) and their proportions in 
the positive dataset (y-axis)

```{r, fig.width=6, fig.height=8}
library(scales)

# 2) Filter positive and get one row per feature
df_pos <- top_50_features %>%
  filter(dataset == "positive") %>% 
  distinct(feature, percent_of_dataset)

# 3) Order features by descending proportion
df_pos <- df_pos %>%
  arrange(rev(desc(percent_of_dataset))) %>%
  mutate(feature = factor(feature, levels = feature))

# 4) Plot
ggplot(df_pos, aes(y = feature, x = percent_of_dataset)) +
  geom_col(fill = "steelblue") +
  scale_x_continuous(                  
    labels = label_percent(accuracy = 1),
    expand = expansion(c(0, 0.15))
  ) +
  coord_cartesian(xlim = c(0.75, 1)) + 
  labs(
    y = "Feature ID",
    x = "Proportion of positive sequences\nwith at least one non-zero Activation"
  ) +
   geom_text(
    aes(label = percent(percent_of_dataset)),
    hjust = -0.1,        # push labels just outside the bar
    size  = 3
   ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    aspect.ratio   = 2
  )
```
```{r, fig.width=6, fig.height=8}
library(tidyverse)
library(scales)

# 2) Filter positive and get one row per feature
df_pos <- top_50_features %>%
  filter(dataset == "negative") %>% 
  distinct(feature, percent_of_dataset)

# 3) Order features by descending proportion
df_pos <- df_pos %>%
  arrange(rev(desc(percent_of_dataset))) %>%
  mutate(feature = factor(feature, levels = feature))

# 4) Plot
ggplot(df_pos, aes(y = feature, x = percent_of_dataset)) +
  geom_col(fill = "steelblue") +
  scale_x_continuous(                  
    labels = label_percent(accuracy = 1),
    expand = expansion(c(0, 0.15))
  ) +
  coord_cartesian(xlim = c(0.75, 1)) + 
  labs(
    y = "Feature ID",
    x = "Proportion of negative sequences\nwith those features"
  ) +
   geom_text(
    aes(label = percent(percent_of_dataset)),
    hjust = -0.1,        # push labels just outside the bar
    size  = 3
   ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5),
    aspect.ratio   = 2
  )
```
Conclusion: most prevalent features in the positive dataset are also the most prevalent features in the negative dataset. 




# here, talk about the fact that there almost no features exclusive to the neuropeptides (not even close!!)


```{r}
winning_thresholds <- read.csv(file = "../../data/concept_groups/F1_scores/group_1/winning_thresholds.csv")
```

```{r}
clean_scores <- winning_thresholds %>%
  filter(winning_F1_score != 0)
```

All winnning thresholds are zero!!! 

```{r}
ggplot(clean_scores, aes(x = winning_F1_score)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  labs(
    title = "Distribution of winning F1 scores",
    x     = "Winning F1 Score",
    y     = "Density"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
clean_scores2 <- clean_scores %>%
  filter(winning_F1_score >= 0.2)

ggplot(clean_scores2, aes(x = winning_F1_score)) +
  geom_histogram(
    binwidth = 0.05,
    center = 0.025,
    fill     = "steelblue",
    color    = "white"
  ) +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),  # ticks every 0.1
    limits = c(0.2, 1)
  ) +
  geom_text(
    stat     = "bin",
    binwidth = 0.05,
    center = 0.025,
    aes(label = ifelse(..count.. > 0, ..count.., NA_integer_)),
    vjust    = -0.5,
    size      = 3
  ) +
  labs(
    x     = "Winning F1 Score",
    y     = "Count (Number of features)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank()
  )
```

Now let's look at teh F1 scores produced by the maxpooled data. 

```{r}
winning_thresholds_max <- read.csv(file = "../../data/concept_groups/F1_scores/group_1/winning_thresholds_maxpooled.csv")
```

```{r}
clean_scores <- winning_thresholds_max %>%
  filter(winning_F1_score != 0)
```

# Here there are very few non-zero winning thresholds, but their F1 scores are very tiny.

```{r}
ggplot(clean_scores, aes(x = winning_F1_score)) +
  geom_density(fill = "steelblue", alpha = 0.6) +
  labs(
    title = "Distribution of winning F1 scores",
    x     = "Winning F1 Score",
    y     = "Density"
  ) +
  theme_minimal(base_size = 14)
```

```{r}
clean_scores2 <- clean_scores %>%
  filter(winning_F1_score >= 0.2)

ggplot(clean_scores2, aes(x = winning_F1_score)) +
  geom_histogram(
    binwidth = 0.05,
    center = 0.025,
    fill     = "steelblue",
    color    = "white"
  ) +
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),  # ticks every 0.1
    limits = c(0.2, 1)
  ) +
  geom_text(
    stat     = "bin",
    binwidth = 0.05,
    center = 0.025,
    aes(label = ifelse(..count.. > 0, ..count.., NA_integer_)),
    vjust    = -0.5,
    size      = 3
  ) +
  labs(
    x     = "Winning F1 Score",
    y     = "Count (Number of features)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.minor = element_blank()
  )

```

Now, let's look at feature 3652- the winning feature.

```{r}
feature_3652 <- top_50_features %>%
  filter(feature == 3652)
```

```{r}
feature_3652 %>%
  ggplot(aes(x = percent_of_dataset, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "3652 (winning feature), F1: 0.665",
      x     = "Fraction of sequences containing the feature",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```

```{r}
feature_3652 %>%
  ggplot(aes(x = percent_non_zero, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      x     = "Fraction of the sequence that activates",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```
Max activation is also the same for both positive and negative datasets for that feature. 

Now, let's look at feature 314.

```{r}
feature_314 <- top_50_features %>%
  filter(feature == 314)
```

```{r}
feature_314 %>%
  ggplot(aes(x = percent_of_dataset, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "314, F1: 0.665",
      x     = "Fraction of sequences containing the feature",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```

```{r}
feature_314 %>%
  ggplot(aes(x = percent_non_zero, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "314, F1: 0.665",
      x     = "Fraction of the sequence that activates",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```

Now, let's look at feature 3957.

```{r}
feature_3957 <- top_50_features %>%
  filter(feature == 3957)
```

```{r}
feature_3957 %>%
  ggplot(aes(x = percent_of_dataset, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "3957, F1: 0.663",
      x     = "Fraction of sequences containing the feature",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```

```{r}
feature_3957 %>%
  ggplot(aes(x = percent_non_zero, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "3957, F1: 0.663",
      x     = "Fraction of the sequence that activates",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```
Now, let's look at feature 99.

```{r}
feature_99 <- top_50_features %>%
  filter(feature == 99)
```

```{r}
feature_99 %>%
  ggplot(aes(x = percent_of_dataset, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "99, F1: 0.66",
      x     = "Fraction of sequences containing the feature",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```

```{r}
feature_99 %>%
  ggplot(aes(x = percent_non_zero, color = dataset, fill = dataset)) +
    geom_density(alpha = 0.4, size = 1) +     # semi­transparent fill
    scale_x_continuous(
      labels = label_percent(accuracy = 1),
      limits = c(0, 1)
    ) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer (type = "qual", palette = "Dark2") +
    labs(
      title = "99, F1: 0.66",
      x     = "Fraction of the sequence that activates",
      y     = "Density",
      color = "Dataset",
      fill  = "Dataset"
    ) +
    theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5),
      legend.position = "top"
    )


```
