---
title: "group_2_fine_tuned"
output: html_document
date: "2025-06-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
```


Loading in the F1 scores for the tuned models:

Model where the ESM embeddings were normalized:
```{r}
winning_F1_ft_norm <- read.csv(file = "../../data/concept_groups/F1_scores/group_2/winning_thresholds_ft_norm.csv")

```

Model where the ESM embeddings were not normalized:
```{r}
winning_F1_ft_unnorm <- read.csv(file = "../../data/concept_groups/F1_scores/group_2/winning_thresholds_ft_unnorm.csv")

```

Plot comparing the distribution of F1 scores:

combining the hybrid/modified (pretrained), fine-tuned norm and fine-tuned unnorm models into a single df.
```{r}
mod_F1 <- winning_F1_modified %>%
  filter(winning_F1_score >= 0.50000) %>%
  mutate(group = "original")

ft_norm_F1 <- winning_F1_ft_norm %>%
  filter(winning_F1_score >= 0.50000) %>%
  mutate(group = "fine-tuned normalized")

ft_unnorm_F1 <- winning_F1_ft_unnorm %>%
  filter(winning_F1_score >= 0.50000) %>%
  mutate(group = "fine-tuned")

F1_all <- rbind(mod_F1, ft_unnorm_F1, ft_norm_F1)
  
```

```{r}
F1_all %>%
  group_by(group) %>%
  count()
```

```{r, warning=FALSE}
# Load required libraries
library(ggplot2)
library(ggforce)
library(colorspace)
library(forcats)

# Reorder the factor levels
F1_all <- F1_all %>%
  mutate(
    group = fct_relevel(group,
      "original",
      "fine-tuned",
      "fine-tuned normalized"
    )
  )

# Create the combined violin + sina plot
ggplot(F1_all, aes(x = group, y = winning_F1_score, fill = group)) +
  # fixed width violins
  geom_violin(trim = TRUE, alpha = 0.4, width = 1.2) +
  #geom_sina(aes(color = group), size = 0.3, alpha = 0.7) +
  # colorspace qualitative palette
  scale_fill_manual(
    values = qualitative_hcl(n = 3, palette = "Dark 3")
  ) +
  scale_color_manual(
    values = qualitative_hcl(n = 3, palette = "Dark 3")
  ) +
  scale_y_continuous(limits = c(0.49, 1.0)) +
  theme_minimal() +
  labs(
    x = "Group",
    y = "Winning F1 Score",
    title = "Distribution of Winning F1 Scores by Group"
  ) +
  theme(
    legend.position = "none",
    axis.title = element_text(size = 12),
    axis.text  = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )



```



```{r, warning=FALSE}
# Load required libraries
library(ggplot2)
library(colorspace)
library(forcats)
library(dplyr)

# Prepare the data: reorder groups and treat threshold as a factor
F1_all <- F1_all %>%
  mutate(
    group     = fct_relevel(group, "original", "fine-tuned", "fine-tuned normalized"),
    threshold = factor(winning_threshold)
  )

# Create faceted bar chart with counts above each bar
ggplot(F1_all, aes(x = threshold, fill = group)) +
  geom_bar(width = 0.7) +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    vjust = -0.5,
    size = 3
  ) +
  scale_y_continuous(limits = c(0,1350)) +
  scale_fill_manual(values = qualitative_hcl(3, "Dark 3")) +
  facet_grid(group ~ ., scales = "fixed") +
  labs(
    x     = "Winning Threshold",
    y     = "Count",
    title = "Count of Features per Winning Threshold by Group"
  ) +
  theme_minimal() +
  theme(
    legend.position   = "none",
    axis.text.x       = element_text(hjust = 1),
    axis.title        = element_text(size = 12),
    axis.text         = element_text(size = 10),
    plot.title        = element_text(size = 14, face = "bold", hjust = 0.5),
    strip.text.y.left = element_text(angle = 0, face = "bold", size = 12),
    panel.spacing.y      = unit(1, "cm") 
  )

```

```{r, warning=FALSE, fig.height=6.5, fig.width=7}
# Load required libraries
library(ggplot2)
library(dplyr)
library(forcats)
library(colorspace)
library(ggforce)  # for geom_sina
library(grid)     # for unit()

# Prepare the data: ensure group ordering and make threshold a factor
F1_all <- F1_all %>%
  mutate(
    group     = fct_relevel(group, "original", "fine-tuned", "fine-tuned normalized"),
    threshold = factor(winning_threshold)
  )

# Define a darker palette for points
base_pal   <- qualitative_hcl(3, "Dark 3")
dark_pal   <- darken(base_pal, amount = 0.3)  # make it 30% darker

# Violin + darker sina points, faceted by group
f1_vs_thresh_plot <- ggplot(F1_all, aes(x = threshold, y = winning_F1_score, fill = group)) +
  geom_violin(
    trim  = TRUE,    
    scale = "width", 
    width = 0.8,
    alpha = 0.6,
    drop  = FALSE
  ) +
  geom_sina(
    aes(color = group),
    size  = 0.3,
    alpha = 1,
    scale = "width"
  ) +
  scale_fill_manual(values = base_pal) +
  scale_color_manual(values = dark_pal) +
  facet_grid(group ~ ., scales = "fixed") +
  labs(
    x     = "Winning Threshold",
    y     = "Winning F1 Score",
    title = "Distribution of F1 Scores Across Thresholds by Model"
  ) +
  theme_minimal() +
  theme(
    legend.position     = "none",
    axis.text.x         = element_text(angle = 45, hjust = 1),
    axis.title          = element_text(size = 12),
    axis.text           = element_text(size = 10),
    plot.title          = element_text(size = 14, face = "bold", hjust = 0.5),
    strip.text.y.left   = element_text(angle = 0, face = "bold", size = 12),
    panel.spacing.y     = unit(1, "cm")
  )

ggsave(
  filename = "../../Figures/F1_thresholds_violin.png",  # desired output file
  plot     = last_plot(),                 # or e.g. p if you assigned your ggplot to `p`
  width    = 6,                           # width in inches
  height   = 8,                           # height in inches
  units    = "in",
  dpi      = 400                          # resolution
)

```

#=======================================================================================================================
# Here we have trained 5 models (runs 2-6). I will be analyzing run6 for first.
#=======================================================================================================================

# I have found the F1 scores across all 10 folds for the neuro toxin dataset and using the finetuned SAE embeddings.
Now, I will compare the F1 scores of the finetuned and not original models.

Let's review the original F1 scores. 

```{r}
winning_F1_orig_exc_fold_9 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/F1_scores_graphpart_40/excluded_fold_9/winning_thresholds_excluded_fold_9.csv")
```

```{r}
winning_F1_ft_exc_fold_9_run_6 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/F1_scores_hybrid/F1_scores_graphpart_40/run_6/excluded_fold_9/winning_thresholds_excluded_fold_9.csv")

winning_F1_ft_exc_fold_9_run_5 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/F1_scores_hybrid/F1_scores_graphpart_40/run_5/excluded_fold_9/winning_thresholds_excluded_fold_9.csv")

winning_F1_ft_exc_fold_9_run_4 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/F1_scores_hybrid/F1_scores_graphpart_40/run_4/excluded_fold_9/winning_thresholds_excluded_fold_9.csv")

winning_F1_ft_exc_fold_9_run_3 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/F1_scores_hybrid/F1_scores_graphpart_40/run_3/excluded_fold_9/winning_thresholds_excluded_fold_9.csv")

winning_F1_ft_exc_fold_9_run_2 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/F1_scores_hybrid/F1_scores_graphpart_40/run_2/excluded_fold_9/winning_thresholds_excluded_fold_9.csv")
```

```{r}
winning_all_models <- bind_rows(
  winning_F1_orig_exc_fold_9        %>% mutate(model = "original"),
  winning_F1_ft_exc_fold_9_run_2    %>% mutate(model = "run_2"),
  winning_F1_ft_exc_fold_9_run_3    %>% mutate(model = "run_3"),
  winning_F1_ft_exc_fold_9_run_4    %>% mutate(model = "run_4"),
  winning_F1_ft_exc_fold_9_run_5    %>% mutate(model = "run_5"),
  winning_F1_ft_exc_fold_9_run_6    %>% mutate(model = "run_6")
) %>%
  relocate(model, .before = 1)

# (optional) keep a nice order for plotting/tables
winning_all_models$model <- factor(winning_all_models$model,
                            levels = c("original","run_2","run_3","run_4","run_5","run_6"))
```


Now, let's make a figure comparing the Finetuned and original F1 scores of both models. 

```{r, warning=FALSE}
# Load required libraries
library(ggplot2)
library(ggforce)
library(colorspace)
library(forcats)

# Create the combined violin + sina plot
ggplot(winning_all_models, aes(x = model, y = winning_F1_score, fill = model)) +
  # fixed width violins
  geom_violin(trim = TRUE, alpha = 0.4, width = 1.2) +
  #geom_sina(aes(color = group), size = 0.3, alpha = 0.7) +
  # colorspace qualitative palette
  scale_fill_manual(
    values = qualitative_hcl(n = 6, palette = "Dark 3")
  ) +
  scale_color_manual(
    values = qualitative_hcl(n = 6, palette = "Dark 3")
  ) +
  scale_y_continuous(limits = c(0.49, 1.0)) +
  theme_minimal() +
  labs(
    x = "Model",
    y = "Winning F1 Score",
    title = "Distribution of Winning F1 Scores by Model"
  ) +
  theme(
    legend.position = "none",
    axis.title = element_text(size = 12),
    axis.text  = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )



```



Now let's look at the top 100 features in all groups

```{r}
top100_original <- winning_F1_orig_exc_fold_9 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(model = "original")

top100_run_2 <- winning_F1_ft_exc_fold_9_run_2 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(model = "run_2")

top100_run_3 <- winning_F1_ft_exc_fold_9_run_3 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(model = "run_3")

top100_run_4 <- winning_F1_ft_exc_fold_9_run_4 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(model = "run_4")

top100_run_5 <- winning_F1_ft_exc_fold_9_run_5 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(model = "run_5")

top100_run_6 <- winning_F1_ft_exc_fold_9_run_6 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(model = "run_6")

# combine into one data.frame
top100_orig_run2 <- bind_rows(top100_original, top100_run_2)
top100_orig_run3 <- bind_rows(top100_original, top100_run_3)
top100_orig_run4 <- bind_rows(top100_original, top100_run_4)
top100_orig_run5 <- bind_rows(top100_original, top100_run_5)
top100_orig_run6 <- bind_rows(top100_original, top100_run_6)
```

```{r}
set.seed(42)  # so the jitter is reproducible

# 1) turn your bin factor into numeric positions
# 2) add a jitter offset (the same for line & point)
top100_orig_run4 <- top100_orig_run4 %>%
  mutate(
    model      = factor(model, levels = c("original","run_4")),
    dataset_num  = as.numeric(model),
    x_jitter     = dataset_num + runif(n(), -0.15, 0.15)
  )

```

```{r}
# 1) Flag how many bins each feature appears in
top100_flagged <- top100_orig_run4 %>%
  add_count(feature, name = "n_bins") %>%
  mutate(
    share_group = factor(
      n_bins,
      levels = 1:2,
      labels = c(
        "Unique",
        "Shared"
      )
    )
  )
```

```{r}
library(ggrepel)

ggplot(top100_flagged, aes(group = feature)) +
  # 1) Orange “Unique” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Unique"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 2) Dark-green “Shared” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Shared"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 3) Labels for Shared in the neuropeptide dataset
  geom_text_repel(
    data = filter(top100_flagged,
                  share_group == "Shared",
                  model     == "original"),  # make sure this matches your factor
    aes(x = x_jitter, y = winning_F1_score, label = feature),
    size          = 3,
    nudge_y       = 0.05,
    nudge_x       = -0.05,
    segment.color = "grey60",
    segment.size  = 0.3,
    max.overlaps  = Inf
  ) +
  # 4) Lines underneath
  geom_line(
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 0.4, alpha = 0.6
  ) +
  # 5) Point-style legend
  scale_color_manual(
    name   = "Feature sharing",
    values = c("Unique" = "orange", "Shared" = "darkgreen"),
    guide  = guide_legend(
      override.aes = list(
        linetype = 0,   # no line
        shape    = 16,  # solid circle
        size     = 4
      )
    )
  ) +
  # 6) X axis
  scale_x_continuous(
    breaks = 1:2,
    labels = levels(top100_flagged$model),
    name   = "Model",
    expand = expansion(add = 0.2)
  ) +
  labs(y = "Winning F1 Score") +
  theme_minimal() +
  theme(
    legend.position  = "right",
    legend.title     = element_blank(),
    legend.text     = element_text(size = 11),
    axis.title.x     = element_blank(),
    axis.text.x      = element_text(size = 12),
    axis.title.y     = element_text(size = 12)
  )

```


```{r}
# 1) Compute slice fractions and “mid‐angle” rotated so 0 = 12 o’clock
levels_order <- c(
  "Unique",
  "Shared"
)

pie_data <- top100_flagged %>%
  distinct(feature, share_group) %>%
  dplyr::count(share_group) %>%
  mutate(
    share_group = factor(share_group, levels = levels_order),
    fraction    = n / sum(n)
  ) %>%
  arrange(share_group) %>%
  mutate(
    ymin    = cumsum(lag(fraction, default = 0)),
    ymax    = cumsum(fraction),
    # shift start to 12 o’clock by subtracting 1/4 circle (pi/2)
    midang  = pi/2 - 2 * pi * (ymin + fraction/2),
    pct     = fraction * 100,
    label   = paste0(n, " (", round(pct), "%)")
  )

total_features <- sum(pie_data$n)

# 2) Draw pie + outside labels at radius 1.3
ggplot(pie_data) +
  geom_arc_bar(
    aes(
      x0     = 0, y0 = 0,
      r0     = 0, r  = 1,
      amount = n,
      fill   = share_group
    ),
    stat  = "pie",
    color = "white",
    size  = 0.2
  ) +
  geom_text(
    aes(
      x     = 1.3 * cos(midang),
      y     = 1.3 * sin(midang),
      label = label
    ),
    size        = 4,
    color       = "black",
    inherit.aes = FALSE
  ) +
  annotate(
    "text",
    x     = 0, y     = -1.4,
    label = paste0("Total features: ", total_features),
    size  = 5,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Unique"            = "orange",
      "Shared"   = "darkgreen"
    ),
    guide = guide_legend(
      title        = NULL,
      override.aes = list(linetype = 0, shape = 16, size = 4)
    )
  ) +
  coord_fixed() +
  xlim(-1.5, 1.5) +
  ylim(-1.5, 1.5) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.text     = element_text(size = 11),
    plot.margin     = margin(10, 10, 10, 10)
  )

```

# Now I am going to plot the logistic regression results for the new finetuned models:


Loading validation results
```{r}
percent_identity <- 40
run <- 3

base_dir <- file.path(
  paste0("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/classifier_results/run_", run, "/"),
  "log_regression_csv_results",
  paste0("graphpart_", percent_identity)
)

# 1) Locate all per‐fold validation CSVs
files <- list.files(
  path       = base_dir,
  pattern    = "logistic_regression_results_excluded_fold_\\d+\\.csv$",
  recursive  = TRUE,
  full.names = TRUE
)

# 2) Read & combine
all_val_results <- files %>%
  set_names() %>% 
  map_dfr(~ {
    df <- read_csv(.x, show_col_types = FALSE)
    fold_num <- str_extract(basename(.x), "\\d+") %>% as.integer()
    df %>% mutate(fold = fold_num)
  })

```

plotting the validation results:
```{r}
ggplot(all_val_results, aes(x = num_features, y = validation, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Average Validation Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )

```


Test results:
Loading all the test results from the 40% nested k-fold cross validation
```{r}
# install.packages("tidyverse")  # if you don't already have it
library(tidyverse)

# 1) Point to the directory where your fold sub-folders live:
base_dir <- paste0("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/finetuned_SAE_analysis/classifier_results/run_", run, "/test_log_regression_csv_results/graphpart_40")

# 2) Find all the CSV files (one per fold):
files <- list.files(
  path       = base_dir, 
  pattern    = "^linear_regression_test_results_test_fold_\\d+\\.csv$", 
  recursive  = TRUE,
  full.names = TRUE
)

# 3) Read & combine, extracting the fold number from each filename:
all_results <- files %>% 
  set_names() %>% 
  map_dfr(read_csv, .id = "file", show_col_types = FALSE) %>% 
  mutate(
    fold = str_extract(basename(file), "\\d+")
  ) %>% 
  select(fold, num_features, test_accuracy)
```

Ploting the graph:
```{r}
ggplot(all_results, aes(x = num_features, y = test_accuracy, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Test Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```


Getting df with the average test accuracy:
```{r}
avg_results <- all_results %>%
  # if test_accuracy is character, first turn it numeric:
  mutate(test_accuracy = as.numeric(test_accuracy)) %>%
  group_by(num_features) %>%
  summarise(
    avg_test_accuracy = mean(test_accuracy, na.rm = TRUE),
    .groups = "drop"
  )
```

plotting the average test result:
```{r}
ggplot(
  data = avg_results,
  aes(x = num_features, y = avg_test_accuracy)
  ) +
  geom_line(size = 0.8, color = "darkmagenta") +
  labs(
    x = "Number of Features",
    y = "Average Test Accuracy (%)"
  ) +
  theme_minimal()
```

Detecting the final number of features I need to use:
(Elbow/knee method)
```{r}
# all_results should have: fold, num_features, test_accuracy
# Make sure accuracy is numeric (not character):
all_results <- all_results %>%
  mutate(test_accuracy = as.numeric(test_accuracy))

# 1) Summarise across folds → mean, sd, se per num_features
avg_results <- all_results %>%
  group_by(num_features) %>%
  summarise(
    avg_test_accuracy = mean(test_accuracy, na.rm = TRUE),
    sd_acc            = sd(test_accuracy,   na.rm = TRUE),
    n_folds           = dplyr::n(),
    se_acc            = sd_acc / sqrt(n_folds),
    .groups = "drop"
  ) %>%
  arrange(num_features)

```

```{r}
# ---- Method A: Elbow/knee on slope ----
slopes <- diff(avg_results$avg_test_accuracy) / diff(avg_results$num_features)
threshold <- 0.001  # e.g., 0.1% gain per 10 features (adjust as you wish)

idx <- which(slopes < threshold)
opt_n_elbow <- if (length(idx)) {
  # take the first time the marginal gain drops below threshold
  avg_results$num_features[idx[1] + 0]
} else {
  # fallback: never dropped → pick the largest tested (or choose your own rule)
  max(avg_results$num_features)
}

opt_n_elbow
```

















