---
title: "group_2_part_4"
output: html_document
date: "2025-05-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
```

Here, I try using only a single flanking residue around the mature peptide as the cleavage site. 

# start with "right_after_signal" from "group_2_analysis.Rmd"
=========================== POSITIVE DATASET ===========================================================================

Getting the final cleavage site df for creating labels (positive)
```{r}
final_cleavage_site_df <- right_after_signal %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) # filtering out mature peptides
```

Getting the positive data labels for the cleavage sites
```{r}
library(stringr)

positive_labels <- final_cleavage_site_df %>%
  # for each row, build a list of the 4 flanking positions
  rowwise() %>%
    mutate(
      # build a tiny tibble of flank positions + their labels
      flank = list(tibble::tibble(
        residue_number = c(as.numeric(peptide_start) - 1,
                           as.numeric(peptide_end)   + 1),
        cleavage       = c("cleavage_1",   
                           "cleavage_2")
      ))
    ) %>%
    unnest(cols = c(flank)) %>% 
    # drop any cleavage_1 when start == 1 or after_signal == TRUE
    filter(!(cleavage == "cleavage_1" & (peptide_start == 1 | peptide_after_signal))) %>%
    # drop any cleavage_2 when end == Length
    filter(!(cleavage == "cleavage_2" & peptide_end == Length)) %>%
    # also drop any positions that fall outside 1…Length
    filter(residue_number >= 1, residue_number <= Length) %>%
    # extract the amino acid at each flank position
    mutate(residue = str_sub(Sequence, residue_number, residue_number))
    
positive_labels_2 <- positive_labels %>%
  select(Entry, Sequence, Length, residue, residue_number) %>%
  distinct()

positive_labels_save <- positive_labels_2 %>%
  select(Entry, residue, residue_number) %>%
  ungroup() %>%
  distinct()

write_csv(positive_labels_save, "../../data/concept_groups/sequences/group_2/training/group_2_positive_single_flank.csv")

```

==================== NEGATIVE DATASET ================================================
Now we need to sample the negative dataset:
First we make a table that contains everything we want to exclude. 
Besides the cleavage sites, I also want to exclude the adjacent neighbors to the cleavage sites (just in case). 
```{r}
pos_labels_with_neighbors <- positive_labels_2 %>%
  select(Entry, residue_number) %>%
  # add the -1 and +1 neighbors
  mutate(
    neighbor_minus = residue_number - 1,
    neighbor_minus_2 = residue_number - 2,
    neighbor_plus  = residue_number + 1,
    neighbor_plus_2  = residue_number + 2
  ) %>%
  pivot_longer(
    cols      = c(residue_number, neighbor_minus, neighbor_minus_2, neighbor_plus, neighbor_plus_2),
    values_to = "residue_number"
  ) %>%
  filter(residue_number >= 1) %>%       # drop any <1
  select(Entry, residue_number) %>%
  distinct()
```

Now, we need to count how many cleavage site resides (positives) we have per sequence/Entry. 
Remember that some sequences have multiple alternative cleavage sites.
```{r}
positives_count <- positive_labels_2 %>%
  dplyr::count(Entry, name = "n_pos")
```


Each sequence is expanded to have one row per residue (will be used later)
I kept the sequence row as is to later extract the actual residue
```{r}
expanded_seqs <- positive_labels_2 %>%
  select(Entry, Sequence, Length) %>%
  distinct() %>%
  uncount(Length) %>%                   # replicate each row 'Length' times
  group_by(Entry) %>%
  mutate(residue_number = row_number()) %>%
  ungroup() 
```

Next, we need to remove the excluded positions
```{r}
expanded_seqs <- expanded_seqs %>%
  anti_join(pos_labels_with_neighbors, by = c("Entry", "residue_number"))
```

Here we sample the negative dataset positions. (Try not to rerun this) each time it will resample
```{r}
negative_labels <- expanded_seqs %>%
  left_join(positives_count, by = "Entry") %>% # adding the count of positive dataset residues per Entry/sequence
  group_by(Entry) %>%
  sample_n(size = unique(n_pos)) %>%    # sample same number as positives
  ungroup() %>%
  mutate(
    residue = str_sub(Sequence, residue_number, residue_number) # str_sub(string, start, end)
  ) %>%
  select(Entry, residue, residue_number)
```

Saving the sampled negative data 
```{r}
write_csv(negative_labels, "../../data/concept_groups/sequences/group_2/training/group_2_negative_single_flank.csv")

```

Making sure the sampling worked and we have equal number of positives. 
```{r}
# How many positives per Entry?
pos_counts <- positive_labels_save %>%
  dplyr::count(Entry, name = "n_pos")

# How many negatives did I actually get per Entry?
neg_counts <- negative_labels %>%
  dplyr::count(Entry, name = "n_neg")

# Join and look for any Entry where n_neg < n_pos
mismatches <- pos_counts %>%
  left_join(neg_counts, by = "Entry") %>%
  mutate(less_neg = ifelse(n_neg < n_pos, TRUE, FALSE))

mismatches          # see which ones and by how much
```

===================================================================================================================
# ========================== Single flanking residue F1 scores computed in python =================================
===================================================================================================================

```{r}
winning_F1_single_flank <- read.csv(file = "../../data/concept_groups/F1_scores/group_2/single_flank_F1/winning_thresholds.csv")
# new modified: 
# old unmodified: winning F1 is 0.67 (feature 4614)
```

#--------------------------- FEATURE 8216 -------------------------------------------------------------------

```{r}
feature <- 8216
```

```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```

```{r}

means <- feature_df %>%
  group_by(dataset) %>%
  summarise(mean_activation = mean(activation, na.rm=TRUE)) %>%
  pull(mean_activation)

feature_df %>%
  ggplot(aes(x = dataset, y = activation, color = dataset, fill = dataset)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    #geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Dataset",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -1,
      hjust     = 3,
      size      = 4,
      color     = "black"
    ) +
    geom_hline(yintercept = means, 
               color = "gray50", 
               linetype = "dashed",
               size       = 0.5) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```














