---
title: "group_2_part_3"
output: html_document
date: "2025-05-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(readr)
library(glue)
library(seqinr)    # for read.fasta()
```

Here, I try calculating F1 with both a domain-based precision and domain-based recall. 

# resampling negative dataset in pairs:
```{r, warning=FALSE}

# ─── your positive‐domain table ────────────────────────────────────────────────
positive_labels_domain <- read_csv("../../data/concept_groups/sequences/group_2/training/group_2_positive_domain.csv")

# Note: get the "right_after_fasta" df from "group_2_analysis.Rmd" right before the positive dataset is made
seq_df <- right_after_signal %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) %>% # filtering out mature peptides
  select(Entry, Sequence, Length) %>%
  distinct(Sequence, .keep_all = TRUE)

```


```{r}
# ─── join sequence info onto your positive domains ────────────────────────────
pos_dom <- positive_labels_domain %>%
  left_join(seq_df, by = "Entry")
head(pos_dom)
```

```{r}
# ─── 1) build exclusion set: all positive residues ±1 ─────────────────────────
exclusion <- pos_dom %>%
  select(Entry, residue_number) %>%
  mutate(
    neighbor_minus = residue_number - 1,
    neighbor_plus  = residue_number + 1
  ) %>%
  pivot_longer(
    c(residue_number, neighbor_minus, neighbor_plus),
    values_to = "residue_number"
  ) %>%
  filter(residue_number >= 1) %>% # getting rid of "fake neighbors" with indices of -1
  distinct(Entry, residue_number)

head(exclusion)
```

```{r}
# ─── 2) how many positive domains per Entry? ─────────────────────────────────
# each domain_id in pos_dom is one domain

positives_count <- pos_dom %>%
  distinct(Entry, domain_id) %>%
  dplyr::count(Entry, name = "n_domains")

positives_count
```

```{r}
# ─── 3) expand each sequence to all allowed positions ─────────────────────────
seq_info <- positive_labels_2 %>% 
  select(Entry, Sequence, Length) %>% 
  distinct()

expanded_seqs <- seq_info %>%
  uncount(Length, .remove = FALSE) %>%  
  group_by(Entry) %>%
  mutate(residue_number = row_number()) %>%
  ungroup() %>%
  anti_join(exclusion, by = c("Entry","residue_number"))
```

```{r}

# ─── 4) build candidate consecutive‐pair starts for the negative dataset ──────────────────────────────
candidate_pairs <- expanded_seqs %>%
  group_by(Entry, Sequence, Length) %>%
  filter(residue_number < Length) %>%      # so +1 stays in bounds
  transmute(
    start = residue_number,
    end   = residue_number + 1
  ) %>%
  # ensure both start & end rows exist in expanded_seqs
  inner_join(expanded_seqs %>% select(Entry, res = residue_number),
             by = c("Entry","start" = "res")) %>%
  inner_join(expanded_seqs %>% select(Entry, res = residue_number),
             by = c("Entry","end"   = "res")) %>%
  distinct(Entry, Sequence, Length, start, end) %>%
  ungroup()
```

Ensure that candidate pairs do not overlap with the excluded positions:
```{r}
candidate_pairs_clean <- candidate_pairs %>%
  filter(
    !start %in% exclusion$residue_number &
    !end   %in% exclusion$residue_number
  )

nrow(candidate_pairs)
nrow(candidate_pairs_clean)
  
```

```{r}
# ─── 5) sample exactly as many negative domains per Entry ─────────────────────
neg_dom <- candidate_pairs %>%
  left_join(positives_count, by = "Entry") %>%
  group_by(Entry) %>%
  sample_n(size = unique(n_domains)) %>%
  ungroup() %>%
  mutate(domain_id = row_number())
```

# ─── 6) “unnest” each pair into two rows + capture residue letter ────────────
```{r}
negative_labels_domain <- neg_dom %>%
  # grab the residue letters
  rowwise() %>%
  mutate(
    res1 = str_sub(Sequence, start, start),
    res2 = str_sub(Sequence, end,   end)
  ) %>%
  select(Entry, domain_id, start, end, res1, res2) %>%
  pivot_longer(
    c(start, end),
    names_to  = "which_pos", # is this the start (first residue) or end (second residue)?
    values_to = "residue_number"
  ) %>%
  mutate(
    residue = if_else(which_pos == "start", res1, res2) # only keep the first or second residue per row
  ) %>%
  select(Entry, residue, residue_number, domain_id) %>%
  arrange(Entry, domain_id, residue_number)
```

```{r}
# — write it out —
write_csv(negative_labels_domain, "../../data/concept_groups/sequences/group_2/training/group_2_negative_domain.csv")
```

========================================================================================================
# ========================== Domain-based F1 scores computed in python =================================
========================================================================================================

```{r}
winning_F1_domain_based <- read.csv(file = "../../data/concept_groups/F1_scores/group_2/domain_based_F1/winning_thresholds.csv")
# new modified: 
# old unmodified: winning F1 is 0.67 (feature 4614)
```

Now let's look at the winning features for this modified method (per-residue precision and domain recall)

#--------------------------- FEATURE 8284 -------------------------------------------------------------------
```{r}
feature <- 8284
```

Loading ALL activations for feature 8284 (won with F1 of 0.989)
Run the following script in python: "get_activations_for_feature.py"
```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

Some additional F1 info about feature 8284:

Threshold -> F1
0.0 -> 0.989
0.15 -> 0.972
0.30 -> 0.792
0.40 -> 0.525
0.50 -> 0.198
0.60 -> 0.043
0.80 -> 0.0008

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```
```{r}

means <- feature_df %>%
  group_by(dataset) %>%
  summarise(mean_activation = mean(activation, na.rm=TRUE)) %>%
  pull(mean_activation)

feature_df %>%
  ggplot(aes(x = dataset, y = activation, color = dataset, fill = dataset)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    #geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Dataset",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -1,
      hjust     = 3,
      size      = 4,
      color     = "black"
    ) +
    geom_hline(yintercept = means, 
               color = "gray50", 
               linetype = "dashed",
               size       = 0.5) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```
What is the percent zeros in both datasets?
```{r}
feature_df %>%
  group_by(dataset) %>%
  summarise(
    n               = n(),
    non_zero        = sum(activation != 0),
    percent_non_zero = non_zero / n * 100
  )
```

For feature 8284, I want to see what the distribution of activations is for the first outer, first inner, second inner and second outer cleavage residues. 
NOTE: get the cleavage_positions df from "group_2_analysis.Rmd"

```{r}
# joining this dataframe to the activations df
joined_cleavage_positions <- left_join(cleavage_positions, feature_df)
```
Now let's plot the distribution of activations per cleavage site position:
```{r, warning=FALSE}
joined_cleavage_positions <- joined_cleavage_positions %>%
  mutate(
    cleavage_pos = fct_relevel(
      cleavage_pos,
      "first\nouter",
      "first\ninner",
      "second\ninner",
      "second\nouter"
    )
  )

joined_cleavage_positions %>%
  ggplot(aes(x = cleavage_pos, y = activation, color = cleavage_pos, fill = cleavage_pos)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Cleavage site",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -0.5,
      hjust     = 2,
      size      = 4,
      color     = "black"
    ) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```

#------------------------------------------ FEATURE 4515 -------------------------------------------------------------------

```{r}
feature <- 4515
```

```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```

#--------------------------- FEATURE 4058 -------------------------------------------------------------------

```{r}
feature <- 4058
```

```{r}

feature_df <- read.csv(file = glue("../../data/concept_groups/stats/group_2/activations_for_feature_{feature}.csv"))
```

```{r}

feature_df %>%
  ggplot(aes(x = activation, fill = dataset, color = dataset)) +
    # use ..count.. instead of the default ..density..
    geom_density(alpha = 0.4, size = 0.2) +
    scale_fill_manual(
      values = c(
        positive = "#66C2A5",  # teal
        negative = "#FC8D62"   # orange
    )) +
    scale_x_continuous(expand = expansion(mult = c(0.0, 0.0))) +
    scale_y_continuous(
      #limits = c(0, 4000),
      expand = expansion(mult = c(0, 0))) +
    labs(
      x = "Activation",
      y = "Density",
      title = glue("Feature {feature}")
    ) +
    theme_cowplot() +
    theme(
      legend.position = c(0.8, 0.8)
    )
```

```{r}

means <- feature_df %>%
  group_by(dataset) %>%
  summarise(mean_activation = mean(activation, na.rm=TRUE)) %>%
  pull(mean_activation)

feature_df %>%
  ggplot(aes(x = dataset, y = activation, color = dataset, fill = dataset)) +
    geom_violin(scale = "count", alpha = 0.3, show.legend = FALSE) +
    #geom_sina(scale = "count", alpha = 0.6, size = 0.5, show.legend = FALSE) +
    scale_y_continuous(
      limits = c(0,1),
      expand = expansion(mult = c(0,0))
    ) +
    labs(
      x     = "Dataset",
      y     = "Activation",
      title = glue("Feature {feature}")
    ) +
    stat_summary( 
      fun   = mean,
      geom  = "point",
      shape = 23,
      size  = 2,  # diamond at the mean 
      color = "black",
      fill  = "black"
    ) +
    stat_summary(
      fun       = mean, 
      geom      = "text",
      aes(label = sprintf("%.2f", ..y..)),
      vjust     = -1,
      hjust     = 3,
      size      = 4,
      color     = "black"
    ) +
    geom_hline(yintercept = means, 
               color = "gray50", 
               linetype = "dashed",
               size       = 0.5) +
    theme_cowplot() +
    theme(
      axis.title.x = element_text(margin = margin(t = 10)),
      axis.title.y = element_text(margin = margin(r = 10)),
      plot.title   = element_text(hjust = 0.5),
      legend.position = "none"
    )
```









