---
title: "group_2_raw_activations_interactive"
output: html_document
date: "2025-05-16"
---

```{r setup, include=FALSE}
# ─── LIBRARIES ─────────────────────────────────────────────────────────────────
library(tidyverse)
library(cowplot)
library(ggrepel)
library(plotly)
library(htmlwidgets)
library(htmltools)
library(glue)
library(seqinr)
library(ggiraph)
library(fs) 
```


Here is the ggiraph version of the interactive plot:
I used this one:
```{r}

# ─── PARAMETERS & PATHS ─────────────────────────────────────────────────────────
feature         <- 9047
fasta_path      <- "../../data/concept_groups/sequences/group_2/training/cleavage_site_sequences.fasta"
activations_csv <- glue("../../data/concept_groups/stats/group_2/feature_{feature}_activations_ALL.csv")
pos_domain_csv  <- "../../data/concept_groups/sequences/group_2/training/group_2_positive_domain.csv"
#pos_domain_csv  <- "../../data/concept_groups/sequences/group_2/training/group_2_positive_triple_flank_domain.csv"
output_dir      <- glue("../../figures/group_2/")
html_out        <- file.path(output_dir, glue("feature_{feature}_interactive.html"))
dir_create(output_dir, recurse = TRUE)
```


```{r}

# ─── READ FULL SEQUENCES ─────────────────────────────────────────────────────────
seqs_list <- read.fasta(fasta_path, seqtype="AA", as.string=TRUE, strip.desc=TRUE)
seq_df    <- tibble(Entry = names(seqs_list),
                    Sequence = toupper(unlist(seqs_list)))
```

Here I need to add in annotations:
```{r}
annotations_all <- clean_data_5 %>%
  select(Entry, Organism, Function, Protein.families, protein_names) %>%
  distinct()

seq_names <- seq_df %>%
  select(Entry) %>%
  distinct()

annotations_df <- left_join(seq_names, annotations_all)

```

```{r}

# ─── READ ACTIVATIONS & EXTRACT RESIDUE ─────────────────────────────────────────
activations_df <- read_csv(activations_csv, col_types = cols(
  Entry      = col_character(),
  position   = col_integer(),
  activation = col_double()
)) %>%
  # strip trailing “.pt” from Entry
  mutate(Entry = str_remove(Entry, "\\.pt$")) %>%
  left_join(seq_df, by="Entry") %>%
  mutate(residue = str_sub(Sequence, position, position))
```


Note: These labels are going to change depending on if you are doing 1, 2 or 3 flanking residues
so modify/check the code block below carefully
```{r}
# ─── READ CLEAVAGE‐SITE LABELS DOMAINS/DOUBLE/TRIPLE FLANK ──────────────────────────────────────────────────
pos_domains_df <- read_csv(pos_domain_csv, col_types = cols(
  Entry          = col_character(),
  residue_number = col_integer(),
  residue        = col_character(),
  domain_id      = col_integer()
))

entries_for_plot <- intersect(
  activations_df$Entry %>% unique(),
  pos_domains_df$Entry %>% unique()
)
```


For the next code block, rerun the first part of "group_2_part_2.Rmd" in order to get "positive_labels_2" with one flanking residue
```{r}
# ─── READ CLEAVAGE‐SITE LABELS SINGLE FLANK ──────────────────────────────────────────────────
cleavage_single <- positive_labels_2 %>%
  select(Entry, residue, residue_number)

entries_for_plot <- intersect(
  activations_df$Entry %>% unique(),
  cleavage_single$Entry %>% unique()
)
```


```{r}

# ─── FUNCTION TO BUILD A GGIRAPH PLOT FOR ONE ENTRY ──────────────────────────
make_profile_girafe <- function(entry_id) {
  df_e    <- filter(activations_df, Entry == entry_id)
  seq_len <- max(df_e$position)
  y_max   <- max(df_e$activation, na.rm = TRUE)
  y_lim   <- if (y_max == 0) 1 else y_max * 1.2
  tick_h  <- y_lim * 0.09

  # cleavage‐site boxes
  clefts <- pos_domains_df %>%  # change this part for different flanks
    filter(Entry == entry_id) %>%
    distinct(residue_number, residue)
  
  # pull tooltip text from your annotations_df
  tooltip_title <- annotations_df %>%
    filter(Entry==entry_id) %>%
    slice(1) %>%   # just in case
    glue_data(
      "Organism: {Organism}\n",
      "Function: {Function}\n",
      "Protein family: {Protein.families}\n",
      "Protein name: {protein_names}"
    )

  rects <- tibble(
    xmin = clefts$residue_number - 0.5,
    xmax = clefts$residue_number + 0.5,
    ymin = -tick_h,
    ymax =  y_lim
  )
  
  # for each position, build a tooltip rectangle under the line
  hover_strips <- df_e %>%
    transmute(
      xmin    = position - 0.5,
      xmax    = position + 0.5,
      ymin    = -tick_h,
      ymax    = 0,
      tooltip = glue::glue(
        "Position: {position}\n",
        "Residue:  {residue}"
      ),
      data_id = paste0(Entry, "_", position)
    )

  # dynamic x‐axis breaks
  x_by     <- case_when(
    seq_len > 500 ~ 50,
    seq_len > 350 ~ 40,
    seq_len > 100 ~ 20,
    TRUE          ~ 10
  )
  x_breaks <- seq(0, seq_len, by = x_by)

  # build ggplot with interactive geoms
  p <- ggplot() +
    # 1) interactive title
    geom_text_interactive(
      data = tibble(
        x = (seq_len + 1) / 10, 
        y = y_lim * 1.08, 
        label = paste0("Feature ", feature, ": ", entry_id),
        tooltip = tooltip_title,
        data_id = entry_id
      ),
      aes(x = x, y = y, label = label, tooltip = tooltip, data_id = data_id),
      size = 4,
      inherit.aes = FALSE
    ) +
    
    # 1) neon‐green cleavage‐site highlight
    geom_rect_interactive(
      data        = rects,
      aes(
        xmin    = xmin,
        xmax    = xmax,
        ymin    = ymin,
        ymax    = ymax,
        tooltip = "Cleavage site",
        data_id = xmin   # arbitrary but unique per rect
      ),
      fill        = "#39FF14",
      color       = "gray70",
      size        = 0.1,
      alpha       = 0.5,
      inherit.aes = FALSE
    ) +
    
    # 2) invisible hover-strips under the baseline:
    geom_rect_interactive(
      data = hover_strips,
      aes(
        xmin    = xmin,
        xmax    = xmax,
        ymin    = ymin,
        ymax    = ymax,
        tooltip = tooltip,
        data_id = data_id
      ),
      fill        = "transparent",
      color       = NA,
      inherit.aes  = FALSE
    ) +

    # 2) blue bars with full‐bar hover text
    geom_col_interactive(
      data        = df_e,
      aes(
        x       = position,
        y       = activation,
        tooltip = glue(
          "Position: {position}\n",
          "Residue:  {residue}\n",
          "Activation: {round(activation,3)}"
        ),
        data_id = paste0(Entry, "_", position) 
      ),
      fill   = "steelblue",
      color  = "gray40",
      size = 0.1,
      width  = 1,
      alpha  = 0.7
    ) +

    # 3) minor tick‐marks on x
    geom_segment(
      data = tibble(
        x    = seq(0.5, seq_len + 0.5, by = 1),
        xend = x, 
        y    = 0, 
        yend = -tick_h * 0.9
      ),
      aes(x = x, xend = xend, y = y, yend = yend),
      color       = "gray40",
      size        = 0.2,
      inherit.aes = FALSE
    ) +

    # 4) baseline and residue‐labels
    geom_hline(yintercept = 0, color = "black", size = 0.3) +
    geom_text(
      data = clefts,
      aes(
        x     = residue_number,
        y     = -tick_h * 0.6,
        label = residue
      ),
      size        = 2.5,
      inherit.aes = FALSE
    ) +

    # scales, labels, theme
    scale_x_continuous(
      breaks = x_breaks,
      limits = c(0.5, seq_len + 0.5),
      expand = expansion(0,0)
    ) +
    scale_y_continuous(
      breaks = scales::pretty_breaks(6),
      #limits = c(-tick_h, y_lim),
      expand = expansion(0,0)
    ) +
    coord_cartesian(clip="off") +   
    labs(
      #title = glue("Feature {feature}: {entry_id}"),
      x     = "Residue Position",
      y     = "Activation"
    ) +
    #theme_minimal(base_size = 11) +
    theme(
      plot.margin  = margin(t = 40, r = 10, b = 10, l = 10),
      panel.grid.major   = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.line.y        = element_line(color = "black", size = 0.2),
      axis.ticks.x       = element_line(color = "black", size = 0.2),
      axis.ticks.y       = element_line(color = "black", size = 0.2),
      axis.text.x        = element_text(color = "black", margin = margin(t = 4)),
      axis.text.y        = element_text(color = "black"),
      axis.title.x = element_text(margin = margin(t = 6)),
      axis.title.y = element_text(margin = margin(r = 6)),
      panel.background   = element_rect(fill = "white", color = NA),
      plot.background    = element_rect(fill = "white", color = NA)
    )

  # wrap it in a girafe()
  girafe(
    ggobj        = p,
    width_svg    = 11,
    height_svg   = 5,
    options      = list(
      opts_hover(css = "fill-opacity:1;cursor:pointer;")
    )
  )
}

# ─── BUILD ONE WIDGET PER ENTRY & SAVE ALL TO HTML ─────────────────────────────
widgets <- map(entries_for_plot, make_profile_girafe)

spaced_widgets <- lapply(widgets, function(w){
  tags$div(
    style = "margin: 50px 20px;",  # top/bottom = 40px, left/right = 10px
    w
  )
})

all_page <- tagList(
  lapply(spaced_widgets, function(w){
    list(w, htmltools::br())
  })
)

htmltools::save_html(all_page, html_out)

```


A version using plotly (not the best) 
DO NOT USE!
```{r, warning=FALSE}

# ─── PARAMETERS & PATHS ─────────────────────────────────────────────────────────
feature         <- 7547
fasta_path      <- "../../data/concept_groups/sequences/group_2/training/cleavage_site_sequences.fasta"
activations_csv <- glue("../../data/concept_groups/stats/group_2/feature_{feature}_activations_ALL.csv")
pos_domain_csv  <- "../../data/concept_groups/sequences/group_2/training/group_2_positive_domain.csv"
output_dir      <- glue("../../figures/group_2/")
html_out        <- file.path(output_dir, glue("feature_{feature}_interactive_profiles.html"))
dir_create(output_dir, recurse = TRUE)

# ─── READ SEQUENCES ──────────────────────────────────────────────────────────────
seqs_list <- read.fasta(fasta_path, seqtype="AA", as.string=TRUE, strip.desc=TRUE)
seq_df    <- tibble(Entry=names(seqs_list), Sequence= toupper(unlist(seqs_list)))

# ─── READ ACTIVATIONS & EXTRACT RESIDUE ───────────────────────────────────────────
activations_df <- read_csv(activations_csv, col_types=cols(
  Entry      = col_character(),
  position   = col_integer(),
  activation = col_double()
)) %>%
  left_join(seq_df, by="Entry") %>%
  mutate(residue = str_sub(Sequence, position, position))

# ─── READ CLEAVAGE‐SITE LABELS ────────────────────────────────────────────────────
pos_domains_df <- read_csv(pos_domain_csv, col_types=cols(
  Entry          = col_character(),
  residue_number = col_integer(),
  residue        = col_character(),
  domain_id      = col_integer()
))

entries_for_plot <- intersect(
  unique(activations_df$Entry),
  unique(pos_domains_df$Entry)
)

# ─── BUILD ONE PROFILE (no numeric labels) ───────────────────────────────────────
make_profile_plot_interactive <- function(entry_id) {
  df_e    <- filter(activations_df, Entry==entry_id)
  seq_len <- max(df_e$position)
  y_max   <- max(df_e$activation, na.rm=TRUE)
  y_lim   <- if(y_max==0) 1 else y_max*1.2
  tick_h  <- y_lim*0.09

  # cleavage‐site backgrounds
  clefts <- pos_domains_df %>%
    filter(Entry==entry_id) %>%
    distinct(residue_number, residue)

   # 4) X‐axis breaks: 
  x_by <- case_when(
  seq_len > 500 ~ 50,
  seq_len > 350 ~ 40,
  seq_len > 100 ~ 20,
  TRUE          ~ 10
  )
  
  x_breaks <- seq(0, seq_len, by=x_by)
  
  rects <- tibble(
    xmin = clefts$residue_number - 0.5,
    xmax = clefts$residue_number + 0.5,
    ymin = -tick_h,
    ymax =  y_lim
  )


  p <- ggplot(df_e, aes(x=position,y=activation)) +
    # green cleavage‐site highlight if any
    geom_rect(
      data = rects,
      aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax,
          text = "cleavage site"),
      fill  = "#39FF14",   # neon green
      color = "gray70",
      size     = 0.1,
      alpha = 0.5,
      inherit.aes = FALSE
    ) +
    geom_col(aes(text = glue(
      "Entry: {Entry}<br>",
      "Position: {position}<br>",
      "Residue: {residue}<br>",
      "Activation: {round(activation,3)}"
    )),
    fill="steelblue", 
    color ="gray40", 
    width =1, 
    size = 0.1,
    alpha = 0.7) +
    # little minor tick segments
    geom_segment(
      data = tibble(x=seq(0.5, seq_len+0.5, by=1),
                    xend=x, y=0, yend=-tick_h),
      aes(x=x,xend=xend,y=y,yend=yend),
      color="gray40", size=0.2, inherit.aes=FALSE
    ) +
    geom_hline(yintercept=0, color="black", size=0.3) +
    # optional: residue labels at bottom
    geom_text(
        data=clefts,
        aes(x=residue_number, y=-tick_h+tick_h*0.5, label=residue),
        fontface="bold", size=3, inherit.aes=FALSE
      ) +
    scale_x_continuous(
      breaks = x_breaks,
      #minor_breaks= seq(0.5, seq_len+0.5, by = 1),
      limits = c(0.5, seq_len+0.5),
      expand = expansion(mult = c(0, 0))
    ) +
    scale_y_continuous(
      limits = c(-tick_h, y_lim),
      breaks = pretty_breaks(n = 6),
      expand = expansion(mult = c(0, 0))
    ) +
    # geom_text(
    #     data = cleft_labels,
    #     aes(x = residue_number, y = 0, label = residue),
    #     vjust = 0.5,                # push below the baseline
    #     size  = 1.7,
    #     color    = "black"
    #) +
    labs(title=glue("Feature {feature}: {entry_id}"),
         x="Residue Position", y="Activation") +
    theme(
      panel.grid.major   = element_blank(),
      panel.grid.minor   = element_blank(),
      axis.line.y   = element_line(color = "black", size = 0.2),
      axis.ticks.x     = element_line(color = "black", size = 0.2),
      axis.ticks.y     = element_line(color = "black", size = 0.2),
      
      # (optional) nudge the x labels so they don’t overlap the ticks
      axis.text.x        = element_text(margin = margin(t = 4), color = "black"),
      axis.text.y     = element_text(color = "black"),
      
      plot.title         = element_text(hjust = 0.5),
      panel.background   = element_rect(fill = "white", color = NA),
      plot.background    = element_rect(fill = "white", color = NA)
      )

  return(p)
}


# ─── CONVERT & SAVE ALL TO SINGLE HTML ──────────────────────────────────────────
widgets <- entries_for_plot %>%
  map(~ ggplotly(make_profile_plot_interactive(.x), tooltip="text"))

all_page <- tagList(widgets)
htmltools::save_html(all_page, file = html_out)


# ─── SAVE ALL WIDGETS TO SINGLE HTML ────────────────────────────────────────────
all_page <- tagList(widgets)
htmltools::save_html(all_page, file = html_out)
```



















