---
title: "known_peptides.Rmd"
output: html_document
date: "2025-08-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(tidytext)    # for reorder_within() and scale_x_reordered()
```

```{r}
known_peptide_annotations <- read.delim(file = "../../data/known_peptides_test/uniprot_neuropep_entries.tsv", sep="\t", na = c("", "NA", "#N/A"))
```

```{r}
clean_known_peptides <- known_peptide_annotations %>%
  select(Entry, Sequence, Length, Peptide) %>%
  distinct() %>%
  na.omit(Peptide)
```


Here is a list of all the peptide starts and ends. 
```{r}
clean_known_peptides2 <- clean_known_peptides %>%
  # extract all “123..456” substrings as a list-column
  mutate(range_str = str_extract_all(Peptide, "\\d+\\.\\.\\d+")) %>%
  # turn each element of that list into its own row
  unnest_longer(range_str) %>%
  # split “123..456” into two new columns
  separate(range_str, into = c("peptide_start","peptide_end"), sep = "\\.\\.") %>%
  # convert to integer
  mutate(
     start = as.integer(peptide_start),
     end   = as.integer(peptide_end)) %>%
  select(Entry, Sequence, Length, peptide_start, peptide_end)

```

get Signal sequence info: (where signal starts and ends so that later we can remove cleavage site right after signal)
```{r}
signal_info <- known_peptide_annotations %>%
  select(Entry, Signal.peptide) %>%
  distinct() %>%
  mutate(range_str = str_extract_all(Signal.peptide, "\\d+\\.\\.\\d+")) %>%
  unnest_longer(range_str) %>%
  separate(range_str, into = c("signal_start","signal_end"), sep = "\\.\\.") %>%
  mutate(
     start = as.integer(signal_start),
     end   = as.integer(signal_end)) %>%
  select(Entry, signal_start, signal_end)
```

Joining the peptide and signal info dataframes together.

```{r}
peptide_data_full <- clean_known_peptides2 %>%
  left_join(signal_info)
```

```{r}
right_after_signal_known <- peptide_data_full %>%
  mutate(peptide_after_signal = ifelse(as.numeric(signal_end) + 1 == peptide_start, TRUE, FALSE))
```  
  
Let's see how many peptides are NOT right after the signal:
```{r}
right_after_signal_known %>%
  filter(peptide_after_signal == FALSE) %>%
  count()

```
Now let's see the sequences where the start is 1 and the end is equal to the length. These are the mature peptide and we do not want these in the dataset.
```{r}
mature_peptides <- right_after_signal_known %>%
  filter(peptide_start == 1 & peptide_end == Length) # result: 938 mature peptide sequences
```

Now, let's look at the sequences where the start is 1, but the end is not equal to the length (these we want to keep because we can use the second cleavage site residues)
```{r}
semi_mature_peptides <- right_after_signal_known %>%
  filter(peptide_start == 1) %>%
  filter(peptide_end != Length)
```


=========================== POSITIVE DATASET for cleavage site annotations ========================================================================

Getting the final cleavage site df for creating labels (positive)
```{r}
final_cleavage_site_df <- right_after_signal_known %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) # filtering out mature peptides
```

Getting the positive data labels for the cleavage sites
```{r}
library(stringr)

positive_labels <- final_cleavage_site_df %>%
  # for each row, build a list of the 4 flanking positions
  rowwise() %>%
    mutate(
      # build a tiny tibble of flank positions + their labels
      flank = list(tibble::tibble(
        residue_number = c(as.numeric(peptide_start) - 2,
                           as.numeric(peptide_start) - 1,
                           as.numeric(peptide_end)   + 1,
                           as.numeric(peptide_end)   + 2),
        cleavage       = c("cleavage_1",   
                           "cleavage_1",
                           "cleavage_2",
                           "cleavage_2"),
        cleavage_pos       = c("first\nouter",   
                           "first\ninner",
                           "second\ninner",
                           "second\nouter")
      ))
    ) %>%
    unnest(cols = c(flank)) %>% 
    # drop any cleavage_1 when start == 1 or after_signal == TRUE
    filter(!(cleavage == "cleavage_1" & (peptide_start == 1 | peptide_after_signal))) %>%
    # drop any cleavage_2 when end == Length
    filter(!(cleavage == "cleavage_2" & peptide_end == Length)) %>%
    # also drop any positions that fall outside 1…Length
    filter(residue_number >= 1, residue_number <= Length) %>%
    # extract the amino acid at each flank position
    mutate(residue = str_sub(Sequence, residue_number, residue_number))
    
positive_labels_2 <- positive_labels %>%
  select(Entry, Sequence, Length, residue, residue_number) %>%
  distinct()

positive_labels_save <- positive_labels_2 %>%
  select(Entry, residue, residue_number) %>%
  ungroup() %>%
  distinct()

write_csv(positive_labels_save, "../../data/concept_groups/sequences/group_2/training/group_2_positive_known_peps.csv")
```







