---
title: "layer_18_analysis"
output: html_document
date: "2025-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(tidytext)    # for reorder_within() and scale_x_reordered()
```


Raw F1 scores: 
```{r}
```


#=========================================================================================
# VALIDATION RESULTS
#=========================================================================================

```{r}
#layer_18_log_reg <- read.csv("/Volumes/T7 Shield/layer_18_classifiers/val_results/val_results_excluded_fold_{}.csv")

base_dir <- file.path("/Volumes/T7 Shield/layer_18_classifiers/val_results")

# 1) Locate all per‐fold validation CSVs
files <- list.files(
  path       = base_dir,
  pattern    = "val_results_excluded_fold_\\d+\\.csv$",
  recursive  = TRUE,
  full.names = TRUE
)

# 2) Read & combine
all_val_results <- files %>%
  set_names() %>% 
  map_dfr(~ {
    df <- read_csv(.x, show_col_types = FALSE)
    fold_num <- str_extract(basename(.x), "\\d+") %>% as.integer()
    df %>% mutate(fold = fold_num)
  })

```

```{r}
ggplot(all_val_results, aes(x = num_features, y = validation, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_val_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Average Validation Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )

```
#------------------------------------------------------------------------
# TEST RESULTS:
#------------------------------------------------------------------------
```{r}
base_dir <- file.path("/Volumes/T7 Shield/layer_18_classifiers/test_results")

# 2) Find all the CSV files (one per fold):
files <- list.files(
  path       = base_dir, 
  pattern    = "test_results_excluded_fold_\\d+\\.csv$", 
  recursive  = TRUE,
  full.names = TRUE
)

# 3) Read & combine, extracting the fold number from each filename:
all_results <- files %>% 
  set_names() %>% 
  map_dfr(read_csv, .id = "file", show_col_types = FALSE) %>% 
  mutate(
    fold = str_extract(basename(file), "\\d+")
  ) %>% 
  select(fold, num_features, test_accuracy)
```

Ploting the graph:
```{r}
ggplot(all_results, aes(x = num_features, y = test_accuracy, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Test Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```

Average test results:
```{r}
avg_results <- all_results %>%
  # if test_accuracy is character, first turn it numeric:
  mutate(test_accuracy = as.numeric(test_accuracy)) %>%
  group_by(num_features) %>%
  summarise(
    avg_test_accuracy = mean(test_accuracy, na.rm = TRUE),
    .groups = "drop"
  )
```

plotting the average test result:
```{r}
ggplot(
  data = avg_results,
  aes(x = num_features, y = avg_test_accuracy)
  ) +
  geom_line(size = 0.8, color = "darkmagenta") +
  labs(
    x = "Number of Features",
    y = "Average Test Accuracy (%)"
  ) +
  theme_minimal()
```




```{r}

library(dplyr)
library(ggplot2)

# --- 0) Ensure clean & sorted; coerce to percent if needed ---
df <- avg_results %>%
  transmute(
    num_features = as.integer(num_features),
    acc = as.numeric(avg_test_accuracy)
  ) %>%
  arrange(num_features)

# If accuracy is in [0,1], convert to percentage points
if (max(df$acc, na.rm = TRUE) <= 1.001) {
  df <- mutate(df, acc = acc * 100)
}

# --- 1) Slope-threshold elbow ---
# Smooth slope to reduce noise (simple trailing mean of length k)
k <- 3
kernel <- rep(1/k, k)

df <- df %>%
  mutate(
    d_acc  = c(NA, diff(acc)),
    d_feat = c(NA, diff(num_features)),
    slope  = d_acc / d_feat,                           # pp (percentage points) per feature
    slope_smooth = as.numeric(stats::filter(slope, kernel, sides = 1))
  )

threshold <- 0.01  # 0.01 pp/feature = 0.1% per 10 features
idx1 <- which(df$slope_smooth < threshold)[1]
elbow_slope <- if (!is.na(idx1)) df$num_features[idx1] else max(df$num_features)

# --- 2) Max-distance ("kneedle") elbow ---
# Geometric distance from each point to the line between endpoints
x <- df$num_features
y <- df$acc

x1 <- x[1];  y1 <- y[1]
x2 <- x[length(x)]; y2 <- y[length(y)]

# Handle degenerate case where all x are equal (shouldn't happen here)
if (x2 == x1 && y2 == y1) {
  elbow_kneedle <- x1
} else {
  # Distance from point (x0,y0) to line through (x1,y1)-(x2,y2)
  num <- abs((y2 - y1)*x - (x2 - x1)*y + x2*y1 - y2*x1)
  den <- sqrt((y2 - y1)^2 + (x2 - x1)^2)
  dist_to_line <- num / den

  elbow_kneedle <- x[which.max(dist_to_line)]
}

cat("Elbow (slope-threshold):", elbow_slope, "\n")
cat("Elbow (kneedle/max-distance):", elbow_kneedle, "\n")
```

```{r}
# --- 3) Plot with both elbows marked ---
ggplot(df, aes(x = num_features, y = acc)) +
  geom_line(size = 1, color = "darkmagenta") +
  #geom_vline(xintercept = elbow_slope, linetype = 2) +
  geom_vline(xintercept = elbow_kneedle, linetype = 3) +
  # annotate("text", x = elbow_slope, y = max(df$acc, na.rm = TRUE),
  #          label = paste0("slope elbow = ", elbow_slope), vjust = -0.5, hjust = 0.9, angle = 90) +
  annotate("text", x = elbow_kneedle, y = max(df$acc, na.rm = TRUE),
           label = paste0("kneedle elbow = ", elbow_kneedle), vjust = -0.5, hjust = 1.5, angle = 90) +
  labs(x = "Number of Features", y = "Average Test Accuracy (%)") +
  theme_minimal()

```

```{r}
# ---- Optional: plot with elbow marker ----
ggplot(df, aes(num_features, acc)) +
  geom_line() +
  geom_point(size = 1.2, alpha = 0.7) +
  geom_vline(xintercept = opt_n_elbow, linetype = "dashed") +
  annotate("text", x = opt_n_elbow, y = max(df$acc, na.rm = TRUE),
           vjust = -0.5, label = paste("Elbow ≈", opt_n_elbow)) +
  labs(x = "Number of Features", y = "Validation Accuracy (%)") +
  theme_minimal()

```

plot with elbow marker:

```{r}
library(dplyr)
library(ggplot2)

# assume layer_18_log_reg already loaded
layer_18_log_reg <- layer_18_log_reg %>%
  mutate(
    num_features = as.integer(num_features),
    validation   = as.numeric(validation)
  ) %>%
  arrange(num_features)

opt_n_elbow <- 210
elbow_y <- layer_18_log_reg %>%
  filter(num_features == opt_n_elbow) %>%
  summarise(validation = mean(validation, na.rm = TRUE)) %>%
  pull(validation)

ggplot(layer_18_log_reg, aes(x = num_features, y = validation)) +
  geom_line(size = 1, color = "purple") +
  # elbow markers
  geom_vline(xintercept = opt_n_elbow, linetype = "dashed") +
  geom_point(aes(x = opt_n_elbow, y = elbow_y), size = 2.2) +
  annotate("text",
           x = opt_n_elbow, y = elbow_y,
           label = paste0("Elbow ≈ ", opt_n_elbow),
           vjust = -1, hjust = 1.1) +
  labs(
    x = "Number of Features",
    y = "Validation Accuracy (%)"
  ) +
  scale_y_continuous(limits = c(84, 100),
                     breaks = seq(84, 100, 2),
                     expand = expansion(mult = c(0.01, 0.03))) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5),
    panel.grid.minor = element_blank(),
    legend.position = "none"
  )

```

#==============================================================================================================
# SECRETOME ANALYSIS 
#==============================================================================================================


```{r}
model_320_logreg <- read.csv("/Volumes/T7 Shield/layer_18_F1_scores/secretome_log_reg_folds_320.csv")

colnames(model_320_logreg)
```

```{r}
model_320_logreg <- model_320_logreg %>%
  mutate(Entry = str_match(Entry, "^[^|]*\\|([^|]+)\\|")[,2]) %>%
  dplyr::rename("l18_logreg_prob" = probability)
```

```{r}
filtered_secretome <- read.csv("../../data/filtered_secretome_15316.csv")
```  
  
```{r}  
filtered_secretome_wl18pred <- filtered_secretome %>%
  inner_join(model_320_logreg) %>%
  select(Entry, neuropeptide, l18_logreg_prob, Protein_names, avg_prob, everything())
```
```{r}
low_prob_group <- filtered_secretome_wl18pred %>%
  select(Entry, neuropeptide, l18_logreg_prob, Protein_names) %>%
  filter(neuropeptide == TRUE) %>%
  filter(l18_logreg_prob <= 0.10)

#write.csv(low_prob_group, "../../data/lowest_group_l18_logreg.csv", row.names = FALSE)
```

```{r}
mid_prob_group <- filtered_secretome_wl18pred %>%
  select(Entry, neuropeptide, l18_logreg_prob, Protein_names) %>%
  filter(neuropeptide,                      # same as neuropeptide == TRUE
         l18_logreg_prob > 0.10,
         l18_logreg_prob <= 0.50)

#write.csv(mid_prob_group, "../../data/middle_group_l18_logreg.csv", row.names = FALSE)
```

```{r}
high_prob_FN_group <- filtered_secretome_wl18pred %>%
  select(Entry, neuropeptide, l18_logreg_prob, Protein_names) %>%
  filter(!neuropeptide,                      # same as neuropeptide == FALSE
         l18_logreg_prob >= 0.90) %>%
  arrange(desc(l18_logreg_prob))

#write.csv(high_prob_FN_group, "../../data/high_prob_non_neuro_group_l18_logreg.csv", row.names = FALSE)
```

Making a plot that counts the true positives in each ranking bin
```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(forcats)

# 1) Bin the probabilities into 10% buckets
binned <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(neuropeptide)) %>%
  mutate(
    prob_bin = cut(
      l18_logreg_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = TRUE,
      labels = paste0(seq(0, 90, 10), "%–", seq(10, 100, 10), "%")
    )
  ) %>%
  group_by(prob_bin) %>%
  summarise(
    total = n(),
    np_count = sum(neuropeptide == TRUE),
    prop_np = ifelse(total > 0, np_count / total, 0),
    .groups = "drop"
  ) %>%
  mutate(prob_bin = fct_explicit_na(prob_bin, na_level = "Unbinned"))

# 2) Plot: bar height = proportion neuropeptides; label "np/total" above bars
ggplot(binned, aes(x = fct_rev(prob_bin), y = prop_np)) +
  geom_col(fill = "lavender",
           color = "grey40",
           size = 0.3) +
  geom_text(aes(label = paste0(np_count, "/", total)),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 0.8)) +
  labs(
    x = "Predicted probability bin",
    y = "Neuropeptide proportion",
    title = "Proportion of neuropeptides by predicted probability bin"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

# here I want to filter the secretome entries to the allowed entries and rerun the plot above (I will be just removing the unused entries)

```{r}
colnames(filtered_secretome_wl18pred)
```
```{r}
library(tidyverse)

allowed_entries <- read_csv("../../data/final_human_secretome_entries_1201.csv")

filtered_secretome_wl18pred <- filtered_secretome_wl18pred %>%
  filter(Entry %in% allowed_entries$Entry)
```
```{r}
library(dplyr)
library(ggplot2)
library(scales)
library(forcats)

# 1) Bin the probabilities into 10% buckets
binned <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(neuropeptide)) %>%
  mutate(
    prob_bin = cut(
      l18_logreg_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = TRUE,
      labels = paste0(seq(0, 90, 10), "%–", seq(10, 100, 10), "%")
    )
  ) %>%
  group_by(prob_bin) %>%
  summarise(
    total = n(),
    np_count = sum(neuropeptide == TRUE),
    prop_np = ifelse(total > 0, np_count / total, 0),
    .groups = "drop"
  ) %>%
  mutate(prob_bin = fct_explicit_na(prob_bin, na_level = "Unbinned"))

# 2) Plot: bar height = proportion neuropeptides; label "np/total" above bars
ggplot(binned, aes(x = fct_rev(prob_bin), y = prop_np)) +
  geom_col(fill = "lavender",
           color = "grey40",
           size = 0.3) +
  geom_text(aes(label = paste0(np_count, "/", total)),
            vjust = -0.3, size = 3.5) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 0.8)) +
  labs(
    x = "Predicted probability bin",
    y = "Neuropeptide proportion",
    title = "Proportion of neuropeptides by predicted probability bin"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```





Plotting the distribution of peptide length within each bin.
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)

# 1) Add 10% probability bins to the row-level data
binned_rows <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(Length)) %>% 
  filter(neuropeptide == TRUE) %>%
  mutate(
    prob_bin = cut(
      l18_logreg_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = TRUE,
      labels = paste0(seq(0, 90, 10), "%–", seq(10, 100, 10), "%")
    ),
    prob_bin = fct_explicit_na(prob_bin, na_level = "Unbinned")
  )

# 2) Counts per bin for nicer x-axis labels
bin_counts <- binned_rows %>%
  count(prob_bin) %>%
  mutate(label = paste0(as.character(prob_bin), "\n(n=", scales::comma(n), ")"))

# Ensure label order matches factor order (we'll reverse in the plot)
lab_vec <- bin_counts$label
names(lab_vec) <- as.character(bin_counts$prob_bin)

# 3) Boxplot: Length distribution per bin (highest prob on the LEFT)
ggplot(binned_rows, aes(x = fct_rev(prob_bin), y = Length)) +
  geom_boxplot(outlier.alpha = 0.25, width = 0.6) +
  scale_x_discrete(labels = function(x) lab_vec[as.character(x)]) +
  labs(
    x = "Predicted probability bin",
    y = "Length",
    title = "Length distribution by predicted probability bin"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```

```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)

# Define the 10% bin labels once
bin_labels <- paste0(seq(0, 90, 10), "%–", seq(10, 100, 10))
first_bin <- bin_labels[1]                 # "0%–10%"
last_bin  <- bin_labels[length(bin_labels)]# "90%–100%"

# Row-level data with bins
binned_rows <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(Length)) %>%
  filter(neuropeptide) %>%
  mutate(
    prob_bin = cut(
      l18_logreg_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = TRUE,
      labels = bin_labels
    )
  )

# Keep only first & last bins; order so highest prob is on the LEFT
two_bins <- binned_rows %>%
  filter(prob_bin %in% c(first_bin, last_bin)) %>%
  mutate(prob_bin = fct_relevel(prob_bin, last_bin, first_bin))

# (Optional) add n per bin on x-axis labels
lab_df <- two_bins %>%
  count(prob_bin) %>%
  mutate(label = paste0(as.character(prob_bin), "\n(n=", comma(n), ")"))
lab_map <- setNames(lab_df$label, lab_df$prob_bin)

ggplot(two_bins, aes(x = prob_bin, y = Length)) +
  geom_boxplot(width = 0.6, 
               outlier.alpha = 0.25,
               outlier.shape = NA) +
  scale_x_discrete(labels = function(x) lab_map[x]) +
  scale_y_continuous(limits = c(0, 500)) +
  labs(
    x = "Predicted probability bin",
    y = "Length",
    title = "Length distribution in the lowest vs highest probability bins"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```
```{r}
library(dplyr)
library(ggplot2)
library(forcats)
library(scales)

# 1) Bin and compute mean Length + n per bin
bin_labels <- paste0(seq(0, 90, 10), "%–", seq(10, 100, 10))

binned_means <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(Length)) %>%
  mutate(
    prob_bin = cut(
      l18_logreg_prob,
      breaks = seq(0, 1, by = 0.1),
      include.lowest = TRUE,
      right = TRUE,
      labels = bin_labels
    )
  ) %>%
  group_by(prob_bin) %>%
  summarise(
    n = n(),
    avg_len = mean(Length, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    prob_bin = fct_rev(prob_bin),                         # highest on left
    label = sprintf("%.1f", avg_len)                      # text above bars
  )

# 2) Bar plot with mean above each bin
ggplot(binned_means, aes(x = prob_bin, y = avg_len)) +
  geom_col(fill = "lavender", color = "grey40", size = 0.3) +
  geom_text(aes(label = label), vjust = -0.25, size = 3.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Predicted probability bin",
    y = "Average Length",
    title = "Average sequence length by predicted probability bin"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```



zooom in on top 10%:
```{r}
library(dplyr)
library(ggplot2)
library(forcats)

topbin_1pct <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(neuropeptide)) %>%   # fix: remove fct_rev(), close parens
  filter(l18_logreg_prob >= 0.90) %>%                        # keep only 90–100%
  mutate(
    pct_bin = pmin(floor(l18_logreg_prob * 100), 100L),      # 90, 91, …, 100
    bin_lab = factor(paste0(pct_bin, "%"),
                     levels = paste0(90:100, "%"))
  ) %>%
  group_by(bin_lab) %>%
  summarise(
    total    = n(),
    np_count = sum(neuropeptide == TRUE),
    .groups  = "drop"
  )

# If all bins are 0, avoid ylim being 0
ymax <- max(topbin_1pct$np_count, na.rm = TRUE)
ymax <- ifelse(is.finite(ymax) && ymax > 0, ymax * 1.1, 1)

topbin_1pct <- topbin_1pct %>%
  mutate(bin_lab = forcats::fct_rev(bin_lab))   # highest (100%) on the left

ggplot(topbin_1pct, aes(x = bin_lab, y = np_count)) +
  geom_col(fill = "lavender", color = "grey40", size = 0.3) +
  geom_text(aes(label = paste0(np_count, "/", total)),
            vjust = -0.3, size = 3.5) +
  labs(
    x = "Predicted probability (1% bins within 90–100%)",
    y = "True positives (count)",
    title = "True positives by predicted probability (90–100%)"
  ) +
  coord_cartesian(ylim = c(0, ymax)) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

```
```
  
  
  
confusion matrix for model:
  
```{r}
library(dplyr)
library(ggplot2)
library(scales)

# --- choose a threshold (edit me) ---
thr <- 0.5

cm <- filtered_secretome_wl18pred %>%
  filter(!is.na(l18_logreg_prob), !is.na(neuropeptide)) %>%
  mutate(
    pred = l18_logreg_prob >= thr,
    # confusion bucket
    bucket = case_when(
      pred &  neuropeptide ~ "TP",
      pred & !neuropeptide ~ "FP",
      !pred &  neuropeptide ~ "FN",
      TRUE                 ~ "TN"
    )
  ) %>%
  count(bucket, name = "n") %>%
  tidyr::complete(bucket = c("TP","FP","FN","TN"), fill = list(n = 0)) %>%
  mutate(pct = n / sum(n))

# layout coordinates for 2×2 grid (Pred on x, True on y)
layout <- tibble::tribble(
  ~bucket, ~x, ~y, ~pred_lbl, ~true_lbl,
  "TP",    1,  1,  "Pred: +", "True: +",
  "FP",    2,  1,  "Pred: +", "True: −",
  "FN",    1,  2,  "Pred: −", "True: +",
  "TN",    2,  2,  "Pred: −", "True: −"
)

cm_plot <- cm %>%
  inner_join(layout, by = "bucket") %>%
  mutate(label = paste0(bucket, "\n", n, " (", percent(pct, accuracy = 0.1), ")"))

ggplot(cm_plot, aes(x = x, y = y, fill = bucket)) +
  geom_tile(color = "white", width = 0.98, height = 0.98) +
  geom_text(aes(label = label), size = 5, fontface = "bold", color = "white") +
  scale_fill_manual(values = c(TP = "#52856c", TN = "#447a9e", FP = "#943e35", FN = "#b58b2a")) +
  scale_x_continuous(breaks = c(1,2), labels = c("Predicted +", "Predicted −"), expand = c(0,0)) +
  scale_y_reverse(breaks = c(1,2), labels = c("True +", "True −"), expand = c(0,0)) +
  coord_fixed() +
  labs(
    title = paste0("Confusion Matrix at threshold = ", thr, sprintf("\nTotal N = %s", scales::comma(sum(cm$n)))),
    x = NULL, y = NULL, fill = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold"),
    plot.title  = element_text(hjust = 0.5, face = "bold"),
    legend.position = "none"
  )

```
Distribution of sequences by length in the 32,000 neuropeptides:  
```{r}

length_dist_32k <- read.csv("../../data/postive_graphpart_data_binned.csv") 
```
  
```{r}
plot_df <- length_dist_32k %>%
  mutate(
    bin_start = as.integer(sub("-.*", "", bin)),
    bin = factor(bin, levels = bin[order(bin_start)])
  )

ggplot(plot_df, aes(x = bin, y = count)) +
  geom_col(fill = "#326ba8") +
  geom_text(aes(label = count), vjust = -0.3, size = 3) +
  labs(
    x = "Length bin (aa)",
    y = "Count",
    title = "Sequence length distribution (50-aa bins)"
  ) +
  expand_limits(y = max(plot_df$count) * 1.1) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) 
```




  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  
  

