---
title: "layer_18_datasets"
output: html_document
date: "2025-11-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(tidytext)    # for reorder_within() and scale_x_reordered()
```

```{r}
novo_all_data_smORFs <- read.csv("/Volumes/T7 Shield/layer_18_F1_scores/novo_smORF_50000_log_reg_folds_320.csv") 
```

For a sanity check, let's count the number of datapoints where the probability is >90%
```{r}
count_high_prob <- novo_all_data_smORFs %>%
  select(probability) %>%
  filter(probability > 0.9) %>%
  count()

count_high_prob
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)

data_dir <- "/Volumes/T7 Shield/layer_18_F1_scores/"

# Helper to read one file
read_scores <- function(fname) {
  read_csv(file.path(data_dir, fname))
}

novo_150      <- read_scores("novo_150_log_reg_folds_320.csv")
novo_smORF_50k <- read_scores("novo_smORF_50000_log_reg_folds_320.csv")
de_Souza      <- read_scores("de_Souza_log_reg_folds_320.csv")
Whited        <- read_scores("Whited_log_reg_folds_320.csv")
```

```{r}
count_high_prob <- de_Souza %>%
  select(probability) %>%
  filter(probability > 0.9) %>%
  count()

count_high_prob
```


```{r}
# Common plotting function
make_prob_plot <- function(df, title) {
  
  # Labels that *match* the breaks: 0–10, 10–20, ..., 90–100
  bin_labels <- paste0(seq(0, 90, by = 10), "%–", seq(10, 100, by = 10), "%")
  
  df %>%
    mutate(
      probability = as.numeric(probability),
      prob_bin = cut(
        probability,
        breaks = seq(0, 1, by = 0.1),
        include.lowest = TRUE,
        right = TRUE,
        labels = bin_labels
      )
    ) %>%
    count(prob_bin, name = "count") %>%
    # Reorder so that 90–100 is on the left, 0–10 on the right
    mutate(prob_bin = factor(prob_bin, levels = rev(bin_labels))) %>%
    ggplot(aes(x = prob_bin, y = count)) +
    geom_col(width = 0.9, fill = "#9bafe0", color = "gray57") +
    geom_text(aes(label = count), vjust = -0.3, size = 3.5) +
    scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
    labs(
      title = title,
      x = "Probability bin (10%)",
      y = "Count"
    ) +
    theme_minimal(base_size = 13) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.minor = element_blank()
    )
}

# 4 separate plots (each will render separately in the Rmd)
plot_novo_150       <- make_prob_plot(novo_150,       "novo_150_log_reg_folds_320")
plot_novo_smORF_50k <- make_prob_plot(novo_smORF_50k, "novo_smORF_50000_log_reg_folds_320")
plot_de_Souza       <- make_prob_plot(de_Souza,       "de_Souza_log_reg_folds_320")
plot_Whited         <- make_prob_plot(Whited,         "Whited_log_reg_folds_320")
```

```{r}
plot_novo_150
```

```{r}
plot_novo_smORF_50k
```
I have removed enzymes, transmembrane regions, known neuropeptides and filtered for length <= 500 residues. 
Here is the new plot. 
```{r}
library(dplyr)
library(stringr)

fasta_path <- "../../data/novo_alldata_l18_no_TMRs.fasta"

# 1) Read FASTA headers
fa_lines <- readLines(fasta_path)
hdrs <- fa_lines[grepl("^>", fa_lines)]

# Strip leading ">" and keep only the first whitespace-separated token
id_tokens <- hdrs %>%
  sub("^>", "", .) %>%
  sub(" .*", "", .)

# 2) Extract UniProt-style accession from tokens
#    If format is sp|ACC|NAME or tr|ACC|NAME → take ACC (middle part)
#    Otherwise use the token itself
entries_from_fasta <- ifelse(
  str_detect(id_tokens, "\\|"),
  vapply(strsplit(id_tokens, "\\|"), `[`, character(1), 2),
  id_tokens
)

# Optionally strip isoform suffix: P12345-2 -> P12345
entries_from_fasta <- sub("-[0-9]+$", "", entries_from_fasta)

entries_from_fasta <- unique(entries_from_fasta)

# 3) Filter your dataframe
novo_smORF_50k_filtered <- novo_smORF_50k %>%
  filter(Entry %in% entries_from_fasta)

# (Optional) overwrite original:
# novo_smORF_50k <- novo_smORF_50k_filtered

# Quick sanity check: how many kept vs removed
n_before <- nrow(novo_smORF_50k)
n_after  <- nrow(novo_smORF_50k_filtered)
n_removed <- n_before - n_after

cat("Rows before:", n_before, "\n")
cat("Rows after: ", n_after,  "\n")
cat("Rows removed:", n_removed, "\n")

```

Now making the new plot:
```{r}
plot_novo_smORF_50k_filtered <- make_prob_plot(novo_smORF_50k_filtered, "novo_smORF_50000_log_reg_folds_320")
plot_novo_smORF_50k_filtered
```
Saving the 277 top hits:
```{r}
library(dplyr)

# Keep entries with probability >= 90%
novo_smORF_50k_highprob <- novo_smORF_50k_filtered %>%
  filter(probability > 0.6 & probability <= 0.7)

# Save to CSV (adjust path if you like)
highprob_csv <- "../../data/novo_alldata_l18_no_TMRs_60_to_70.csv"
write.csv(novo_smORF_50k_highprob, highprob_csv, row.names = FALSE)

nrow(novo_smORF_50k_filtered)
nrow(novo_smORF_50k_highprob)

```

Saving the 8 Whited top hits:
```{r}
library(dplyr)

# Keep entries with probability >= 90%
Whited_highprob <- Whited %>%
  filter(probability >= 0.7)

# Save to CSV (adjust path if you like)
highprob_csv <- "../../data/Whited_l18_highprob.csv"
write.csv(Whited_highprob, highprob_csv, row.names = FALSE)

nrow(Whited)
nrow(Whited_highprob)

```

Saving the 13 de_Souza top hits:
```{r}
library(dplyr)

# Keep entries with probability >= 90%
de_Souza_highprob <- de_Souza %>%
  filter(probability >= 0.7)

# Save to CSV (adjust path if you like)
highprob_csv <- "../../data/de_Souza_l18_highprob.csv"
write.csv(de_Souza_highprob, highprob_csv, row.names = FALSE)

nrow(de_Souza)
nrow(de_Souza_highprob)

```


```{r}
plot_de_Souza
```





```{r}
plot_Whited
```

#======================== plotting the distribution between 100 and 90% probability ==================================================

```{r}
# 1% bins within the 90–100% range for novo_smORF_50k
novo_smORF_50k_90_100 <- novo_smORF_50k %>%
  mutate(
    probability = as.numeric(probability),
    probability = pmin(pmax(probability, 0), 1)  # clamp to [0,1]
  ) %>%
  filter(!is.na(probability)) %>%
  mutate(
    # integer percentage 0–100
    pct = floor((probability + 1e-8) * 100),
    # map exact 100% into 99–100 bin
    pct = pmin(pct, 99L)
  ) %>%
  # keep only 90–99 (i.e., 90–91, ..., 99–100)
  filter(pct >= 90, pct <= 99) %>%
  count(pct, name = "count") %>%
  # ensure all bins 90–99 exist
  tidyr::complete(pct = 90:99, fill = list(count = 0)) %>%
  mutate(
    bin_label = paste0(pct, "%–", pct + 1, "%"),
    # order so 99–100 on the left, 90–91 on the right
    bin_label = factor(bin_label,
                       levels = paste0(99:90, "%–", 100:91, "%"))
  )

ggplot(novo_smORF_50k_90_100, aes(x = bin_label, y = count)) +
  geom_col(width = 0.9) +
  geom_text(aes(label = count), vjust = -0.3, size = 3.5) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "novo_smORF_50000_log_reg_folds_320 (90–100%, 1% bins)",
    x = "Probability bin (1%)",
    y = "Count"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.minor = element_blank()
  )
```

#============================================================================================================  
#Here I want to correlate the layer 18 predictions with the average predictions of the cleavage site model
#============================================================================================================
```{r}
library(tidyverse)

clean_entry <- function(x) {
  x <- as.character(x)
  ifelse(
    grepl("\\|", x),
    vapply(strsplit(x, "\\|"), `[`, character(1), 2),  # sp|ACC|NAME → ACC
    x
  )
}

```



1. Compute per-Entry max cleavage probabilities for each model
```{r}
max_logreg <- secretome_logreg_predicitons %>%
  mutate(Entry = clean_entry(Entry)) %>%
  group_by(Entry) %>%
  summarise(max_logreg = max(probability, na.rm = TRUE), .groups = "drop")

max_crf <- secretome_crf_predicitons %>%
  mutate(Entry = clean_entry(Entry)) %>%
  group_by(Entry) %>%
  summarise(max_crf = max(probability, na.rm = TRUE), .groups = "drop")

max_rf <- secretome_ranfor_predicitons %>%
  mutate(Entry = clean_entry(Entry)) %>%
  group_by(Entry) %>%
  summarise(max_rf = max(probability, na.rm = TRUE), .groups = "drop")

max_svm <- secretome_svm_predicitons %>%
  mutate(Entry = clean_entry(Entry)) %>%
  group_by(Entry) %>%
  summarise(max_svm = max(probability, na.rm = TRUE), .groups = "drop")

max_cleavage_df <- max_logreg %>%
  full_join(max_crf, by = "Entry") %>%
  full_join(max_rf,  by = "Entry") %>%
  full_join(max_svm, by = "Entry") %>%
  mutate(
    mean_max_cleav = rowMeans(
      dplyr::select(., max_logreg, max_crf, max_rf, max_svm),
      na.rm = TRUE
    )
  )


```

2. Prep layer-18 per-Entry probabilities
```{r}
model_320_logreg <- read.csv(
  "/Volumes/T7 Shield/layer_18_F1_scores/secretome_log_reg_folds_320.csv"
)

layer18_df <- model_320_logreg %>%
  mutate(Entry = clean_entry(Entry)) %>%
  group_by(Entry) %>%
  summarise(
    prob_l18 = mean(probability, na.rm = TRUE),
    .groups  = "drop"
  )


```

3. Join with neuropeptide flags
```{r}
secr_neuro_flags_raw <- read.csv("../../data/secretome_neuropeptide_flags.csv")

flags_df <- secr_neuro_flags_raw %>%
  mutate(
    Entry = clean_entry(Entry),
    neuro_flag = case_when(
      is.logical(neuropeptide) ~ neuropeptide,
      tolower(as.character(neuropeptide)) %in% c("true", "t", "1", "yes") ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  select(Entry, neuro_flag)

table(flags_df$neuro_flag, useNA = "ifany")
# should show FALSE and TRUE, like your earlier output

```
Checking overlap:
```{r}
neuro_ids <- flags_df$Entry[flags_df$neuro_flag]

# Overlap with layer-18
length(intersect(neuro_ids, layer18_df$Entry))

# Overlap with cleavage ensemble
length(intersect(neuro_ids, max_cleavage_df$Entry))
```



```{r}

table(secr_neuro_flags$neuropeptide, useNA = "ifany")
head(secr_neuro_flags$Entry[secr_neuro_flags$neuropeptide == TRUE], 10)
```

```{r}
# 0) Load filter tables
secretome_no_enzymes <- read.csv(
  "../../data/secretome_filter_out_enzymes_5/secretome_filtered_nonenzymes_nocsdhit.csv"
)

secr_no_tm <- read.csv(
  "../../data/secretome_filter_out_nocdhit/uniprot_tm_filter/secretome_nontransmembrane_entries.csv"
)

# allowed entries = non-enzyme ∩ non-TM (after cleaning IDs)
allowed_entries <- intersect(
  clean_entry(secretome_no_enzymes$Entry),
  clean_entry(secr_no_tm$Entry)
)


```

```{r}
plot_df <- layer18_df %>%
  inner_join(max_cleavage_df, by = "Entry") %>%
  left_join(flags_df, by = "Entry") %>%
  filter(Entry %in% allowed_entries)   # <-- small fix


valid_pairs <- plot_df %>%
  filter(!is.na(prob_l18), !is.na(mean_max_cleav))

table(valid_pairs$neuro_flag, useNA = "ifany")
```


4. Correlations
```{r}
cor_pearson  <- cor(plot_df$prob_l18, plot_df$mean_max_cleav,
                    use = "complete.obs", method = "pearson")
cor_spearman <- cor(plot_df$prob_l18, plot_df$mean_max_cleav,
                    use = "complete.obs", method = "spearman")

cor_pearson
cor_spearman

```

5. 2D density “3D histogram” + neuropeptide points (styled like your example)
```{r}
library(cowplot)

orf_data   <- valid_pairs
neuro_data <- valid_pairs %>% dplyr::filter(neuro_flag)   # <- THIS is the key change

#nrow(neuro_data)   # should be 276

sizes <- c(background = 1.2, neuropeptide = 2.3)

point_plot_secr <- ggplot() +
  # background points
  geom_point(
    data  = orf_data,
    aes(x = mean_max_cleav, y = prob_l18, size = "background"),
    color = "#CAD8EF",
    stroke = 0.2
  ) +
  # density polygons
  stat_density_2d(
    geom    = "polygon",
    contour = TRUE,
    data    = orf_data,
    aes(x = mean_max_cleav, y = prob_l18,
        fill = factor(after_stat(level))),
    bins      = 10,
    key_glyph = "blank"
  ) +
  # highlighted neuropeptide points
  geom_point(
    data  = neuro_data,
    aes(x = mean_max_cleav, y = prob_l18, size = "neuropeptide"),
    shape  = 21,
    color  = "#70685d",
    fill   = "#ffb22e",
    stroke = 0.25
  ) +
  scale_fill_manual(
    values = c(
      '#BACAE7', '#ACBCD8', '#A2B1CD', '#93A1BC', '#8494AF',
      '#7787A1', '#6A7B95', '#60708A', '#52647F', '#3C597E'
    ),
    guide = "none"
  ) +
  scale_size_manual(
    values = sizes,
    breaks = c("background", "neuropeptide"),
    labels = c("Secretome", "Neuropeptide"),
    drop   = FALSE
  ) +
  guides(
    size = guide_legend(
      override.aes = list(
        fill   = "#CAD8EF",
        color  = "#8aa4ba",
        shape  = 16,
        stroke = 0.2
      )
    )
  ) +
  scale_x_continuous(
    name   = "Mean max cleavage-site probability (4-model ensemble)",
    limits = c(0, 1)
  ) +
  scale_y_continuous(
    name   = "Layer-18 logreg probability",
    limits = c(0, 1)
  ) +
  labs(size = "") +
  theme_cowplot(16) +
  theme(
    legend.position = "top"
  )

point_plot_secr

```























