---
title: "logistic_reg_analysis.Rmd"
output: html_document
date: "2025-07-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(tidytext)    # for reorder_within() and scale_x_reordered()
```


#=========================================================================================
# VALIDATION RESULTS
#=========================================================================================

```{r}
log_reg_40 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/linear_regression_results_40_hom_dup.csv")
log_reg_50 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/linear_regression_results_50_hom_dup.csv")
log_reg_60 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/linear_regression_results_60_hom_dup.csv")
log_reg_70 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/linear_regression_results_70_hom_dup.csv")
log_reg_80 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/linear_regression_results_80_hom_dup.csv")
log_reg_90 <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/linear_regression_results_90_hom_dup.csv")
```

combining data into a single df
```{r}

combined <- bind_rows(
  `40` = log_reg_40,
  `50` = log_reg_50,
  `60` = log_reg_60,
  `70` = log_reg_70,
  `80` = log_reg_80,
  `90` = log_reg_90,
  .id  = "perc_homology"
) %>%
  mutate(
    # make sure perc_homology is numeric, not a string
    perc_homology = as.numeric(perc_homology)
  )
```

Now I want a line plot of validation accuracy across number of features. Each percent homology
will be a line of a different color. 
```{r}

ggplot(combined, aes(x = num_features, y = validation, color = factor(perc_homology))) +
  geom_line(size = 1) +
  scale_color_manual(
    name   = "training/validation\nset homology (%)",
    values = qualitative_hcl(
      n       = length(unique(combined$perc_homology)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Validation Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```
Now, let's finish building out first test set (human neuropeptides) and then run this logistic regression on our test set and get the result as 
single percent accuracy. 

#=========================================================================================
# TEST RESULTS
#=========================================================================================

Test set done. Let's plot the test results.

```{r}
log_reg_40_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_40_hom_dup.csv")
log_reg_50_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_50_hom_dup.csv")
log_reg_60_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_60_hom_dup.csv")
log_reg_70_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_70_hom_dup.csv")
log_reg_80_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_80_hom_dup.csv")
log_reg_90_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_90_hom_dup.csv")

```

```{r}
combined_test <- bind_rows(
  `40` = log_reg_40_test,
  `50` = log_reg_50_test,
  `60` = log_reg_60_test,
  `70` = log_reg_70_test,
  `80` = log_reg_80_test,
  `90` = log_reg_90_test,
  .id  = "perc_homology"
) %>%
  mutate(
    # make sure perc_homology is numeric, not a string
    perc_homology = as.numeric(perc_homology)
  )
```

```{r}
ggplot(combined_test, aes(x = num_features, y = test_accuracy, color = factor(perc_homology))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "training/validation\nset homology (%)",
    values = qualitative_hcl(
      n       = length(unique(combined$perc_homology)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Test Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```

# Logistic regression using folds from graphpart. Reminder there were 10 total folds at 50% identity. 
# fold 9 was set asside as the test dataset
# folds 0-8 were used in training and cross validation

```{r}
log_reg_folds <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/log_regression_csv_results/logistic_regression_results_folds_0-8_40.csv")
```

```{r}
log_reg_folds_test <- read.csv("/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/linear_regression_test_results_fold_9_40.csv")
```

```{r}
val_data <- log_reg_folds %>%
  rename(accuracy = validation) %>%
  mutate(dataset = "validation")

test_data <- log_reg_folds_test %>%
  rename(accuracy = test_accuracy) %>%
  mutate(dataset = "test")

combined_val_test_folds <- bind_rows(val_data, test_data)

```

```{r}
ggplot(
  data = combined_val_test_folds,
  aes(x = num_features, y = accuracy, color = dataset)
  ) +
  scale_color_manual(
    name   = "dataset",
    values = c("darkmagenta", "darkgreen")
  ) +
  geom_line(size = 0.8) +
  labs(
    x = "Number of Features",
    y = "Validation Accuracy (%)"
  ) +
  theme_minimal()


```
Now let's plot the test results:



```{r}
ggplot(
  data = log_reg_folds_test,
  aes(x = num_features, y = test_accuracy)
) +
  geom_line(size = 0.8, color = "darkmagenta") +
  labs(
    x = "Number of Features",
    y = "Validation Accuracy (%)"
  ) +
  theme_minimal()


```


Loading all the test results from the 40% nested k-fold cross validation
```{r}
# install.packages("tidyverse")  # if you don't already have it
library(tidyverse)

# 1) Point to the directory where your fold sub-folders live:
base_dir <- "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results/test_log_regression_csv_results/graphpart_40"

# 2) Find all the CSV files (one per fold):
files <- list.files(
  path       = base_dir, 
  pattern    = "^linear_regression_test_results_test_fold_\\d+\\.csv$", 
  recursive  = TRUE,
  full.names = TRUE
)

# 3) Read & combine, extracting the fold number from each filename:
all_results <- files %>% 
  set_names() %>% 
  map_dfr(read_csv, .id = "file", show_col_types = FALSE) %>% 
  mutate(
    fold = str_extract(basename(file), "\\d+")
  ) %>% 
  select(fold, num_features, test_accuracy)
```

Ploting the graph:
```{r}
ggplot(all_results, aes(x = num_features, y = test_accuracy, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Test Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )


```

Getting df with the average test accuracy:
```{r}
avg_results <- all_results %>%
  # if test_accuracy is character, first turn it numeric:
  mutate(test_accuracy = as.numeric(test_accuracy)) %>%
  group_by(num_features) %>%
  summarise(
    avg_test_accuracy = mean(test_accuracy, na.rm = TRUE),
    .groups = "drop"
  )
```

plotting the average test result:
```{r}
ggplot(
  data = avg_results,
  aes(x = num_features, y = avg_test_accuracy)
  ) +
  geom_line(size = 0.8, color = "darkmagenta") +
  labs(
    x = "Number of Features",
    y = "Average Test Accuracy (%)"
  ) +
  theme_minimal()
```
Detecting the final number of features I need to use:
(Elbow/knee method)
```{r}
# all_results should have: fold, num_features, test_accuracy
# Make sure accuracy is numeric (not character):
all_results <- all_results %>%
  mutate(test_accuracy = as.numeric(test_accuracy))

# 1) Summarise across folds → mean, sd, se per num_features
avg_results <- all_results %>%
  group_by(num_features) %>%
  summarise(
    avg_test_accuracy = mean(test_accuracy, na.rm = TRUE),
    sd_acc            = sd(test_accuracy,   na.rm = TRUE),
    n_folds           = dplyr::n(),
    se_acc            = sd_acc / sqrt(n_folds),
    .groups = "drop"
  ) %>%
  arrange(num_features)

```

```{r}
# ---- Method A: Elbow/knee on slope ----
slopes <- diff(avg_results$avg_test_accuracy) / diff(avg_results$num_features)
threshold <- 0.001  # e.g., 0.1% gain per 10 features (adjust as you wish)

idx <- which(slopes < threshold)
opt_n_elbow <- if (length(idx)) {
  # take the first time the marginal gain drops below threshold
  avg_results$num_features[idx[1] + 0]
} else {
  # fallback: never dropped → pick the largest tested (or choose your own rule)
  max(avg_results$num_features)
}

opt_n_elbow
```
# [1] 70  (in your previous run)

```{r}
# ---- Method B: 1-SE rule (pick simplest within 1 SE of best) ----
best_row <- avg_results %>% slice_max(avg_test_accuracy, n = 1, with_ties = FALSE)
best_acc <- best_row$avg_test_accuracy
best_se  <- best_row$se_acc

opt_n_1se <- avg_results %>%
  filter(avg_test_accuracy >= best_acc - best_se) %>%
  slice_min(num_features) %>%
  pull(num_features)

opt_n_1se
```




Loading validation results
```{r}
percent_identity <- 40
base_dir <- file.path(
  "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/classifier_results",
  "log_regression_csv_results",
  paste0("graphpart_", percent_identity)
)

# 1) Locate all per‐fold validation CSVs
files <- list.files(
  path       = base_dir,
  pattern    = "logistic_regression_results_excluded_fold_\\d+\\.csv$",
  recursive  = TRUE,
  full.names = TRUE
)

# 2) Read & combine
all_val_results <- files %>%
  set_names() %>% 
  map_dfr(~ {
    df <- read_csv(.x, show_col_types = FALSE)
    fold_num <- str_extract(basename(.x), "\\d+") %>% as.integer()
    df %>% mutate(fold = fold_num)
  })

```

plotting the validation results:
```{r}
ggplot(all_val_results, aes(x = num_features, y = validation, color = factor(fold))) +
  geom_line(size = 0.8) +
  scale_color_manual(
    name   = "Test fold",
    values = qualitative_hcl(
      n       = length(unique(all_results$fold)),
      palette = "Set 2"
    )
  ) +
  labs(
    x = "Number of Features",
    y = "Average Validation Accuracy (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title      = element_text(hjust = 0.5),
    legend.position = "right"
  )

```


























