---
title: "secretome_analysis_2"
output: html_document
date: "2026-01-21"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggiraph)
library(glue)
library(htmltools)
library(fs)
library(readr)
library(stringr)
```

Loading the secratome annotations:
```{r}
secratome_annotations <- read.delim(file = "../../data/secratome_annotations.tsv", sep="\t", na = c("", "NA", "#N/A"))
```

Let's clean the secratome annotations and then join the log reg predictions. 

```{r}
final_secratome_df <- secratome_annotations %>%
  dplyr::select(
    Entry,
    Sequence,
    Length,
    Signal.peptide,
    Peptide,
    Protein.names,
    Function..CC.,
    Gene.Ontology..molecular.function.
  ) %>%
  dplyr::rename(
    Signal_peptide   = Signal.peptide,
    Protein_names    = Protein.names,
    Protein_function = Function..CC.,
    Go_function      = Gene.Ontology..molecular.function.
  ) %>%
  dplyr::select(
    Entry,
    Protein_names,
    Protein_function,
    Go_function,
    Signal_peptide,
    Peptide,
    Sequence,
    Length
  )
```

```{r}
colnames(secratome_annotations)
```

Now I want to add an additional column ("Neuropeptide") 
that will flag the secreted proteins as neuropeptide/hormone peptides. 

```{r}
secr_neuro_flags <- read.csv(file = "../../data/secratome_neuropeptide_flags.csv")
```

Now let's join the two df together.
```{r}
final_secratome_df <- final_secratome_df %>%
  inner_join(secr_neuro_flags) %>%
  select(Entry, Protein_names, neuropeptide, everything())
```

I now need to filter the final secratome_df to the allowed entries:

```{r}
#allowed_entries <- read.csv(file = "../../data/deduplicated_secretome_entries.csv")
allowed_entries <- read.csv(file = "../../data/signalP_cdhit_secretome_entries.csv")

final_secratome_df <- final_secratome_df %>%
  inner_join(allowed_entries, by = "Entry")

```

sanity check
```{r}
nrow(final_secratome_df)
anyDuplicated(final_secratome_df$Entry)
```
```{r}
final_secratome_df %>%
  count(neuropeptide)
```

```{r}
write.csv(
  final_secratome_df,
  file = "../../data/deduplicated_secretome_annotations_signalp_cdhit.csv",
  row.names = FALSE
)
```

Now, I am going to parse Helena and Thomas' neuropeptide re-annotations

```{r}
library(dplyr)

# Load data
hs_df  <- read.csv("../../data/deduplicated_secretome_annotations_HS.csv")
tlk_df <- read.csv("../../data/deduplicated_secretome_annotations_TLK.csv")

# Join: keep HS length, add Thomas from TLK
secretome_HS_TLK <- hs_df %>%
  left_join(
    tlk_df %>% select(Entry, Thomas),
    by = "Entry"
  ) %>%
  select(
    Entry,
    neuropeptide,
    Helena,
    Thomas,
    everything()
  )


```

cleaning up the Helena and Thomas columns to booleans 

```{r}
secretome_HS_TLK <- secretome_HS_TLK %>%
  mutate(
    Helena = Helena == 2,
    Thomas = Thomas == 2
  )
```

Now I want the rows where HELENA != THOMAS (where they do not agree)

```{r}
secretome_no_agreement <- secretome_HS_TLK %>%
  filter(Helena != Thomas)

```

saving the no agreement as a csv:
```{r}
write.csv(
  secretome_no_agreement,
  "../../data/secretome_no_agreement.csv",
  row.names = FALSE
)

```

Just to know, how many peptides did Helena reannotate? See the code block below

```{r}
secretome_HS_TLK %>%
  filter(!is.na(Helena)) %>%
  summarise(
    true_to_false = sum(neuropeptide == TRUE  & Helena == FALSE),
    false_to_true = sum(neuropeptide == FALSE & Helena == TRUE)
  )
```

```{r}
secretome_HS_TLK %>%
  filter(!is.na(Thomas)) %>%
  summarise(
    true_to_false = sum(neuropeptide == TRUE  & Thomas == FALSE),
    false_to_true = sum(neuropeptide == FALSE & Thomas == TRUE)
  )
```

```{r}
library(dplyr)

reannotation_summary <- secretome_HS_TLK %>%
  summarise(
    Annotator = c("Helena", "Thomas"),

    Total = c(
      sum(!is.na(Helena)),
      sum(!is.na(Thomas))
    ),

    TRUE_to_FALSE = c(
      sum(neuropeptide == TRUE  & Helena == FALSE, na.rm = TRUE),
      sum(neuropeptide == TRUE  & Thomas == FALSE, na.rm = TRUE)
    ),

    FALSE_to_TRUE = c(
      sum(neuropeptide == FALSE & Helena == TRUE, na.rm = TRUE),
      sum(neuropeptide == FALSE & Thomas == TRUE, na.rm = TRUE)
    )
  )

print(reannotation_summary)

```

Final secretome!!
```{r}
final_neuroflags <- read.csv(file = "../../data/secretome_revised_TLK.csv")

final_neuroflags %>%
  count(Thomas.revised)

```

This is just the 33 they disagreed on.

The others should remain the same.
```{r}
revised_neuroflags <- final_neuroflags %>%
  dplyr::select(Entry, Thomas.revised) %>%
  dplyr::rename(neuropeptide = Thomas.revised) %>%
  dplyr::mutate(
    neuropeptide = as.logical(neuropeptide)
  )
```


```{r}
secretome_perfect_agreement <- secretome_HS_TLK %>%
  dplyr::filter(Helena == Thomas) %>%
  dplyr::select(Entry, Thomas) %>%
  dplyr::rename(neuropeptide = Thomas) %>%
  dplyr::mutate(neuropeptide = as.logical(neuropeptide))

neuroflags_revised_final <- rbind(revised_neuroflags, secretome_perfect_agreement)
```

```{r}
write.csv(
  neuroflags_revised_final, "../../data/secretome_neuroflags_revised_final.csv",
  row.names = FALSE
)

```

```{r}
neuroflags_revised_final %>%
  count(neuropeptide)
```


