---
title: "toxin_test_set"
output: html_document
date: "2025-07-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggforce)
library(ggridges)
options(scipen = 999)
library(colorspace)
library(tidytext)    # for reorder_within() and scale_x_reordered()
```


```{r}
toxin_annotations <- read.delim(file = "../../data/test_dataset_stats/toxins_KW-0800.tsv", sep="\t", na = c("", "NA", "#N/A"))
```

Let's filter out the NAs and see how many toxins we have that are annotated with "peptide"
```{r}
clean_toxins <- toxin_annotations %>%
  select(Entry, Sequence, Length, Peptide, Signal.peptide) %>%
  filter(!is.na(Peptide))
```

Now, let's save the clean_toxins Entries as a fasta file so that we can align them to the
303 neuropeptides we used 

```{r}
for_fasta <- clean_toxins %>%
  select(Entry, Sequence)

# Create FASTA-formatted lines: ">ID" on one line, followed by the sequence on the next
fasta_lines <- apply(for_fasta, 1, function(row) {
  paste0(">", row["Entry"], "\n", row["Sequence"])
})

# Write these lines to a .fasta file
writeLines(fasta_lines, con = "../../data/test_dataset_stats/clean_toxins.fasta")
```

Pairwise alignments were made between the toxin sequences and the neuropeptides. 

Here is the distribution of percent similarity between the two datasets:
```{r}
align_toxins <- read.csv(file = "../../data/test_dataset_stats/toxin_pairwise_alignment_results.csv")
```

Let's plot the distribution of percent similarity:
```{r}
align_toxins %>%
  ggplot(aes(x = percent_seq_sim)) +
  geom_histogram(
    binwidth = 5,
    center   = 2.5,
    color    = "gray60",
    fill     = "cornsilk"
  ) +
  # add counts above each bar
  stat_bin(
    binwidth = 5,
    center   = 2.5,
    geom     = "text",
    aes(label = ..count..),
    vjust    = -0.5
  ) +
  xlab(label = "percent sequence similarity") +
  # give a little headroom so the labels don't get cut off
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme_minimal()

```

```{r}
high_sim_toxins <- align_toxins %>%
  filter(percent_seq_sim > 30)
```

Removing the toxin entry the is exactly the same as an entry in the neuropeptide dataset:
```{r}
clean_align_toxins <- align_toxins %>%
  filter(Entry_test != "P0DP56")
```


```{r}
clean_align_toxins %>%
  ggplot(aes(x = percent_seq_sim)) +
  geom_histogram(
    binwidth = 5,
    center   = 2.5,
    color    = "gray60",
    fill     = "cornsilk"
  ) +
  # add counts above each bar
  stat_bin(
    binwidth = 5,
    center   = 2.5,
    geom     = "text",
    aes(label = ..count..),
    vjust    = -0.5
  ) +
  xlab(label = "percent sequence similarity") +
  # give a little headroom so the labels don't get cut off
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  theme_minimal()

```

Let's take the clean toxins dataset and see how many Entries are left after we filter out the mature peptides.

```{r}
clean_toxins_2 <- clean_toxins %>%
  # extract all “123..456” substrings as a list-column
  mutate(range_str = str_extract_all(Peptide, "\\d+\\.\\.\\d+")) %>%
  # turn each element of that list into its own row
  unnest_longer(range_str) %>%
  # split “123..456” into two new columns
  separate(range_str, into = c("peptide_start","peptide_end"), sep = "\\.\\.") %>%
  # convert to integer
  mutate(
     start = as.integer(peptide_start),
     end   = as.integer(peptide_end)) %>%
  select(Entry, Sequence, Length, peptide_start, peptide_end)

```

Getting signal info:

```{r}
signal_info_toxins <- clean_toxins %>%
  select(Entry, Signal.peptide) %>%
  distinct() %>%
  mutate(range_str = str_extract_all(Signal.peptide, "\\d+\\.\\.\\d+")) %>%
  unnest_longer(range_str) %>%
  separate(range_str, into = c("signal_start","signal_end"), sep = "\\.\\.") %>%
  mutate(
     start = as.integer(signal_start),
     end   = as.integer(signal_end)) %>%
  select(Entry, signal_start, signal_end)
```

Joining the peptide and signal info dataframes together.

```{r}
toxins_data_full <- clean_toxins_2 %>%
  left_join(signal_info_toxins)
```
Label the peptides that are right after the signal:

```{r}
right_after_signal_toxins <- toxins_data_full %>%
  mutate(peptide_after_signal = ifelse(as.numeric(signal_end) + 1 == peptide_start, TRUE, FALSE))
``` 

Now let's see the sequences where the start is 1 and the end is equal to the length. These are the mature peptide and we do not want these in the dataset.
```{r}
mature_peptides_toxins <- right_after_signal_toxins %>%
  filter(peptide_start == 1 & peptide_end == Length) # result: 938 mature peptide sequences
```

Now, let's look at the sequences where the start is 1, but the end is not equal to the length (these we want to keep because we can use the second cleavage site residues)
```{r}
semi_mature_peptides_toxins <- right_after_signal_toxins %>%
  filter(peptide_start == 1) %>%
  filter(peptide_end != Length) # result: only 9 of these 
```

I now want to save the unique sequences as a fasta file. 
```{r}
for_fasta_1 <- right_after_signal_toxins %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) %>% # filtering out mature peptides
  select(Entry, Sequence) %>%
  distinct(Entry, Sequence, .keep_all = TRUE)

# Create FASTA-formatted lines: ">ID" on one line, followed by the sequence on the next
fasta_lines <- apply(for_fasta_1, 1, function(row) {
  paste0(">", row["Entry"], "\n", row["Sequence"])
})

#Write these lines to a .fasta file
writeLines(fasta_lines, con = "../../data/toxin_sequences.fasta")
```

Let's count the final number of toxin sequences we will be working with:

```{r}
for_fasta_1 %>%
  select(Entry) %>%
  distinct() %>%
  dplyr::count()
```

Let's get final df with the 1685 seqs. 
```{r}
unique_toxins <- for_fasta_1 %>%
  select(Entry) %>%
  distinct() 

final_toxins_df <- right_after_signal_toxins %>%
  right_join(unique_toxins)

write_csv(final_toxins_df, file = "../../data/test_dataset_stats/final_toxin_dataset.csv")
  
```

Now I need save the positive dataset. I need to sample the cleavage sites. 

#----------------------------------------------------------------------------------------------
# Positive dataset
#----------------------------------------------------------------------------------------------

sampling the positive cleavage sites:
```{r}
final_cleavage_site_df_toxins <- right_after_signal_toxins %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) # filtering out mature peptides
```

Getting the positive data labels for the cleavage sites
```{r}
library(stringr)

positive_labels_toxins <- final_cleavage_site_df_toxins %>%
  # for each row, build a list of the 4 flanking positions
  rowwise() %>%
    mutate(
      # build a tiny tibble of flank positions + their labels
      flank = list(tibble::tibble(
        residue_number = c(as.numeric(peptide_start) - 2,
                           as.numeric(peptide_start) - 1,
                           as.numeric(peptide_end)   + 1,
                           as.numeric(peptide_end)   + 2),
        cleavage       = c("cleavage_1",   
                           "cleavage_1",
                           "cleavage_2",
                           "cleavage_2"),
        cleavage_pos       = c("first\nouter",   
                           "first\ninner",
                           "second\ninner",
                           "second\nouter")
      ))
    ) %>%
    unnest(cols = c(flank)) %>% 
    # drop any cleavage_1 when start == 1 or after_signal == TRUE
    filter(!(cleavage == "cleavage_1" & (peptide_start == 1 | peptide_after_signal))) %>%
    # drop any cleavage_2 when end == Length
    filter(!(cleavage == "cleavage_2" & peptide_end == Length)) %>%
    # also drop any positions that fall outside 1…Length
    filter(residue_number >= 1, residue_number <= Length) %>%
    # extract the amino acid at each flank position
    mutate(residue = str_sub(Sequence, residue_number, residue_number))
    
positive_labels_2_toxins <- positive_labels_toxins %>%
  select(Entry, Sequence, Length, residue, residue_number) %>%
  distinct()

positive_labels_save_toxins <- positive_labels_2_toxins %>%
  select(Entry, residue, residue_number) %>%
  ungroup() %>%
  distinct()

write_csv(positive_labels_save_toxins, "../../data/concept_groups/sequences/group_2/training/group_2_positive_toxins.csv")

```

Let's save the cleavage site sequence again. Since we did a lot of new filtering, we may have dropped some Entries.
```{r}

for_fasta_toxins <- positive_labels_save_toxins %>%
  select(Entry) %>%
  distinct() %>%
  inner_join(for_fasta_1)

# Create FASTA-formatted lines: ">ID" on one line, followed by the sequence on the next
fasta_lines <- apply(for_fasta_toxins, 1, function(row) {
  paste0(">", row["Entry"], "\n", row["Sequence"])
})

# Write these lines to a .fasta file
writeLines(fasta_lines, con = "../../data/cleavage_site_sequences_toxins.fasta")
```


Let's count the final number of toxin sequences we will be 
working with after sampling all the cleavage sites:

```{r} 
for_fasta_toxins %>%
  select(Entry) %>%
  distinct() %>%
  dplyr::count()
```
Now let's make the plot showing the distribution of amino acids within the cleavage sites in the toxin and non-toxin 
datasets. We expect the distributions to be about the same. 

We will need the following two dfs: "positive_labels_save_toxins" and "positive_labels_save"

```{r}
# 1) Add dataset column
neuro <- positive_labels_save %>% 
  mutate(dataset = "neuropeptide")
toxin <- positive_labels_save_toxins %>% 
  mutate(dataset = "toxin")

# 2) Bind together
combined <- bind_rows(neuro, toxin)

# Define the 20 standard amino acids
all_residues <- c("A","C","D","E","F",
                  "G","H","I","K","L",
                  "M","N","P","Q","R",
                  "S","T","V","W","Y")

# 3) Compute the order by descending neuropeptide counts
neuro_order <- combined %>%
  filter(dataset == "neuropeptide") %>%
  dplyr::count(residue) %>%
  arrange(desc(n)) %>%
  pull(residue)

# 4) Make the full x‐axis ordering: first the neuro_order, then any remaining AAs
residue_order <- c(neuro_order, setdiff(all_residues, neuro_order))

# 4) Build a counts dataframe, include zero‐counts so toxins line up
counts_toxin_neuro_df<- combined %>% 
  dplyr::count(dataset, residue, name = "count") %>% 
  # ensure every residue shows up in every dataset
  complete(dataset, residue = residue_order, fill = list(count = 0)) %>% 
  # apply the ordering
  mutate(
    residue = factor(residue, levels = residue_order),
    dataset = factor(dataset, levels = c("neuropeptide", "toxin"))
  )
```

```{r}

# 5) Plot: stacked facets, same x‐order, counts above bars
ggplot(counts_toxin_neuro_df, aes(x = residue, y = count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = count), vjust = -0.3, size = 3) +
  facet_wrap(~ dataset, ncol = 1, scales = "fixed") +
  scale_y_continuous(expand = expansion(mult = c(0,0.1))) +
  scale_x_discrete(drop = FALSE) +
  labs(
    x = "",
    y = "Count"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(vjust = 0.5),
    strip.text = element_text(face = "bold")
  )
```

Now I want to compare and report the number of repeats per peptide in the 
toxin and neuropeptide datasets. We expect that the toxin datasets have far less repeats than the neuropeptide dataset.

Let's first get a df with neuropeptides:
```{r}
for_binning_neuro <- right_after_signal %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) # filtering out mature peptides
```

the df "right_after_signal" is the dataset we have been working with. I will group the peptides by "Entry"
and I will count the number of rows (alternative peptides per Entry)

```{r}
alt_pep_counts_neuro <- for_binning_neuro %>%
  group_by(Sequence) %>%
  mutate(alt_peptides = n()) %>%
  ungroup()
```

Let's now get the df with toxins:
```{r}
for_binning_toxin <- right_after_signal_toxins %>%
  filter(!(peptide_start == 1 & peptide_end == Length)) # filtering out mature peptides
```

Counting the number of alternative peptides:
```{r}
alt_pep_counts_toxin <- for_binning_toxin %>%
  group_by(Sequence) %>%
  mutate(alt_peptides = n()) %>%
  ungroup()
```

Looking at the counts and the number of peptides per count:
```{r}
count_counts_neuro <- alt_pep_counts_neuro %>%
  group_by(alt_peptides) %>%
  dplyr::count() %>%
  ungroup()
```

```{r}
count_counts_toxin <- alt_pep_counts_toxin %>%
  group_by(alt_peptides) %>%
  dplyr::count() %>%
  ungroup()
```


The "positive_labels_save_toxins" df contains only the Entries that made it through the final filterings. When finding the distribution of repeats, I need to only use the entries that made it to the final dataset. 



*Finish making the repeat plots here*






#----------------------------------------------------------------------------------------------
# Negative dataset
#----------------------------------------------------------------------------------------------

Now we need to sample the negative dataset:
First we make a table that contains everything we want to exclude. 
Besides the cleavage sites, I also want to exclude the adjacent neighbors to the cleavage sites (just in case). 
```{r}
pos_labels_with_neighbors_toxins <- positive_labels_2_toxins %>%
  select(Entry, residue_number) %>%
  # add the -1 and +1 neighbors
  mutate(
    neighbor_minus = residue_number - 1,
    neighbor_plus  = residue_number + 1
  ) %>%
  pivot_longer(
    cols      = c(residue_number, neighbor_minus, neighbor_plus),
    values_to = "residue_number"
  ) %>%
  filter(residue_number >= 1) %>%       # drop any <1
  select(Entry, residue_number) %>%
  distinct()
```

Now, we need to count how many cleavage site resides (positives) we have per sequence/Entry. 
Remember that some sequences have multiple alternative cleavage sites.
```{r}
positives_count_toxins <- positive_labels_2_toxins %>%
  dplyr::count(Entry, name = "n_pos")
```


Each sequence is expanded to have one row per residue (will be used later)
I kept the sequence row as is to later extract the actual residue
```{r}
expanded_seqs <- positive_labels_2_toxins %>%
  select(Entry, Sequence, Length) %>%
  distinct() %>%
  uncount(Length) %>%                   # replicate each row 'Length' times
  group_by(Entry) %>%
  mutate(residue_number = row_number()) %>%
  ungroup() 
```

Next, we need to remove the excluded positions
```{r}
expanded_seqs <- expanded_seqs %>%
  anti_join(pos_labels_with_neighbors, by = c("Entry", "residue_number"))
```

Here we sample the negative dataset positions. (Try not to rerun this) each time it will resample
```{r}
negative_labels_toxins <- expanded_seqs %>%
  left_join(positives_count_toxins, by = "Entry") %>% # adding the count of positive dataset residues per Entry/sequence
  group_by(Entry) %>%
  sample_n(size = unique(n_pos)) %>%    # sample same number as positives
  ungroup() %>%
  mutate(
    residue = str_sub(Sequence, residue_number, residue_number) # str_sub(string, start, end)
  ) %>%
  select(Entry, residue, residue_number)
```

Saving the sampled negative data 
```{r}
write_csv(negative_labels_toxins, "../../data/concept_groups/sequences/group_2/training/group_2_negative_toxins.csv")

```

Making sure the sampling worked and we have equal number of positives. 
```{r}
# How many positives per Entry?
pos_counts_toxin <- positive_labels_save_toxins %>%
  dplyr::count(Entry, name = "n_pos")

# How many negatives did I actually get per Entry?
neg_counts_toxins <- negative_labels_toxins %>%
  dplyr::count(Entry, name = "n_neg")

# Join and look for any Entry where n_neg < n_pos
mismatches <- pos_counts_toxin %>%
  left_join(neg_counts_toxins, by = "Entry") %>%
  mutate(less_neg = ifelse(n_neg < n_pos, TRUE, FALSE))

mismatches          # see which ones and by how much
```

Now we need to get domain labels for the positive dataset in order to run the hybrid F1 score (modified) script in python.

```{r}
positive_labels_domain_toxin <- positive_labels_toxins %>%
  # define each window (domain) by Entry + cleavage site + flank type
  # within the same sequence, same peptide (start and end) and cleavage site, this is a single domain
  mutate(domain_string = paste(Entry, peptide_start, peptide_end, cleavage, sep = "_")) %>%
  group_by(domain_string) %>%
    # this gives each domain a unique integer
    mutate(domain_id = cur_group_id()) %>%
  ungroup() #%>%
  #select(-domain_string)
```

saving the new positive labels (for hybrid model)
```{r}
positive_labels_save_domain_toxin <- positive_labels_domain_toxin %>%
  select(Entry, residue, residue_number, domain_id) %>%
  distinct()

write_csv(positive_labels_save_domain_toxin, "../../data/group_2_positive_domain_toxin.csv")

```

#=================================================================================================
# Modified (hybrid) F1 scores
#=================================================================================================

Now we can calculate the F1 scores and load them here for the toxin dataset:

Let's look at the winning thresholds and F1 scores.
```{r}
winning_F1_toxins <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/F1_scores_hybrid/winning_thresholds_toxins.csv")
# winning F1 is 0.733 (feature 9047)
```

Now I need to compare the winning features in the neuropeptide dataset to the winning features in the toxin dataset and compare how many are shared between the two datasets and how many are unique to each dataset.

I have the following two datasets: "winnig_F1_toxins" and "winning_F1_modified"

```{r}
top50_toxins <- winning_F1_toxins %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 50) %>%
  mutate(dataset = "toxins")

top50_neuro <- winning_F1_modified %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 50) %>%
  mutate(dataset = "neuropeptides")

# combine into one data.frame
top50_neuro_toxin <- bind_rows(top50_toxins, top50_neuro)
```

Now I am going to plot the F1 scores per dataset:
```{r}
ggplot(top50_neuro_toxin, aes(x = dataset, y = winning_F1_score)) +
  geom_violin(trim = FALSE, fill = "lightblue") +
  geom_sina(size = 1, alpha = 0.6) +
  #stat_summary(fun = mean, geom = "crossbar", width = 0.5, color = "darkblue") +
  labs(
    y     = "Winning F1 Score"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size=12),
    axis.title.x = element_blank(),  # remove x-axis label
    axis.title.y = element_text(size=14)
  )
```


```{r}

set.seed(42)  # so the jitter is reproducible

# 1) turn your bin factor into numeric positions
# 2) add a jitter offset (the same for line & point)
top50_neuro_toxin <- top50_neuro_toxin %>%
  mutate(
    dataset      = factor(dataset, levels = c("neuropeptides","toxins")),
    dataset_num  = as.numeric(dataset),
    x_jitter     = dataset_num + runif(n(), -0.15, 0.15)
  )

```

```{r}
# 1) Flag how many bins each feature appears in
top50_flagged <- top50_neuro_toxin %>%
  add_count(feature, name = "n_bins") %>%
  mutate(
    share_group = factor(
      n_bins,
      levels = 1:2,
      labels = c(
        "Unique",
        "Shared"
      )
    )
  )
```

```{r}
library(ggrepel)

ggplot(top50_flagged, aes(group = feature)) +
  # 1) Orange “Unique” points (mapped to color)
  geom_point(
    data = filter(top50_flagged, share_group == "Unique"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 2) Dark-green “Shared” points (mapped to color)
  geom_point(
    data = filter(top50_flagged, share_group == "Shared"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 3) Labels for Shared in the neuropeptide dataset
  geom_text_repel(
    data = filter(top50_flagged,
                  share_group == "Shared",
                  dataset     == "neuropeptides"),  # make sure this matches your factor
    aes(x = x_jitter, y = winning_F1_score, label = feature),
    size          = 3,
    nudge_y       = 0.05,
    nudge_x       = -0.05,
    segment.color = "grey60",
    segment.size  = 0.3,
    max.overlaps  = Inf
  ) +
  # 4) Lines underneath
  geom_line(
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 0.4, alpha = 0.6
  ) +
  # 5) Point-style legend
  scale_color_manual(
    name   = "Feature sharing",
    values = c("Unique" = "orange", "Shared" = "darkgreen"),
    guide  = guide_legend(
      override.aes = list(
        linetype = 0,   # no line
        shape    = 16,  # solid circle
        size     = 4
      )
    )
  ) +
  # 6) X axis
  scale_x_continuous(
    breaks = 1:2,
    labels = levels(top50_flagged$dataset),
    name   = "Dataset",
    expand = expansion(add = 0.2)
  ) +
  labs(y = "Winning F1 Score") +
  theme_minimal() +
  theme(
    legend.position  = "right",
    legend.title     = element_blank(),
    legend.text     = element_text(size = 11),
    axis.title.x     = element_blank(),
    axis.text.x      = element_text(size = 12),
    axis.title.y     = element_text(size = 12)
  )

```
Now let's loot at the top 100 features. 

```{r}
top100_toxins <- winning_F1_toxins %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "toxins")

top100_neuro <- winning_F1_modified %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "neuropeptides")

# combine into one data.frame
top100_neuro_toxin <- bind_rows(top100_toxins, top100_neuro)
```

```{r}

set.seed(42)  # so the jitter is reproducible

# 1) turn your bin factor into numeric positions
# 2) add a jitter offset (the same for line & point)
top100_neuro_toxin <- top100_neuro_toxin %>%
  mutate(
    dataset      = factor(dataset, levels = c("neuropeptides","toxins")),
    dataset_num  = as.numeric(dataset),
    x_jitter     = dataset_num + runif(n(), -0.15, 0.15)
  )

```

```{r}
# 1) Flag how many bins each feature appears in
top100_flagged <- top100_neuro_toxin %>%
  add_count(feature, name = "n_bins") %>%
  mutate(
    share_group = factor(
      n_bins,
      levels = 1:2,
      labels = c(
        "Unique",
        "Shared"
      )
    )
  )
```

```{r}
library(ggrepel)

ggplot(top100_flagged, aes(group = feature)) +
  # 1) Orange “Unique” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Unique"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 2) Dark-green “Shared” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Shared"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 3) Labels for Shared in the neuropeptide dataset
  geom_text_repel(
    data = filter(top100_flagged,
                  share_group == "Shared",
                  dataset     == "neuropeptides"),  # make sure this matches your factor
    aes(x = x_jitter, y = winning_F1_score, label = feature),
    size          = 3,
    nudge_y       = 0.05,
    nudge_x       = -0.05,
    segment.color = "grey60",
    segment.size  = 0.3,
    max.overlaps  = Inf
  ) +
  # 4) Lines underneath
  geom_line(
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 0.4, alpha = 0.6
  ) +
  # 5) Point-style legend
  scale_color_manual(
    name   = "Feature sharing",
    values = c("Unique" = "orange", "Shared" = "darkgreen"),
    guide  = guide_legend(
      override.aes = list(
        linetype = 0,   # no line
        shape    = 16,  # solid circle
        size     = 4
      )
    )
  ) +
  # 6) X axis
  scale_x_continuous(
    breaks = 1:2,
    labels = levels(top100_flagged$dataset),
    name   = "Dataset",
    expand = expansion(add = 0.2)
  ) +
  labs(y = "Winning F1 Score") +
  theme_minimal() +
  theme(
    legend.position  = "right",
    legend.title     = element_blank(),
    legend.text     = element_text(size = 11),
    axis.title.x     = element_blank(),
    axis.text.x      = element_text(size = 12),
    axis.title.y     = element_text(size = 12)
  )

```

Now we are going to load the F1 scores for each homology group and see the difference:
```{r}
winning_F1_hom40 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/winning_thresholds_neuro_toxins_hom40.csv")
winning_F1_hom50 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/winning_thresholds_neuro_toxins_hom50.csv")
winning_F1_hom60 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/winning_thresholds_neuro_toxins_hom60.csv")
winning_F1_hom70 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/winning_thresholds_neuro_toxins_hom70.csv")
winning_F1_hom80 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/winning_thresholds_neuro_toxins_hom80.csv")
winning_F1_hom90 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/winning_thresholds_neuro_toxins_hom90.csv")
```

Let's combine them into a single df:
```{r}
winning_F1_hom40 <- winning_F1_hom40 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(homology = "40%")

winning_F1_hom50 <- winning_F1_hom50 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(homology = "50%")

winning_F1_hom60 <- winning_F1_hom60 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(homology = "60%")

winning_F1_hom70 <- winning_F1_hom70 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(homology = "70%")

winning_F1_hom80 <- winning_F1_hom80 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(homology = "80%")

winning_F1_hom90 <- winning_F1_hom90 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(homology = "90%")

winning_F1_hom_all <- rbind(winning_F1_hom40, winning_F1_hom50, winning_F1_hom60, winning_F1_hom70, winning_F1_hom80, winning_F1_hom90)
```

Now I am going to plot the F1 scores per homology set:
```{r}
ggplot(winning_F1_hom_all, aes(x = homology, y = winning_F1_score)) +
  geom_violin(trim = FALSE, fill = "lightblue") +
  geom_sina(size = 1, alpha = 0.6) +
  stat_summary(fun = mean, geom = "crossbar", width = 0.7, color = "darkblue") +
  labs(
    y     = "Winning F1 Score",
    x     = "Homology Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size=12),
    axis.title.y = element_text(size=14)
  )
```

Now let's take these features and look at how mahy features are shared across all homology groups. 

```{r}

set.seed(42)  # so the jitter is reproducible

# 1) turn your bin factor into numeric positions
# 2) add a jitter offset (the same for line & point)
top100_hom_groups <- winning_F1_hom_all %>%
  mutate(
    dataset      = factor(homology, levels = c("40%", "50%", "60%", "70%", "80%", "90%")),
    dataset_num  = as.numeric(dataset),
    x_jitter     = dataset_num + runif(n(), -0.15, 0.15)
  )

```

```{r}
# 1) Flag how many bins each feature appears in
top100_flagged <- top100_hom_groups %>%
  add_count(feature, name = "n_bins") %>%
  mutate(
    share_group = factor(
      n_bins,
      levels = 1:6,
      labels = c(
        "Unique",
        "Shared (2 bins)",
        "Shared (3 bins)",
        "Shared (4 bins)",
        "Shared (5 bins)",
        "Shared (6 bins)"
      )
    )
  )
```

```{r}
library(ggrepel)

ggplot(top100_flagged, aes(group = feature)) +
  # 1) Orange “Unique” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Unique"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 2, alpha = 1
  ) +
  geom_point(
    data = filter(top100_flagged, share_group == "Shared (2 bins)"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 1, alpha = 0.9
  ) +
  geom_point(
    data = filter(top100_flagged, share_group == "Shared (3 bins)"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 1, alpha = 0.9
  ) +
  geom_point(
    data = filter(top100_flagged, share_group == "Shared (4 bins)"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 1, alpha = 0.9
  ) +
  geom_point(
    data = filter(top100_flagged, share_group == "Shared (5 bins)"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 1, alpha = 0.9
  ) +
  geom_point(
    data = filter(top100_flagged, share_group == "Shared (6 bins)"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 1, alpha = 0.9
  ) +
  # 2) Labels for Shared in the neuropeptide dataset
  # geom_text_repel(
  #   data = filter(top100_flagged,
  #                 share_group == "Shared (6 bins)",
  #                 homology     == "40%"),  # make sure this matches your factor
  #   aes(x = x_jitter, y = winning_F1_score, label = feature),
  #   size          = 3,
  #   nudge_y       = 0.05,
  #   nudge_x       = -0.05,
  #   segment.color = "grey60",
  #   segment.size  = 0.3,
  #   max.overlaps  = Inf
  # ) +
  # 3) Lines underneath
  geom_line(
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 0.4, alpha = 0.6
  ) +
  # 4) Point-style legend
  scale_color_manual(
    name   = "Feature sharing",
    values = c("Unique" = "yellow", 
               "Shared (2 bins)" = "darkturquoise", 
               "Shared (3 bins)" = "orange", 
               "Shared (4 bins)" = "darkgreen",
               "Shared (5 bins)" = "maroon", 
               "Shared (6 bins)" = "darkblue"),
    guide  = guide_legend(
      override.aes = list(
        linetype = 0,   # no line
        shape    = 16,  # solid circle
        size     = 4
      )
    )
  ) +
  # 5) X axis
  scale_x_continuous(
    breaks = 1:6,
    labels = levels(top100_flagged$dataset),
    name   = "Homology Group",
    expand = expansion(add = 0.2)
  ) +
  labs(y = "Winning F1 Score") +
  theme_minimal() +
  theme(
    legend.position  = "right",
    legend.title     = element_blank(),
    legend.text     = element_text(size = 11),
    axis.title.x     = element_blank(),
    axis.text.x      = element_text(size = 12),
    axis.text.y      = element_text(size = 12),
    axis.title.y     = element_text(size = 12)
  )

```

```{r pie_chart_aligned, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggforce)

# 1) Compute slice fractions and “mid‐angle” rotated so 0 = 12 o’clock
levels_order <- c(
  "Unique",
  "Shared (2 bins)",
  "Shared (3 bins)",
  "Shared (4 bins)",
  "Shared (5 bins)",
  "Shared (6 bins)"
)

pie_data <- top100_flagged %>%
  distinct(feature, share_group) %>%
  dplyr::count(share_group) %>%
  mutate(
    share_group = factor(share_group, levels = levels_order),
    fraction    = n / sum(n)
  ) %>%
  arrange(share_group) %>%
  mutate(
    ymin    = cumsum(lag(fraction, default = 0)),
    ymax    = cumsum(fraction),
    # shift start to 12 o’clock by subtracting 1/4 circle (pi/2)
    midang  = pi/2 - 2 * pi * (ymin + fraction/2),
    pct     = fraction * 100,
    label   = paste0(n, " (", round(pct), "%)")
  )

total_features <- sum(pie_data$n)

# 2) Draw pie + outside labels at radius 1.3
ggplot(pie_data) +
  geom_arc_bar(
    aes(
      x0     = 0, y0 = 0,
      r0     = 0, r  = 1,
      amount = n,
      fill   = share_group
    ),
    stat  = "pie",
    color = "white",
    size  = 0.2
  ) +
  geom_text(
    aes(
      x     = 1.3 * cos(midang),
      y     = 1.3 * sin(midang),
      label = label
    ),
    size        = 4,
    color       = "black",
    inherit.aes = FALSE
  ) +
  annotate(
    "text",
    x     = 0, y     = -1.4,
    label = paste0("Total unique features: ", total_features),
    size  = 5,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Unique"            = "yellow",
      "Shared (2 bins)"   = "darkturquoise",
      "Shared (3 bins)"   = "orange",
      "Shared (4 bins)"   = "darkgreen",
      "Shared (5 bins)"   = "maroon",
      "Shared (6 bins)"   = "darkblue"
    ),
    guide = guide_legend(
      title        = NULL,
      override.aes = list(linetype = 0, shape = 16, size = 4)
    )
  ) +
  coord_fixed() +
  xlim(-1.5, 1.5) +
  ylim(-1.5, 1.5) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.text     = element_text(size = 11),
    plot.margin     = margin(10, 10, 10, 10)
  )

```

Let's check the number of unique features:
```{r}
winning_F1_hom_all %>%
  select(feature) %>%
  unique() %>%
  dplyr::count()
```

Now I need to save my F1 scores and ranking to use in my classifier models:

Now I want to save the top F1 scores to a csv file to use in my linear regression model. 
```{r}
top_scores_hom90 <- winning_F1_hom90 %>%
  arrange(desc(winning_F1_score)) %>%  # sort so highest F1 is first
  slice_head(n = 2000) %>%               # take the top 2000
  mutate(rank = row_number()) %>%       # assign rank 1 to the highest F1
  select(feature, rank)
```

saving to csv:
```{r}
write.csv(top_scores_hom90, file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/feature_ranking_hybrid_hom90.csv", row.names = FALSE)
```

```{r}
winning_F1_modified <- winning_F1_modified %>%
  arrange(desc(winning_F1_score)) %>%  # sort so highest F1 is first              # take the top 2000
  mutate(rank = row_number()) 
```

#============================================================================================================
Here, I want to make a plot between combo and neuro rankings 
and between combo and toxin rankings.

combo vs neuro:
```{r}
top100_toxins <- winning_F1_toxins %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "toxins")

top100_neuro <- winning_F1_modified %>%
  select(-rank) %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "neuropeptides")

top100_combo <- winning_F1_hom90 %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "combo")

# combine into one data.frame
top100_neuro_toxin <- bind_rows(top100_toxins, top100_neuro)
top100_neuro_combo <- bind_rows(top100_combo, top100_neuro)
top100_toxin_combo <- bind_rows(top100_combo, top100_toxins)
```

```{r}

set.seed(42)  # so the jitter is reproducible

# 1) turn your bin factor into numeric positions
# 2) add a jitter offset (the same for line & point)
top100_toxin_combo <- top100_toxin_combo %>%
  mutate(
    dataset      = factor(dataset, levels = c("combo","toxins")),
    dataset_num  = as.numeric(dataset),
    x_jitter     = dataset_num + runif(n(), -0.15, 0.15)
  )

```

```{r}
# 1) Flag how many bins each feature appears in
top100_flagged <- top100_toxin_combo %>%
  add_count(feature, name = "n_bins") %>%
  mutate(
    share_group = factor(
      n_bins,
      levels = 1:2,
      labels = c(
        "Unique",
        "Shared"
      )
    )
  )
```

```{r}
library(ggrepel)

ggplot(top100_flagged, aes(group = feature)) +
  # 1) Orange “Unique” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Unique"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 2) Dark-green “Shared” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Shared"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 4) Lines underneath
  geom_line(
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 0.4, alpha = 0.6
  ) +
  # 5) Point-style legend
  scale_color_manual(
    name   = "Feature sharing",
    values = c("Unique" = "orange", "Shared" = "darkgreen"),
    guide  = guide_legend(
      override.aes = list(
        linetype = 0,   # no line
        shape    = 16,  # solid circle
        size     = 4
      )
    )
  ) +
  # 6) X axis
  scale_x_continuous(
    breaks = 1:2,
    labels = levels(top100_flagged$dataset),
    name   = "Dataset",
    expand = expansion(add = 0.2)
  ) +
  labs(y = "Winning F1 Score") +
  theme_minimal() +
  theme(
    legend.position  = "right",
    legend.title     = element_blank(),
    legend.text     = element_text(size = 11),
    axis.title.x     = element_blank(),
    axis.text.x      = element_text(size = 12),
    axis.title.y     = element_text(size = 12)
  )

```

```{r}
# 1) Compute slice fractions and “mid‐angle” rotated so 0 = 12 o’clock
levels_order <- c(
  "Unique",
  "Shared"
)

pie_data <- top100_flagged %>%
  distinct(feature, share_group) %>%
  dplyr::count(share_group) %>%
  mutate(
    share_group = factor(share_group, levels = levels_order),
    fraction    = n / sum(n)
  ) %>%
  arrange(share_group) %>%
  mutate(
    ymin    = cumsum(lag(fraction, default = 0)),
    ymax    = cumsum(fraction),
    # shift start to 12 o’clock by subtracting 1/4 circle (pi/2)
    midang  = pi/2 - 2 * pi * (ymin + fraction/2),
    pct     = fraction * 100,
    label   = paste0(n, " (", round(pct), "%)")
  )

total_features <- sum(pie_data$n)

# 2) Draw pie + outside labels at radius 1.3
ggplot(pie_data) +
  geom_arc_bar(
    aes(
      x0     = 0, y0 = 0,
      r0     = 0, r  = 1,
      amount = n,
      fill   = share_group
    ),
    stat  = "pie",
    color = "white",
    size  = 0.2
  ) +
  geom_text(
    aes(
      x     = 1.3 * cos(midang),
      y     = 1.3 * sin(midang),
      label = label
    ),
    size        = 4,
    color       = "black",
    inherit.aes = FALSE
  ) +
  annotate(
    "text",
    x     = 0, y     = -1.4,
    label = paste0("Total features: ", total_features),
    size  = 5,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Unique"            = "orange",
      "Shared"   = "darkgreen"
    ),
    guide = guide_legend(
      title        = NULL,
      override.aes = list(linetype = 0, shape = 16, size = 4)
    )
  ) +
  coord_fixed() +
  xlim(-1.5, 1.5) +
  ylim(-1.5, 1.5) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.text     = element_text(size = 11),
    plot.margin     = margin(10, 10, 10, 10)
  )

```

# Now I need to repeat the analysis, only with the graphpart data (folds 0-8)

```{r}
winning_F1_folds <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/F1_scores_graphpart/winning_thresholds_folds_0-8.csv")
```

#------------------------------------------------------------------------------------------------------------------------------------------
# Here, I want to make a plot between combo and neuro rankings 
# and between combo and toxin rankings.
#------------------------------------------------------------------------------------------------------------------------------------------

combo vs neuro:
```{r}
top100_toxins <- winning_F1_toxins %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "toxins")

top100_neuro <- winning_F1_modified %>%
  select(-rank) %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "neuropeptides")

top100_folds <- winning_F1_folds %>%
  arrange(desc(winning_F1_score)) %>%
  slice_head(n = 100) %>%
  mutate(dataset = "combo\n(folds)")

# combine into one data.frame
top100_neuro_toxin <- bind_rows(top100_toxins, top100_neuro)
top100_neuro_combo <- bind_rows(top100_folds, top100_neuro)
top100_toxin_combo <- bind_rows(top100_folds, top100_toxins)
```

```{r}

set.seed(42)  # so the jitter is reproducible

# 1) turn your bin factor into numeric positions
# 2) add a jitter offset (the same for line & point)
top100_neuro_combo <- top100_neuro_combo %>%
  mutate(
    dataset      = factor(dataset, levels = c("combo\n(folds)","neuropeptides")),
    dataset_num  = as.numeric(dataset),
    x_jitter     = dataset_num + runif(n(), -0.15, 0.15)
  )

```

```{r}
# 1) Flag how many bins each feature appears in
top100_flagged <- top100_neuro_combo %>%
  add_count(feature, name = "n_bins") %>%
  mutate(
    share_group = factor(
      n_bins,
      levels = 1:2,
      labels = c(
        "Unique",
        "Shared"
      )
    )
  )
```

```{r}
library(ggrepel)

ggplot(top100_flagged, aes(group = feature)) +
  # 1) Orange “Unique” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Unique"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 2) Dark-green “Shared” points (mapped to color)
  geom_point(
    data = filter(top100_flagged, share_group == "Shared"),
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 4, alpha = 0.9
  ) +
  # 4) Lines underneath
  geom_line(
    aes(x = x_jitter, y = winning_F1_score, color = share_group),
    size  = 0.4, alpha = 0.6
  ) +
  # 5) Point-style legend
  scale_color_manual(
    name   = "Feature sharing",
    values = c("Unique" = "orange", "Shared" = "darkgreen"),
    guide  = guide_legend(
      override.aes = list(
        linetype = 0,   # no line
        shape    = 16,  # solid circle
        size     = 4
      )
    )
  ) +
  # 6) X axis
  scale_x_continuous(
    breaks = 1:2,
    labels = levels(top100_flagged$dataset),
    name   = "Dataset",
    expand = expansion(add = 0.2)
  ) +
  labs(y = "Winning F1 Score") +
  theme_minimal() +
  theme(
    legend.position  = "right",
    legend.title     = element_blank(),
    legend.text     = element_text(size = 11),
    axis.title.x     = element_blank(),
    axis.text.x      = element_text(size = 12),
    axis.title.y     = element_text(size = 12)
  )
```

```{r}
# 1) Compute slice fractions and “mid‐angle” rotated so 0 = 12 o’clock
levels_order <- c(
  "Unique",
  "Shared"
)

pie_data <- top100_flagged %>%
  distinct(feature, share_group) %>%
  dplyr::count(share_group) %>%
  mutate(
    share_group = factor(share_group, levels = levels_order),
    fraction    = n / sum(n)
  ) %>%
  arrange(share_group) %>%
  mutate(
    ymin    = cumsum(lag(fraction, default = 0)),
    ymax    = cumsum(fraction),
    # shift start to 12 o’clock by subtracting 1/4 circle (pi/2)
    midang  = pi/2 - 2 * pi * (ymin + fraction/2),
    pct     = fraction * 100,
    label   = paste0(n, " (", round(pct), "%)")
  )

total_features <- sum(pie_data$n)

# 2) Draw pie + outside labels at radius 1.3
ggplot(pie_data) +
  geom_arc_bar(
    aes(
      x0     = 0, y0 = 0,
      r0     = 0, r  = 1,
      amount = n,
      fill   = share_group
    ),
    stat  = "pie",
    color = "white",
    size  = 0.2
  ) +
  geom_text(
    aes(
      x     = 1.3 * cos(midang),
      y     = 1.3 * sin(midang),
      label = label
    ),
    size        = 4,
    color       = "black",
    inherit.aes = FALSE
  ) +
  annotate(
    "text",
    x     = 0, y     = -1.4,
    label = paste0("Total features: ", total_features),
    size  = 5,
    fontface = "bold"
  ) +
  scale_fill_manual(
    values = c(
      "Unique"            = "orange",
      "Shared"   = "darkgreen"
    ),
    guide = guide_legend(
      title        = NULL,
      override.aes = list(linetype = 0, shape = 16, size = 4)
    )
  ) +
  coord_fixed() +
  xlim(-1.5, 1.5) +
  ylim(-1.5, 1.5) +
  theme_void() +
  theme(
    legend.position = "right",
    legend.text     = element_text(size = 11),
    plot.margin     = margin(10, 10, 10, 10)
  )

```
# Getting feature rank file for folds 0-8


```{r}
winning_F1_folds_40 <- read.csv(file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/F1_scores_hybrid/F1_scores_graphpart_40/winning_thresholds_folds_0-8.csv")
```

```{r}
top_scores_F1_folds <- winning_F1_folds_40 %>%
  arrange(desc(winning_F1_score)) %>%  # sort so highest F1 is first
  slice_head(n = 2000) %>%               # take the top 2000
  mutate(rank = row_number()) %>%       # assign rank 1 to the highest F1
  select(feature, rank)
```

saving to csv:
```{r}
write.csv(top_scores_F1_folds, file = "/Volumes/T7 Shield/toxin_dataset/toxins_and_neuro/feature_ranking_folds_0-8_40.csv", row.names = FALSE)
```

















